### 2016-05-02 law-x.tex
Style files etc added

%
% File acl2016.tex
%
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% Based on the style files for ACL-2013, which were, in turn,
%% Based on the style files for ACL-2012, which were, in turn,
%% based on the style files for ACL-2011, which were, in turn,
%% based on the style files for ACL-2010, which were, in turn,
%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
%% based on the style files for EACL-2009 and IJCNLP-2008...

%% Based on the style files for EACL 2006 by
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt]{article}
\usepackage{acl2016}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage{apacite}

%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B{\sc ib}\TeX}

\title{Instructions for ACL-2016 Proceedings}

\author{First Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain} \\\And
  Second Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
  This document contains the instructions for preparing a camera-ready
  manuscript for the proceedings of ACL-2016. The document itself
  conforms to its own specifications, and is therefore an example of
  what your manuscript should look like. These instructions should be
  used for both papers submitted for review and for final versions of
  accepted papers.  Authors are asked to conform to all the directions
  reported in this document.
\end{abstract}

\section{Credits}

This document has been adapted from the instructions for earlier ACL
and NAACL proceedings, including those for NAACL-2016 by Margaret
Mitchell, ACL-2012 by Maggie Li and Michael
White, those from ACL-2010 by Jing-Shing Chang and Philipp Koehn,
those for ACL-2008 by Johanna D. Moore, Simone Teufel, James Allan,
and Sadaoki Furui, those for ACL-2005 by Hwee Tou Ng and Kemal
Oflazer, those for ACL-2002 by Eugene Charniak and Dekang Lin, and
earlier ACL and EACL formats. Those versions were written by several
people, including John Chen, Henry S. Thompson and Donald
Walker. Additional elements were taken from the formatting
instructions of the {\em International Joint Conference on Artificial
  Intelligence} and the \emph{Conference on Computer Vision and
Pattern Recognition}.

\section{Introduction}

The following instructions are directed to authors of papers submitted
to ACL-2016 or accepted for publication in its proceedings. All
authors are required to adhere to these specifications. Authors are
required to provide a Portable Document Format (PDF) version of their
papers. \textbf{The proceedings are designed for printing on A4
paper.}

We will make more detailed instructions available at
\url{acl2016.org/index.php?article_id=9}. Please check this website
regularly.

\section{General Instructions}

Manuscripts must be in two-column format.  Exceptions to the
two-column format include the title, authors' names and complete
addresses, which must be centered at the top of the first page, and
any full-width figures or tables (see the guidelines in
Subsection~\ref{ssec:first}). {\bf Type single-spaced.}  Start all
pages directly under the top margin. See the guidelines later
regarding formatting the first page.  The manuscript should be
printed single-sided and its length
should not exceed the maximum page limit described in Section~\ref{sec:length}.
Do not number the pages.

By uncommenting {\small\verb|\aclfinalcopy|} at the top of this
 document, it will compile to produce an example of the camera-ready formatting; by leaving it commented out, the document will be anonymized for initial submission.  When you first create your submission on softconf, please fill in your submitted paper ID where {\small\verb|***|} appears in the {\small\verb|\def\aclpaperid{***}|} definition at the top.

The review process is double-blind, so do not include any author information (names, addresses) when submitting a paper for review.
However, you should maintain space for names and addresses so that they will fit in the final (accepted) version.  The ACL 2016 \LaTeX\ style will create a titlebox space of 2.5in for you when {\small\verb|\aclfinalcopy|} is commented out.

\subsection{The Ruler}
The ACL 2016 style defines a printed ruler which should be presented in the
version submitted for review.  The ruler is provided in order that
reviewers may comment on particular lines in the paper without
circumlocution.  If you are preparing a document without the provided
style files, please arrange for an equivalent ruler to
appear on the final output pages.  The presence or absence of the ruler
should not change the appearance of any other content on the page.  The
camera ready copy should not contain a ruler. (\LaTeX\ users may uncomment
the {\small\verb|\aclfinalcopy|} command in the document preamble.)

Reviewers: note that the ruler measurements do not align well with
lines in the paper --- this turns out to be very difficult to do well
when the paper contains many figures and equations, and, when done,
looks ugly. In most cases one would expect that the approximate
location will be adequate, altough you can also use fractional
references (e.g., the first paragraph on this page ends at mark $114.5$),
although in most cases one would expect that the approximate location
will be adequate.

\subsection{Electronically-available resources}

ACL provides this description in \LaTeX2e{} ({\small\tt acl2016.tex}) and PDF
format ({\small\tt acl2016.pdf}), along with the \LaTeX2e{} style file used to
format it ({\small\tt acl2016.sty}) and an ACL bibliography style ({\small\tt acl2016.bst})
and example bibliography ({\small\tt acl2016.bib}).
These files are all available at
{\small\tt acl2016.org/index.php?article\_id=9}. We
strongly recommend the use of these style files, which have been
appropriately tailored for the ACL 2016 proceedings.

\subsection{Format of Electronic Manuscript}
\label{sect:pdf}

For the production of the electronic manuscript you must use Adobe's
Portable Document Format (PDF). PDF files are usually produced from
\LaTeX\ using the \textit{pdflatex} command. If your version of
\LaTeX\ produces Postscript files, you can convert these into PDF
using \textit{ps2pdf} or \textit{dvipdf}. On Windows, you can also use
Adobe Distiller to generate PDF.

Please make sure that your PDF file includes all the necessary fonts
(especially tree diagrams, symbols, and fonts with Asian
characters). When you print or create the PDF file, there is usually
an option in your printer setup to include none, all or just
non-standard fonts.  Please make sure that you select the option of
including ALL the fonts. \textbf{Before sending it, test your PDF by
  printing it from a computer different from the one where it was
  created.} Moreover, some word processors may generate very large PDF
files, where each page is rendered as an image. Such images may
reproduce poorly. In this case, try alternative ways to obtain the
PDF. One way on some systems is to install a driver for a postscript
printer, send your document to the printer specifying ``Output to a
file'', then convert the file to PDF.

It is of utmost importance to specify the \textbf{A4 format} (21 cm
x 29.7 cm) when formatting the paper. When working with
{\tt dvips}, for instance, one should specify {\tt -t a4}.
Or using the command \verb|\special{papersize=210mm,297mm}| in the latex
preamble (directly below the \verb|\usepackage| commands). Then using
{\tt dvipdf} and/or {\tt pdflatex} which would make it easier for some.

Print-outs of the PDF file on A4 paper should be identical to the
hardcopy version. If you cannot meet the above requirements about the
production of your electronic submission, please contact the
publication chairs as soon as possible.

\subsection{Layout}
\label{ssec:layout}

Format manuscripts two columns to a page, in the manner these
instructions are formatted. The exact dimensions for a page on A4
paper are:

\begin{itemize}
\item Left and right margins: 2.5 cm
\item Top margin: 2.5 cm
\item Bottom margin: 2.5 cm
\item Column width: 7.7 cm
\item Column height: 24.7 cm
\item Gap between columns: 0.6 cm
\end{itemize}

\noindent Papers should not be submitted on any other paper size.
 If you cannot meet the above requirements about the production of
 your electronic submission, please contact the publication chairs
 above as soon as possible.

\subsection{Fonts}

For reasons of uniformity, Adobe's {\bf Times Roman} font should be
used. In \LaTeX2e{} this is accomplished by putting

\begin{quote}
\begin{verbatim}
\usepackage{times}
\usepackage{latexsym}
\end{verbatim}
\end{quote}
in the preamble. If Times Roman is unavailable, use {\bf Computer
  Modern Roman} (\LaTeX2e{}'s default).  Note that the latter is about
  10\% less dense than Adobe's Times Roman font.

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|rl|}
\hline \bf Type of Text & \bf Font Size & \bf Style \\ \hline
paper title & 15 pt & bold \\
author names & 12 pt & bold \\
author affiliation & 12 pt & \\
the word ``Abstract'' & 12 pt & bold \\
section titles & 12 pt & bold \\
document text & 11 pt  &\\
captions & 11 pt & \\
abstract text & 10 pt & \\
bibliography & 10 pt & \\
footnotes & 9 pt & \\
\hline
\end{tabular}
\end{center}
\caption{\label{font-table} Font guide. }
\end{table}

\subsection{The First Page}
\label{ssec:first}

Center the title, author's name(s) and affiliation(s) across both
columns. Do not use footnotes for affiliations. Do not include the
paper ID number assigned during the submission process. Use the
two-column format only when you begin the abstract.

{\bf Title}: Place the title centered at the top of the first page, in
a 15-point bold font. (For a complete guide to font sizes and styles,
see Table~\ref{font-table}) Long titles should be typed on two lines
without a blank line intervening. Approximately, put the title at 2.5
cm from the top of the page, followed by a blank line, then the
author's names(s), and the affiliation on the following line. Do not
use only initials for given names (middle initials are allowed). Do
not format surnames in all capitals (e.g., use ``Mitchell not
``MITCHELL'').  Do not format title and section headings in all
capitals as well except for proper names (such as ``BLEU'') that are
conventionally in all capitals.  The affiliation should contain the
author's complete address, and if possible, an electronic mail
address. Start the body of the first page 7.5 cm from the top of the
page.

The title, author names and addresses should be completely identical
to those entered to the electronical paper submission website in order
to maintain the consistency of author information among all
publications of the conference. If they are different, the publication
chairs may resolve the difference without consulting with you; so it
is in your own interest to double-check that the information is
consistent.

{\bf Abstract}: Type the abstract at the beginning of the first
column. The width of the abstract text should be smaller than the
width of the columns for the text in the body of the paper by about
0.6 cm on each side. Center the word {\bf Abstract} in a 12 point bold
font above the body of the abstract. The abstract should be a concise
summary of the general thesis and conclusions of the paper. It should
be no longer than 200 words. The abstract text should be in 10 point font.

{\bf Text}: Begin typing the main body of the text immediately after
the abstract, observing the two-column format as shown in
the present document. Do not include page numbers.

{\bf Indent} when starting a new paragraph. Use 11 points for text and
subsection headings, 12 points for section headings and 15 points for
the title.

\begin{table}
\centering
\small
\begin{tabular}{cc}
\begin{tabular}{|l|l|}
\hline
{\bf Command} & {\bf Output}\\\hline
\verb|{\"a}| & {\"a} \\
\verb|{\^e}| & {\^e} \\
\verb|{\`i}| & {\`i} \\
\verb|{\.I}| & {\.I} \\
\verb|{\o}| & {\o} \\
\verb|{\'u}| & {\'u}  \\
\verb|{\aa}| & {\aa}  \\\hline
\end{tabular} &
\begin{tabular}{|l|l|}
\hline
{\bf Command} & {\bf  Output}\\\hline
\verb|{\c c}| & {\c c} \\
\verb|{\u g}| & {\u g} \\
\verb|{\l}| & {\l} \\
\verb|{\~n}| & {\~n} \\
\verb|{\H o}| & {\H o} \\
\verb|{\v r}| & {\v r} \\
\verb|{\ss}| & {\ss} \\\hline
\end{tabular}
\end{tabular}
\caption{Example commands for accented characters, to be used in, e.g., \BibTeX\ names.}\label{tab:accents}
\end{table}

\subsection{Sections}

{\bf Headings}: Type and label section and subsection headings in the
style shown on the present document.  Use numbered sections (Arabic
numerals) in order to facilitate cross references. Number subsections
with the section number and the subsection number separated by a dot,
in Arabic numerals. Do not number subsubsections.

{\bf Citations}: Citations within the text appear in parentheses
as~\cite{Gusfield:97} or, if the author's name appears in the text
itself, as Gusfield~\shortcite{Gusfield:97}.
Using the provided \LaTeX\ style, the former is accomplished using
{\small\verb|\cite|} and the latter with {\small\verb|\shortcite|} or {\small\verb|\newcite|}.  Collapse multiple citations as in~\cite{Gusfield:97,Aho:72}; this is accomplished with the provided style using commas within the {\small\verb|\cite|} command, e.g., {\small\verb|\cite{Gusfield:97,Aho:72}|}.
Append lowercase letters to the year in cases of ambiguities.
 Treat double authors as
in~\cite{Aho:72}, but write as in~\cite{Chandra:81} when more than two
authors are involved. Collapse multiple citations as
in~\cite{Gusfield:97,Aho:72}. Also refrain from using full citations
as sentence constituents.

\penalty -5000

We suggest that instead of
\begin{quote}
  ``\cite{Gusfield:97} showed that ...''
\end{quote}
you use
\begin{quote}
``Gusfield \shortcite{Gusfield:97}   showed that ...''
\end{quote}

If you are using the provided \LaTeX{} and Bib\TeX{} style files, you
can use the command \verb|\newcite| to get ``author (year)'' citations.

As reviewing will be double-blind, the submitted version of the papers
should not include the authors' names and affiliations. Furthermore,
self-references that reveal the author's identity, e.g.,
\begin{quote}
``We previously showed \cite{Gusfield:97} ...''
\end{quote}
should be avoided. Instead, use citations such as
\begin{quote}
``Gusfield \shortcite{Gusfield:97}
previously showed ... ''
\end{quote}

\textbf{Please do not use anonymous citations} and do not include
acknowledgements when submitting your papers. Papers that do not
conform to these requirements may be rejected without review.

\textbf{References}: Gather the full set of references together under
the heading {\bf References}; place the section before any Appendices,
unless they contain references. Arrange the references alphabetically
by first author, rather than by order of occurrence in the text.
Provide as complete a citation as possible, using a consistent format,
such as the one for {\em Computational Linguistics\/} or the one in the
{\em Publication Manual of the American
Psychological Association\/}~\cite{APA:83}.  Use of full names for
authors rather than initials is preferred.  A list of abbreviations
for common computer science journals can be found in the ACM
{\em Computing Reviews\/}~\cite{ACM:83}.

The \LaTeX{} and Bib\TeX{} style files provided roughly fit the
American Psychological Association format, allowing regular citations,
short citations and multiple citations as described above.

{\bf Appendices}: Appendices, if any, directly follow the text and the
references (but see above).  Letter them in sequence and provide an
informative title: {\bf Appendix A. Title of Appendix}.

\subsection{Footnotes}

{\bf Footnotes}: Put footnotes at the bottom of the page and use 9
points text. They may be numbered or referred to by asterisks or other
symbols.\footnote{This is how a footnote should appear.} Footnotes
should be separated from the text by a line.\footnote{Note the line
separating the footnotes from the text.}

\subsection{Graphics}

{\bf Illustrations}: Place figures, tables, and photographs in the
paper near where they are first discussed, rather than at the end, if
possible.  Wide illustrations may run across both columns.  Color
illustrations are discouraged, unless you have verified that
they will be understandable when printed in black ink.

{\bf Captions}: Provide a caption for every illustration; number each one
sequentially in the form:  ``Figure 1. Caption of the Figure.'' ``Table 1.
Caption of the Table.''  Type the captions of the figures and
tables below the body, using 11 point text.

\subsection{Accessibility}
\label{ssec:accessibility}

In an effort to accommodate the color-blind (as well as those printing
to paper), grayscale readability for all accepted papers will be
encouraged.  Color is not forbidden, but authors should ensure that
tables and figures do not rely solely on color to convey critical
distinctions.
Here we give a simple criterion on your colored figures, if your paper has to be printed in black and white, then you must assure that every curves or points in your figures can be still clearly distinguished.

\section{XML conversion and supported \LaTeX\ packages}

Following ACL 2014 we will also we will attempt to automatically convert
your \LaTeX\ source files to publish papers in machine-readable
XML with semantic markup in the ACL Anthology, in addition to the
traditional PDF format.  This will allow us to create, over the next
few years, a growing corpus of scientific text for our own future research,
and picks up on recent initiatives on converting ACL papers from earlier
years to XML.

We encourage you to submit a ZIP file of your \LaTeX\ sources along
with the camera-ready version of your paper. We will then convert them
to XML automatically, using the LaTeXML tool
(\url{http://dlmf.nist.gov/LaTeXML}). LaTeXML has \emph{bindings} for
a number of \LaTeX\ packages, including the ACL 2016 stylefile. These
bindings allow LaTeXML to render the commands from these packages
correctly in XML. For best results, we encourage you to use the
packages that are officially supported by LaTeXML, listed at
\url{http://dlmf.nist.gov/LaTeXML/manual/included.bindings}

\section{Translation of non-English Terms}

It is also advised to supplement non-English characters and terms
with appropriate transliterations and/or translations
since not all readers understand all such characters and terms.
Inline transliteration or translation can be represented in
the order of: original-form transliteration ``translation''.

\section{Length of Submission}
\label{sec:length}

The ACL 2016 main conference accepts submissions of long papers and
short papers.
 Long papers may consist of up to eight (8) pages of
content plus unlimited pages for references. Upon acceptance, final
versions of long papers will be given one additional page (up to 9
pages with unlimited pages for references) so that reviewers' comments
can be taken into account. Short papers may consist of up to four (4)
pages of content, plus unlimited pages for references. Upon
acceptance, short papers will be given five (5) pages in the
proceedings and unlimited pages for references.

For both long and short papers, all illustrations and tables that are part
of the main text must be accommodated within these page limits, observing
the formatting instructions given in the present document. Supplementary
material in the form of appendices does not count towards the page limit.

However, note that supplementary material should be supplementary
(rather than central) to the paper, and that reviewers may ignore
supplementary material when reviewing the paper (see Appendix
\ref{sec:supplemental}). Papers that do not conform to the specified
length and formatting requirements are subject to be rejected without
review.

Workshop chairs may have different rules for allowed length and
whether supplemental material is welcome. As always, the respective
call for papers is the authoritative source.

\section*{Acknowledgments}

The acknowledgments should go immediately before the references.  Do
not number the acknowledgments section. Do not include this section
when submitting your paper for review.

% include your own bib file like this:
\bibliographystyle{apacite}
\bibliography{references,acl2016}

\appendix

\section{Supplemental Material}
\label{sec:supplemental}
ACL 2016 also encourages the submission of supplementary material
to report preprocessing decisions, model parameters, and other details
necessary for the replication of the experiments reported in the
paper. Seemingly small preprocessing decisions can sometimes make
a large difference in performance, so it is crucial to record such
decisions to precisely characterize state-of-the-art methods.

Nonetheless, supplementary material should be supplementary (rather
than central) to the paper. It may include explanations or details
of proofs or deriations that do not fit into the paper, lists of
features or feature tempates, sample inputs and outputs for a system,
pseudo-code or source code, and data. (Source code and data should
be separate uploads, rather than part of the paper).

The paper should not rely on the supplementary material: while the paper
may refer to and cite the supplementary material will be available to the
reviewers, they will not be asked to review the
supplementary material.

Appendices (i.e. supplementary material in the form of proofs, tables,
or pseudo-code) should come after the references, as shown here. Use
\verb|\appendix| before any appendix section to switch the section
numbering over to letters.

\section{Multiple Appendices}
\dots can be gotten by using more than one section. We hope you won't
need that.

\end{document}

### 2016-05-02 submission.tex
LAW workshop paper 2016

%
% File acl2016.tex
%
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% Based on the style files for ACL-2013, which were, in turn,
%% Based on the style files for ACL-2012, which were, in turn,
%% based on the style files for ACL-2011, which were, in turn, 
%% based on the style files for ACL-2010, which were, in turn, 
%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
%% based on the style files for EACL-2009 and IJCNLP-2008...

%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt]{article}
\usepackage{acl2016}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}

%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B{\sc ib}\TeX}

%% Greit å vite:

%% Siteringer:
%% Med parentes rundt hele: \cite{Gusfield:97} 
%% Eller med parentes kun rundt år:  \newcite{Gusfield:97} 
%% Flere på en gang: \cite{Gusfield:97,Aho:72}

%% Fra instruksjonene, om \aclfinalcopy :

%% By uncommenting {\small\verb|\aclfinalcopy|} at the top of this document, it
%% will compile to produce an example of the camera-ready formatting; by leaving
%% it commented out, the document will be anonymized for initial submission.  When
%% you first create your submission on softconf, please fill in your submitted
%% paper ID where {\small\verb|***|} appears in the
%% {\small\verb|\def\aclpaperid{***}|} definition at the top.

%% The review process is double-blind, so do not include any author information
%% (names, addresses) when submitting a paper for review.  However, you should
%% maintain space for names and addresses so that they will fit in the final
%% (accepted) version.  The ACL 2016 \LaTeX\ style will create a titlebox space of
%% 2.5in for you when {\small\verb|\aclfinalcopy|} is commented out.

%% \subsection{The Ruler}
%% The ACL 2016 style defines a printed ruler which should be presented in the
%% version submitted for review.  The ruler is provided in order that
%% reviewers may comment on particular lines in the paper without
%% circumlocution.  If you are preparing a document without the provided
%% style files, please arrange for an equivalent ruler to
%% appear on the final output pages.  The presence or absence of the ruler
%% should not change the appearance of any other content on the page.  The
%% camera ready copy should not contain a ruler. (\LaTeX\ users may uncomment
%% the {\small\verb|\aclfinalcopy|} command in the document preamble.)

\title{The greatest PoS tags, just the best really}

\author{First Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  {\tt email@domain} \\\And
  Second Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  {\tt email@domain}  \\\And
  Third Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  {\tt email@domain} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}

\end{abstract}

%\section{Credits}

\section{Introduction}

% \section*{Acknowledgments}

\bibliography{acl2016}
\bibliographystyle{acl2016}

% \appendix
%\section{Supplemental Material}
%\label{sec:supplemental}

\end{document}

### 2016-05-02 submission-petter.tex
Additions to experimental setup and data

--- 2016-05-02 submission.tex
@@
 \usepackage{latexsym}
+\usepackage{booktabs}
 
@@
 % in the camera-ready version and ask you to change it back.
-
-\newcommand\BibTeX{B{\sc ib}\TeX}
 
@@
 \begin{abstract}
-
+Abstraction ensues.
 \end{abstract}
@@
 \section{Introduction}
+\label{sec:intro}
+
+\section{Previous/Related work}
+\label{sec:prev}
+
+Megyesi, Swedish
+
+MacKinlay, English
+
+OBT, Norwegian
+
+\section{Data}
+\label{sec:data}
+\subsection{The Norwegian Dependency Treebank}
+The newly developed Norwegian Dependency Treebank (NDT) \cite{Sol:Skj:Ovr:14}
+is the first publicly available treebank for Norwegian. The treebank was used
+for training and evaluation in our experiments. The tag set consists of 12
+morphosyntactic PoS tags, with 7 additional tags for punctuation and symbols.
+
+\paragraph{PoS Tag Set}
+\paragraph{Dependency Relations}
+
+\section{Experimental Setup}
+\label{sec:setup}
+
+\paragraph{Tagger}
+TnT \cite{Bra:00}
+
+\paragraph{Parser}
+In choosing a syntactic parser for our experiments, we considered previous work
+on dependency parsing of Norwegian, specifically that of \cite{Sol:Skj:Ovr:14}.
+They found the Mate parser \cite{Boh:10} to be the most successful parser for
+the parsing of NDT. Recent dependency parser comparisons \cite{Cho:Tet:Ste:15}
+also showed that Mate performed very well on parsing of the English portion of
+the OntoNotes 5 corpus, beating a range of contemporary state-of-the-art
+parsers.
+
+\paragraph{Tag Set Mapping}
+In order to alter the tag set of NDT, we created a mapping for carrying out the
+tag set modifications. We created a mapping for carrying out the tag set
+modifications that maps the relevant existing tags to new, more fine-grained
+tags including more relevant morphological features for the applicable tokens.
+
+\paragraph{Baseline}
+It is common practice to compare the performance of PoS taggers to a
+pre-computed baseline for an initial point of comparison.
+For PoS tagging, a commonly used baseline is the Most Frequent Tag (MFT)
+baseline, which we use in our experiments. This involves labeling each word
+with the tag it was assigned most frequently in the training. All unknown
+words, i.e., words not seen in the training data, are assigned the tag most
+frequently assigned to words seen only once in the training. Unknown and
+infrequent words have in common that they rarely occur, and we might therefore
+expect them to have similar properties.
+%, there are two main
+%approaches: assign it the most frequent tag overall in the training data, or
+%the tag most frequently assigned to words seen only once in the training data.
+%We adopted the latter approach for our baseline tagger, as we believe that
+%unknown words may have much in common with words that occur only once in the
+%training, more so than simply words of the most frequent part-of-speech. The
+%reason for this is that unknown and infrequent words have in common that they
+%rarely occur, and we might therefore expect them to have similar properties.
+
+\paragraph{Tags \& Features}
+As we seek to quantify the effects of PoS tagging in a realistic setting, we
+want to run the parser on automatically assigned PoS tags. For the training of
+the parser, however, we have two options: using either gold standard or
+automatically assigned tags. In order to settle on a configuration, we
+conducted experiments with gold standard and automatically assigned tags to see
+how they differ with respect to performance. The results of our experiments
+reveal that the combination of training and testing on automatic tags is
+superior to training on gold standard tags and testing on automatic tags,
+surprisingly. Consequently, the parser was both trained and tested on
+automatically assigned tags in our experiments.
+
+Note that it is absolutely crucial that the morphological features in the
+treebank are removed when using automatic tags, as they are still gold
+standard. For instance, if a verb token is erroneously tagged as a noun, we
+could potentially have a noun token with verbal features such as tense, which
+markedly obfuscates the training and parsing. Another important factor is that
+we want to isolate the effect of PoS tags, necessitating the exclusion of
+morphological features.
+
+\section{Tag Set Optimization}
+\label{sec:optimization}
+
+%\begin{table}
+    %\centering
+    %%\smaller[0.5]
+    %\begin{tabular}{@{}ll@{}}
+        %\toprule
+        %\textbf{Tag} & \textbf{Description} \\
+        %\midrule
+        %\texttt{adj|komp} & Comparative adjective \\
+        %\texttt{adj|pos} & Positive adjective \\
+        %\texttt{adj|sup} & Superlative adjective \\
+        %\texttt{det|be} & Definite determiner \\
+        %\texttt{det|ub} & Indefinite determiner \\
+        %\texttt{pron|pers} & Personal pronoun \\
+        %\texttt{pron|pers|akk} & Personal pronoun, accusative \\
+        %\texttt{pron|pers|nom} & Personal pronoun, nominative \\
+        %\texttt{pron|refl} & Reflexive pronoun \\
+        %\texttt{pron|res} & Reciprocal pronoun \\
+        %\texttt{pron|sp} & Interrogative pronoun \\
+        %\texttt{subst|appell} & Common noun \\
+        %\texttt{subst|appell|be} & Common noun, definite \\
+        %\texttt{subst|appell|be|gen} & Common noun, definite, genitive \\
+        %\texttt{subst|appell|ub} & Common noun, indefinite \\
+        %\texttt{subst|appell|ub|gen} & Common noun, indefinite, genitive \\
+        %\texttt{subst|prop} & Proper noun \\
+        %\texttt{subst|prop|gen} & Proper noun, genitive \\
+        %\texttt{verb|fin} & Finite verb \\
+        %\texttt{verb|infin} & Nonfinite verb \\
+        %\bottomrule
+    %\end{tabular}
+    %\caption{The optimized tag set.}
+    %\label{optimizedtagset}
+%\end{table}
+
+\section{Optimized Pipeline}
+\label{sec:pipeline}
+
+\section{Summary/Conclusion and Future Work}
+\label{sec:summary}
 
@@
 
-\bibliography{acl2016}
 \bibliographystyle{acl2016}
-
+\bibliography{references}
 % \appendix

### 2016-05-02 submission-petter.tex
Minor additions to optimization section

@@
 \label{sec:prev}
-
-Megyesi, Swedish
-
-MacKinlay, English
-
-OBT, Norwegian
+Previous work related to our experiments with tag set modifications include
+investigating the effects of PoS tag sets on tagging of Swedish
+\cite{Meg:01,Meg:02,Meg:09} and English \cite{Mac:05}. The important difference
+is that they investigate the effects of PoS tags on tagging, while we assess
+its effects on a downstream application, namely syntactic parsing.
 
@@
 \subsection{The Norwegian Dependency Treebank}
-The newly developed Norwegian Dependency Treebank (NDT) \cite{Sol:Skj:Ovr:14}
-is the first publicly available treebank for Norwegian. The treebank was used
-for training and evaluation in our experiments. The tag set consists of 12
-morphosyntactic PoS tags, with 7 additional tags for punctuation and symbols.
+We used the newly developed Norwegian Dependency Treebank (NDT)
+\cite{Sol:Skj:Ovr:14}, which is the first publicly available treebank for
+Norwegian (for our experiments). It was developed at the National Library of
+Norway in collaboration with the University of Oslo, and contains manual
+syntactic and morphological annotation. The treebank contains 311 000 tokens
+of Bokmål and 303 000 tokens of Nynorsk, the morphological annotation of which
+follows that of the Oslo-Bergen Tagger \cite{Hag:Joh:Nok:00,Sol:13}, which in
+turn is largely based on the work of \newcite{Faa:Lie:Van:97}. The annotated
+texts are mostly newspaper text, but also include government reports,
+parliament transcripts and excerpts from blogs. The annotation process of the
+treebank was supported by the Oslo-Bergen Tagger and then manually corrected by
+annotators.
+
+Noe om OBT? hmmm
 
 \paragraph{PoS Tag Set}
+The tag set consists of 12 morphosyntactic PoS tags, with 7 additional tags for
+punctuation and symbols. The tag set is thus rather coarse-grained, with broad
+categories such as \texttt{subst} (noun) and \texttt{verb} (verb).
+
 \paragraph{Dependency Relations}
@@
 \label{sec:setup}
+In preparation to conducting our experiments with linguistically motivated tag
+set modifications, a concrete setup for the experiments needed to be
+established, which is presented in the following.
+
+\paragraph{Data Set Split}
+As there was no standardized data set split of NDT due to its very recent
+development, we needed to establish a data set split
+(training/development/test) in preparation to our experiments. Our data set
+split of the treebank follows the standard 80-10-10 (training/development/test)
+split and will be distributed with the treebank and proposed as the new
+standard(?). In creating the data set split, care has been taken to preserve
+contiguous texts in the various data sets while keeping the split balanced in
+terms of genre (and source). Our proposed data set split was used in the
+Norwegian contribution to the Universal Dependencies project \cite{Ovr:Hoh:16}.
+The split will be made available at a companion website.
 
 \paragraph{Tagger}
-TnT \cite{Bra:00}
+For our experiments with tag set modifications, we wanted a PoS tagger that is
+both fast and accurate. There is often a trade-off between the two, as the best
+taggers tend to suffer in terms of speed due to their complexity. However, a
+tagger that achieves both close to state-of-the-art accuracy as well as very
+high speed is TnT \cite{Bra:00}. The fact that TnT was used for evaluating the
+universal tag set \cite{Pet:Das:McD:12}, served as another good indication of
+TnT being appropriate for our task. The sum of these factors led to TnT being
+the tagger of choice for our experiments.
 
@@
 They found the Mate parser \cite{Boh:10} to be the most successful parser for
-the parsing of NDT. Recent dependency parser comparisons \cite{Cho:Tet:Ste:15}
-also showed that Mate performed very well on parsing of the English portion of
-the OntoNotes 5 corpus, beating a range of contemporary state-of-the-art
-parsers.
+the parsing of NDT. Furthermore, recent dependency parser comparisons
+\cite{Cho:Tet:Ste:15} showed that Mate performed very well on parsing of the
+English portion of the OntoNotes 5 corpus, beating a range of contemporary
+state-of-the-art parsers.
 
@@
 
+For each tag, we first experiment with each of the features in isolation before
+employing various combinations of them. We base our choices of combinations on
+how promising the features are and what we deem worth investigating in terms of
+linguistic utility, in order to see how the features might interact.
+
 %\begin{table}

### 2016-05-03 submission-petter.tex
Additions to data, more tables

@@
 \usepackage{booktabs}
+\usepackage{relsize}
 
@@
 \label{sec:prev}
-Previous work related to our experiments with tag set modifications include
-investigating the effects of PoS tag sets on tagging of Swedish
-\cite{Meg:01,Meg:02,Meg:09} and English \cite{Mac:05}. The important difference
-is that they investigate the effects of PoS tags on tagging, while we assess
-its effects on a downstream application, namely syntactic parsing.
+The lack of previous efforts on extrinsic evaluation of the effects of PoS
+tag sets on downstream applications is considerable. We will, instead/however,
+consider(/noe annet, considerable...consider)  papers that evaluate the effects
+of PoS tag set granularity on PoS tagging.  This includes investigating the
+effects of PoS tag sets on tagging of Swedish \cite{Meg:01,Meg:02,Meg:09} and
+English \cite{Mac:05}. Increases in tagger accuracy does not readily translate
+to improved parsing.
 
@@
 \cite{Sol:Skj:Ovr:14}, which is the first publicly available treebank for
-Norwegian (for our experiments). It was developed at the National Library of
-Norway in collaboration with the University of Oslo, and contains manual
-syntactic and morphological annotation. The treebank contains 311 000 tokens
-of Bokmål and 303 000 tokens of Nynorsk, the morphological annotation of which
-follows that of the Oslo-Bergen Tagger \cite{Hag:Joh:Nok:00,Sol:13}, which in
-turn is largely based on the work of \newcite{Faa:Lie:Van:97}. The annotated
-texts are mostly newspaper text, but also include government reports,
-parliament transcripts and excerpts from blogs. The annotation process of the
-treebank was supported by the Oslo-Bergen Tagger and then manually corrected by
-annotators.
+Norwegian. It was developed at the National Library of Norway in collaboration
+with the University of Oslo, and contains manual syntactic and morphological
+annotation. The treebank contains 311 000 tokens of Bokmål and 303 000 tokens
+of Nynorsk.  The annotated texts are mostly newspaper text, but also include
+government reports, parliament transcripts and excerpts from blogs. The
+annotation process of the treebank was supported by the Oslo-Bergen Tagger and
+then manually corrected by annotators.
 
@@
 
-\paragraph{PoS Tag Set}
-The tag set consists of 12 morphosyntactic PoS tags, with 7 additional tags for
-punctuation and symbols. The tag set is thus rather coarse-grained, with broad
-categories such as \texttt{subst} (noun) and \texttt{verb} (verb).
+Linguistic motivation, insight, utility...
 
-\paragraph{Dependency Relations}
+\paragraph{Morphological Annotation}
+The morphological annotation follows that of the Oslo-Bergen Tagger
+\cite{Hag:Joh:Nok:00,Sol:13}, which in turn is largely based on the work of
+\newcite{Faa:Lie:Van:97}. The tag set of NDT is based on the tag set of the
+Oslo-Bergen Tagger. The PoS tags are complemented by a set of available
+morphological features, such as definiteness, number and tense. These features
+are used in our tag set modifications, where the coarse PoS tag of relevant
+tokens are concatenated with one or more of these features.  This allows us to
+include more linguistic information in the tags. The tag set consists of 12
+morphosyntactic PoS tags, with 7 additional tags for punctuation and symbols.
+The tag set is thus rather coarse-grained, with broad categories such as
+\texttt{subst} (noun) and \texttt{verb} (verb).
+
+\paragraph{Syntactic Annotation} The syntactic annotation choices in NDT are
+largely based on the Norwegian Reference Grammar \cite{Faa:Lie:Van:97}. The
+annotation choices are outlined in Table \ref{ndtannotation}, taken from
+\newcite{Sol:Skj:Ovr:14}, providing overview of the analyses of syntactic
+constructions that often distinguish dependency treebanks, such as coordination
+and the treatment of auxiliary and main verbs.
+
+\begin{table} \centering
+    %\smaller[0.5]
+    \begin{tabular}{@{}ll@{}}
+        \toprule
+        \textbf{Head} & \textbf{Dependent} \\
+        \midrule
+        Preposition & Prepositional complement \\
+        Finite verb & Complementizer \\
+        First conjunct & Subsequent conjuncts \\
+        Finite auxiliary & Lexical/main verb \\
+        Noun & Determiner \\
+        \bottomrule
+    \end{tabular}
+    \caption{Annotation choices in NDT.}
+    \label{ndtannotation}
+\end{table}
 
@@
     %\centering
-    %%\smaller[0.5]
-    %\begin{tabular}{@{}ll@{}}
+    %\smaller[0.5]
+    %\begin{tabular}{@{}lllll@{}}
         %\toprule
-        %\textbf{Tag} & \textbf{Description} \\
+        %\textbf{Tag Set} & \textbf{MFT} & \textbf{Accuracy} &
+        %\textbf{LAS} & \textbf{UAS} \\
         %\midrule
-        %\texttt{adj|komp} & Comparative adjective \\
-        %\texttt{adj|pos} & Positive adjective \\
-        %\texttt{adj|sup} & Superlative adjective \\
-        %\texttt{det|be} & Definite determiner \\
-        %\texttt{det|ub} & Indefinite determiner \\
-        %\texttt{pron|pers} & Personal pronoun \\
-        %\texttt{pron|pers|akk} & Personal pronoun, accusative \\
-        %\texttt{pron|pers|nom} & Personal pronoun, nominative \\
-        %\texttt{pron|refl} & Reflexive pronoun \\
-        %\texttt{pron|res} & Reciprocal pronoun \\
-        %\texttt{pron|sp} & Interrogative pronoun \\
-        %\texttt{subst|appell} & Common noun \\
-        %\texttt{subst|appell|be} & Common noun, definite \\
-        %\texttt{subst|appell|be|gen} & Common noun, definite, genitive \\
-        %\texttt{subst|appell|ub} & Common noun, indefinite \\
-        %\texttt{subst|appell|ub|gen} & Common noun, indefinite, genitive \\
-        %\texttt{subst|prop} & Proper noun \\
-        %\texttt{subst|prop|gen} & Proper noun, genitive \\
-        %\texttt{verb|fin} & Finite verb \\
-        %\texttt{verb|infin} & Nonfinite verb \\
+        %Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
+        %Full & 85.15\% & 93.48\% & \textbf{87.15\%} & \textbf{90.39\%} \\
         %\bottomrule
     %\end{tabular}
-    %\caption{The optimized tag set.}
-    %\label{optimizedtagset}
+    %\caption{Evaluation of tagging and parsing the development data with the two
+        %initial tag sets.}
+    %\label{inittagseteval}
 %\end{table}
+
+\begin{table*}
+    \centering
+    \smaller[0.5]
+    \begin{tabular}{@{}llllll@{}}
+        \toprule
+        \textbf{Category} & \textbf{Feature(s)} & \textbf{MFT} &
+        \textbf{Accuracy} & \textbf{LAS} & \textbf{UAS} \\
+        \midrule
+        \emph{Baseline} & --- & 94.14\% & 97.47\% & 87.01\% & 90.19\% \\
+        Noun & Type, case \& definiteness & 89.61\% & 97.05\% & 88.81\% &
+        91.73\% \\
+        Verb & Finiteness & 93.72\% & 97.35\% & 87.30\% & 90.43\% \\
+        Adjective & Degree & 94.13\% & 97.41\% & 87.29\% & 90.44\% \\
+        Determiner & Definiteness & 94.13\% & 97.49\% & 87.30\% & 90.42\% \\
+        Pronoun & Type \& case & 94.12\% & 97.51\% & 87.30\% & 90.41\% \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of tagging and parsing with the most successful tag set
+        modification for each category.}
+    \label{respectiveresults}
+\end{table*}
+
+\begin{table}
+    \centering
+    \smaller[2]
+    \begin{tabular}{@{}ll@{}}
+        \toprule
+        \textbf{Tag} & \textbf{Description} \\
+        \midrule
+        \texttt{adj|komp} & Comparative adjective \\
+        \texttt{adj|pos} & Positive adjective \\
+        \texttt{adj|sup} & Superlative adjective \\
+        \texttt{det|be} & Definite determiner \\
+        \texttt{det|ub} & Indefinite determiner \\
+        \texttt{pron|pers} & Personal pronoun \\
+        \texttt{pron|pers|akk} & Personal pronoun, accusative \\
+        \texttt{pron|pers|nom} & Personal pronoun, nominative \\
+        \texttt{pron|refl} & Reflexive pronoun \\
+        \texttt{pron|res} & Reciprocal pronoun \\
+        \texttt{pron|sp} & Interrogative pronoun \\
+        \texttt{subst|appell} & Common noun \\
+        \texttt{subst|appell|be} & Common noun, definite \\
+        \texttt{subst|appell|be|gen} & Common noun, definite, genitive \\
+        \texttt{subst|appell|ub} & Common noun, indefinite \\
+        \texttt{subst|appell|ub|gen} & Common noun, indefinite, genitive \\
+        \texttt{subst|prop} & Proper noun \\
+        \texttt{subst|prop|gen} & Proper noun, genitive \\
+        \texttt{verb|fin} & Finite verb \\
+        \texttt{verb|infin} & Nonfinite verb \\
+        \bottomrule
+    \end{tabular}
+    \caption{The optimized tag set.}
+    \label{optimizedtagset}
+\end{table}
+
+\begin{table}
+    \centering
+    \smaller[0.5]
+    \begin{tabular}{@{}lllll@{}}
+        \toprule
+        \textbf{Tag set} & \textbf{MFT} & \textbf{Accuracy} &
+        \textbf{LAS} & \textbf{UAS} \\
+        \midrule
+        Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
+        Full & 85.12\% & 93.46\% & 87.13\% & 90.32\% \\
+        Optimized & 89.20\% & 96.85\% & \textbf{88.87\%} & \textbf{91.78\%} \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of tagging and parsing with the optimized tag set, compared to
+        the initial tag sets.}
+    \label{finalresults}
+\end{table}
 

### 2016-05-03 submission.tex
minimal skeleton
Additions to experimental setup and data
File cleanup, submission is main file

@@
 
-\documentclass[11pt]{article}
+\documentclass[11pt,a4paper]{article}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
 \usepackage{acl2016}
@@
 \usepackage{latexsym}
+\usepackage{booktabs}
+\usepackage{relsize}
 
@@
 
-\title{The greatest PoS tags, just the best really}
+\title{Optimizing a PoS Tag Set for Dependency Parsing}
 
@@
 \section{Introduction}
+\label{sec:intro}
+
+\section{Previous/Related work}
+\label{sec:prev}
+There is a considerable lack of previous efforts on extrinsic evaluation of the
+effects of PoS tag sets on downstream applications. We will consider(/noe
+annet, considerable...consider) papers that evaluate the effects of PoS tag
+set granularity on PoS tagging. This includes investigation of the effects of
+PoS tag sets on tagging of Swedish \cite{Meg:01,Meg:02} and English
+\cite{Mac:05}. Increases in tagger accuracy does not readily translate to
+improved parsing.
+
+\cite{Meg:01,Meg:02} trained and evaluated a range of PoS taggers on the
+Stockholm-Umeå Corpus (SUC) \cite{Gus:Har:06}, annotated with a Swedish version
+of PAROLE tags, with the tag set totaling 139 tags. Furthermore, they
+investigated the effects of tag set size on tagging by mapping the original tag
+set into smaller subsets designed for parsing. They argue that a tag set with
+complete morphological tags may not be necessary for all NLP applications, for
+instance syntactic parsing. They found that the smallest tag set comprising 26
+tags yields the lowest tagger error rate. However, for some of the taggers,
+augmenting the tag set with more linguistically informative tags may actually
+lead to a drop in error rate. They argue that this shows that the size of the
+tag set as well as the type of information in the tags are crucial factors for
+tagger performance. However/unfortunately, they do not report results of parsing
+with the various PoS tag sets.
+
+Similarly, \newcite{Mac:05} investigated the effects of PoS tag sets on PoS
+tagger performance in English, specifically the Penn Treebank
+\cite{Mar:San:Mar:93}. They mapped the original tag set of Penn Treebank to
+more fine-grained tag sets "in terms of" linguistic utility. the additional
+linguistic information included in the tags may assist the tagger.
+utility/motivation to investigate whether linguistically motivated tag set
+modifications can improve the tagger accuracy while increasing the linguistic
+information in the tag set. Experimenting with both morphologically and
+syntactically motivated modifications, they found that more fine-grained
+tag sets rarely led to improvements in tagger accuracy,
+where the most successful modification yielded an improvement in tagger
+accuracy of 0.05 percentage points.
+
+\section{Data}
+\label{sec:data}
+\subsection{The Norwegian Dependency Treebank}
+We used the newly developed Norwegian Dependency Treebank (NDT)
+\cite{Sol:Skj:Ovr:14}, which is the first publicly available treebank for
+Norwegian. It was developed at the National Library of Norway in collaboration
+with the University of Oslo, and contains manual syntactic and morphological
+annotation. The treebank contains 311 000 tokens of Bokmål and 303 000 tokens
+of Nynorsk.  The annotated texts are mostly newspaper text, but also include
+government reports, parliament transcripts and excerpts from blogs. The
+annotation process of the treebank was supported by the Oslo-Bergen Tagger
+\cite{Hag:Joh:Nok:00} and then manually corrected by annotators.
+
+%Linguistic motivation, insight, utility...
+
+\paragraph{Morphological Annotation}
+The morphological annotation and PoS tag set of NDT follows the Oslo-Bergen
+Tagger \cite{Hag:Joh:Nok:00,Sol:13}, which in turn is largely based on the work
+of \newcite{Faa:Lie:Van:97}. The tag set consists of 12 morphosyntactic PoS
+tags, with 7 additional tags for punctuation and symbols.  The tag set is thus
+rather coarse-grained, with broad categories such as \texttt{subst} (noun) and
+\texttt{verb} (verb). The PoS tags are complemented by a large set of
+morphological features, providing information about morphological properties such as
+definiteness, number and tense. These features are used in our tag set
+modifications, where the coarse PoS tag of relevant tokens are concatenated
+with one or more of these features.  This allows us to include more linguistic
+information in the tags.
+
+\begin{table}
+    \centering
+    \smaller[0.5]
+    \begin{tabular}{@{}ll@{}}
+        \toprule
+        \textbf{Tag} & \textbf{Description} \\
+        \midrule
+        \texttt{adj} & Adjective \\
+        \texttt{adv} & Adverb \\
+        \texttt{det} & Determiner\\
+        \texttt{inf-merke} & Infinitive marker \\
+        \texttt{interj} & Interjection \\
+        \texttt{konj} & Conjunction \\
+        \texttt{prep} & Preposition \\
+        \texttt{pron} & Pronoun \\
+        \texttt{sbu} & Subordinate conjunction \\
+        \texttt{subst} & Noun \\
+        \texttt{ukjent} & Unknown (foreign word) \\
+        \texttt{verb} & Verb \\
+        \bottomrule
+    \end{tabular}
+    \caption{Overview of the tag set of NDT.}
+    \label{ndttagest}
+\end{table}
+
+\paragraph{Syntactic Annotation}
+The syntactic annotation choices in NDT are largely based on the Norwegian
+Reference Grammar \cite{Faa:Lie:Van:97}. The annotation choices are outlined in
+Table \ref{ndtannotation}, taken from \newcite{Sol:Skj:Ovr:14}, providing
+overview of the analyses of syntactic constructions that often distinguish
+dependency treebanks, such as coordination and the treatment of auxiliary and
+main verbs. The set of dependency relations contains 29 dependency relations,
+including \textsc{adv} (adverbial), \textsc{subj} (subject) and \textsc{koord}
+(coordination).
+
+\begin{table} \centering
+    %\smaller[0.5]
+    \begin{tabular}{@{}ll@{}}
+        \toprule
+        \textbf{Head} & \textbf{Dependent} \\
+        \midrule
+        Preposition & Prepositional complement \\
+        Finite verb & Complementizer \\
+        First conjunct & Subsequent conjuncts \\
+        Finite auxiliary & Lexical/main verb \\
+        Noun & Determiner \\
+        \bottomrule
+    \end{tabular}
+    \caption{Annotation choices in NDT.}
+    \label{ndtannotation}
+\end{table}
+
+\section{Experimental Setup}
+\label{sec:setup}
+In preparation to conducting our experiments with linguistically motivated tag
+set modifications, a concrete setup for the experiments needed to be
+established, which is presented in the following.
+
+\paragraph{Data Set Split}
+As there was no standardized data set split of NDT due to its very recent
+development, we needed to establish a data set split
+(training/development/test) in preparation to our experiments. Our data set
+split of the treebank follows the standard 80-10-10 (training/development/test)
+split and will be distributed with the treebank and proposed as the new
+standard(?). In creating the data set split, care has been taken to preserve
+contiguous texts in the various data sets while keeping the split balanced in
+terms of genre (and source). Our proposed data set split was used in the
+Norwegian contribution to the Universal Dependencies project \cite{Ovr:Hoh:16}.
+The split will be made available at a companion website.
+
+\paragraph{Tagger}
+For our experiments with tag set modifications, we wanted a PoS tagger that is
+both fast and accurate. There is often a trade-off between the two, as the best
+taggers tend to suffer in terms of speed due to their complexity. However, a
+tagger that achieves both close to state-of-the-art accuracy as well as very
+high speed is TnT \cite{Bra:00}. The fact that TnT was used for evaluating the
+universal tag set \cite{Pet:Das:McD:12}, served as another good indication of
+TnT being appropriate for our task. The sum of these factors led to TnT being
+the tagger of choice for our experiments.
+
+\paragraph{Parser}
+In choosing a syntactic parser for our experiments, we considered previous work
+on dependency parsing of Norwegian, specifically that of \cite{Sol:Skj:Ovr:14}.
+They found the Mate parser \cite{Boh:10} to be the most successful parser for
+the parsing of NDT. Furthermore, recent dependency parser comparisons
+\cite{Cho:Tet:Ste:15} showed that Mate performed very well on parsing of the
+English portion of the OntoNotes 5 corpus, beating a range of contemporary
+state-of-the-art parsers.
+
+\paragraph{Tag Set Mapping}
+In order to alter the tag set of NDT, we created a mapping for carrying out the
+tag set modifications. We created a mapping for carrying out the tag set
+modifications that maps the relevant existing tags to new, more fine-grained
+tags including more relevant morphological features for the applicable tokens.
+
+\paragraph{Baseline}
+It is common practice to compare the performance of PoS taggers to a
+pre-computed baseline for an initial point of comparison.
+For PoS tagging, a commonly used baseline is the Most Frequent Tag (MFT)
+baseline, which we use in our experiments. This involves labeling each word
+with the tag it was assigned most frequently in the training. All unknown
+words, i.e., words not seen in the training data, are assigned the tag most
+frequently assigned to words seen only once in the training. Unknown and
+infrequent words have in common that they rarely occur, and we might therefore
+expect them to have similar properties.
+%, there are two main
+%approaches: assign it the most frequent tag overall in the training data, or
+%the tag most frequently assigned to words seen only once in the training data.
+%We adopted the latter approach for our baseline tagger, as we believe that
+%unknown words may have much in common with words that occur only once in the
+%training, more so than simply words of the most frequent part-of-speech. The
+%reason for this is that unknown and infrequent words have in common that they
+%rarely occur, and we might therefore expect them to have similar properties.
+
+\paragraph{Tags \& Features}
+As we seek to quantify the effects of PoS tagging in a realistic setting, we
+want to run the parser on automatically assigned PoS tags. For the training of
+the parser, however, we have two options: using either gold standard or
+automatically assigned tags. In order to settle on a configuration, we
+conducted experiments with gold standard and automatically assigned tags to see
+how they differ with respect to performance. The results of our experiments
+reveal that the combination of training and testing on automatic tags is
+superior to training on gold standard tags and testing on automatic tags,
+surprisingly. Consequently, the parser was both trained and tested on
+automatically assigned tags in our experiments.
+
+Note that it is absolutely crucial that the morphological features in the
+treebank are removed when using automatic tags, as they are still gold
+standard. For instance, if a verb token is erroneously tagged as a noun, we
+could potentially have a noun token with verbal features such as tense, which
+markedly obfuscates the training and parsing. Another important factor is that
+we want to isolate the effect of PoS tags, necessitating the exclusion of
+morphological features.
+
+\section{Tag Set Optimization/Experiments}
+\label{sec:optimization}
+With more fine-grained linguistically
+motivated distinctions, we increase the linguistic information represented in
+the tags, which may assist the tagger in disambiguating ambiguous and unknown
+words, which in turn may aid the parser in recognizing and generalizing
+syntactic patterns. However, the addition of more linguistic information to the
+tags and thus a more fine-grained tag set will most likely lead to a drop in
+tagger accuracy, due to the increase in complexity. The best tagging does not
+necessarily lead to the best parse, and it is therefore interesting to
+investigate how the tag set modifications may affect the interplay between
+tagging and parsing.
+
+\subsection{Tag Set Experiments}
+We modified the tags for nouns, verbs, adjectives, determiners and pronouns in
+NDT by appending selected sets of morphological features to each tag in order
+to increase the linguistic information expressed by the tags.  For each tag, we
+first experimented with each of the features in isolation before employing
+various combinations of them. We based our choices of combinations on how
+promising the features are and what we deem worth investigating in terms of
+linguistic utility, in order to see how the features might interact.
+
+\paragraph{Nouns}
+In Norwegian, nouns are assigned gender (feminine, masculine, neuter),
+definiteness (indefinite or definite) and number (singular or plural). There
+is agreement in gender, definiteness and number between nouns and their
+modifiers (adjectives and determiners).
+
+\paragraph{Verbs}
+Verbs are inflected for tense (infinitive, present, preterite, past perfect)
+in Norwegian and can additionally take on mood (imperative, indicative) and
+voice (active or passive).
+
+\paragraph{Adjectives}
+Adjectives agree with the noun they modify in terms of gender, number and
+definiteness in Norwegian.
+
+\paragraph{Determiners}
+Like adjectives, determiners in Norwegian agree with the noun they modify in
+terms of gender, number and definiteness.
+
+\paragraph{Pronouns}
+Pronouns in Norwegian include personal, reciprocal, reflexive and
+interrogative. They can exhibit gender, number and person. Personal pronouns
+have case (accusative or nominative).
+
+%\begin{table}
+    %\centering
+    %\smaller[0.5]
+    %\begin{tabular}{@{}lllll@{}}
+        %\toprule
+        %\textbf{Tag Set} & \textbf{MFT} & \textbf{Accuracy} &
+        %\textbf{LAS} & \textbf{UAS} \\
+        %\midrule
+        %Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
+        %Full & 85.15\% & 93.48\% & \textbf{87.15\%} & \textbf{90.39\%} \\
+        %\bottomrule
+    %\end{tabular}
+    %\caption{Evaluation of tagging and parsing the development data with the two
+        %initial tag sets.}
+    %\label{inittagseteval}
+%\end{table}
+
+\begin{table*}
+    \centering
+    \smaller[0.5]
+    \begin{tabular}{@{}llllll@{}}
+        \toprule
+        \textbf{Category} & \textbf{Feature(s)} & \textbf{MFT} &
+        \textbf{Accuracy} & \textbf{LAS} & \textbf{UAS} \\
+        \midrule
+        \emph{Baseline} & --- & 94.14\% & 97.47\% & 87.01\% & 90.19\% \\
+        Noun & Type, case \& definiteness & 89.61\% & 97.05\% & 88.81\% &
+        91.73\% \\
+        Verb & Finiteness & 93.72\% & 97.35\% & 87.30\% & 90.43\% \\
+        Adjective & Degree & 94.13\% & 97.41\% & 87.29\% & 90.44\% \\
+        Determiner & Definiteness & 94.13\% & 97.49\% & 87.30\% & 90.42\% \\
+        Pronoun & Type \& case & 94.12\% & 97.51\% & 87.30\% & 90.41\% \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of tagging and parsing with the most successful tag set
+        modification for each category.}
+    \label{respectiveresults}
+\end{table*}
+
+\begin{table}
+    \centering
+    \smaller[1]
+    \begin{tabular}{@{}ll@{}}
+        \toprule
+        \textbf{Tag} & \textbf{Description} \\
+        \midrule
+        \texttt{adj|komp} & Comparative adjective \\
+        \texttt{adj|pos} & Positive adjective \\
+        \texttt{adj|sup} & Superlative adjective \\
+        \texttt{det|be} & Definite determiner \\
+        \texttt{det|ub} & Indefinite determiner \\
+        \texttt{pron|pers} & Personal pronoun \\
+        \texttt{pron|pers|akk} & Personal pronoun, accusative \\
+        \texttt{pron|pers|nom} & Personal pronoun, nominative \\
+        \texttt{pron|refl} & Reflexive pronoun \\
+        \texttt{pron|res} & Reciprocal pronoun \\
+        \texttt{pron|sp} & Interrogative pronoun \\
+        \texttt{subst|appell} & Common noun \\
+        \texttt{subst|appell|be} & Common noun, definite \\
+        \texttt{subst|appell|be|gen} & Common noun, definite, genitive \\
+        \texttt{subst|appell|ub} & Common noun, indefinite \\
+        \texttt{subst|appell|ub|gen} & Common noun, indefinite, genitive \\
+        \texttt{subst|prop} & Proper noun \\
+        \texttt{subst|prop|gen} & Proper noun, genitive \\
+        \texttt{verb|fin} & Finite verb \\
+        \texttt{verb|infin} & Nonfinite verb \\
+        \bottomrule
+    \end{tabular}
+    \caption{Overview of the optimized tag set.}
+    \label{optimizedtagset}
+\end{table}
+
+\begin{table}
+    \centering
+    \smaller[0.5]
+    \begin{tabular}{@{}lllll@{}}
+        \toprule
+        \textbf{Tag set} & \textbf{MFT} & \textbf{Accuracy} &
+        \textbf{LAS} & \textbf{UAS} \\
+        \midrule
+        Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
+        Full & 85.12\% & 93.46\% & 87.13\% & 90.32\% \\
+        Optimized & 89.20\% & 96.85\% & \textbf{88.87\%} & \textbf{91.78\%} \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of tagging and parsing with the optimized tag set, compared to
+        the initial tag sets.}
+    \label{finalresults}
+\end{table}
+
+\section{Optimized Pipeline}
+\label{sec:pipeline}
+
+\section{Summary/Conclusion and Future Work}
+\label{sec:summary}
 

### 2016-05-03 submission.tex
Added preliminary intro

@@
 \label{sec:intro}
+PoS tagging is an important preprocessing step for many NLP
+tasks, such as parsing, named entity recognition and sentiment
+analysis. Whereas much effort has gone into the development of
+PoS-taggers, to the effect that this task is often considered
+more or less a solved task, considerably less effort has been
+devoted to the empirical evaluation of the PoS tag sets
+themselves. 
+PoS tag sets are usually employed in a "one size fits
+all" fashion, regardless of the requirements posed by the
+downstream NLP tasks for which PoS-tagging is performed.
+
+It is well known that syntactic parsing often benefits from quite
+fine-grained morphological distinctions. However, in a realistic
+setting where the aim is to parse raw text, this will often
+require a separate step of morphological analysis in addition to
+PoS-tagging that can be quite expensive.
+
+In this paper, we optimize a Norwegian PoS tag set for the task of
+data-driven dependency parsing.  We show that the introduction of
+morphological distinctions not present in the original tag set, whilst
+compromising tagger accuracy, actually leads to significantly improved
+parsing accuracy.
+
+The article is structured as follows. We start out by reviewing
+previous work on tag set evaluation in Section \ref{sec:prev} and
 

### 2016-05-03 submission.tex
Additions to previous work by MacKinlay and Megyesi

--- 2016-05-03 submission.tex
@@
 PoS tag sets on tagging of Swedish \cite{Meg:01,Meg:02} and English
-\cite{Mac:05}. Increases in tagger accuracy does not readily translate to
-improved parsing.
+\cite{Mac:05}.
+
+%Increases in tagger accuracy does not readily translate to
+%improved parsing.
 
@@
 
-Similarly, \newcite{Mac:05} investigated the effects of PoS tag sets on PoS
-tagger performance in English, specifically the Penn Treebank
-\cite{Mar:San:Mar:93}. They mapped the original tag set of Penn Treebank to
-more fine-grained tag sets "in terms of" linguistic utility. the additional
-linguistic information included in the tags may assist the tagger.
-utility/motivation to investigate whether linguistically motivated tag set
-modifications can improve the tagger accuracy while increasing the linguistic
-information in the tag set. Experimenting with both morphologically and
-syntactically motivated modifications, they found that more fine-grained
-tag sets rarely led to improvements in tagger accuracy,
-where the most successful modification yielded an improvement in tagger
+Similarly, \newcite{Mac:05} investigated the effects of PoS tag sets on tagger
+performance in English, specifically the Wall Street Journal portion of Penn
+Treebank \cite{Mar:San:Mar:93}. They mapped the original tag set of Penn
+Treebank to more fine-grained tag sets using linguistic insight to investigate
+whether the additional linguistic information included in the finer-grained
+tags could assist the tagger. Experimenting with both lexically and
+syntactically conditioned modifications such as distinguishing between count
+nouns and noncount nouns as well as between transitive and intransitive verbs,
+they found that more fine-grained tag sets rarely led to improvements in tagger
+accuracy; the most successful modification yielded an improvement in tagger
 accuracy of 0.05 percentage points.
+
+Transition/overgang her ...
 
@@
     \centering
-    \smaller[0.5]
+    %\smaller[0.5]
     \begin{tabular}{@{}ll@{}}
@@
     \end{tabular}
-    \caption{Overview of the tag set of NDT.}
+    \caption{Overview of the PoS tag set of NDT.}
     \label{ndttagest}

### 2016-05-03 submission.tex
Added overview of article

--- 2016-05-03 submission.tex
@@
 \label{sec:intro}
+PoS tagging is an important preprocessing step for many NLP
+tasks, such as parsing, named entity recognition and sentiment
+analysis. Whereas much effort has gone into the development of
+PoS-taggers, to the effect that this task is often considered
+more or less a solved task, considerably less effort has been
+devoted to empirical evaluation of the PoS tag sets
+themselves. 
+PoS tag sets are usually employed in a ``one size fits
+all'' fashion, regardless of the requirements posed by the
+downstream NLP tasks for which PoS-tagging is performed.
+
+It is well known that syntactic parsing often benefits from quite
+fine-grained morphological distinctions since morphology interacts
+with syntax though phenomena such as agreement and case
+marking. However, in a realistic setting where the aim is to
+automatically parse raw text, the generation of morphological
+information will often require a separate step of morphological
+analysis that can be quite costly.
+
+In this paper, we optimize a Norwegian PoS tag set for the task of
+data-driven dependency parsing. In a set of experiments, various
+morphological distinctions are introduced and evaluated both
+intrinsically, i.e. in terms of PoS-tagging accuracy and
+extrinsically, in terms of parsing accuracy.  Our results show that
+the introduction of morphological distinctions not present in the
+original tag set, whilst compromising tagger accuracy, actually leads
+to significantly improved parsing accuracy. This optimization allows
+us to bypass the additional step of morphological analysis, framing
+the whole preprocessing problem as a simple tagging task.
+
+The article is structured as follows. We start out by reviewing
+previous work on tag set evaluation in Section \ref{sec:prev} and
+\ref{sec:data} provides an overview of the treebank used for
+experimentation. Section \ref{sec:setup} describes the experimental
+setup used in our work and Section \ref{sec:optimization} goes on to
+provide the results from our optimization experiments. Finally,
+Section \ref{sec:summary} summarizes our main findings and discusses
+some avenues for future work.
 

### 2016-05-03 submission.tex

--- 2016-05-03 submission.tex
@@
 \label{sec:intro}
+PoS tagging is an important preprocessing step for many NLP
+tasks, such as parsing, named entity recognition and sentiment
+analysis. Whereas much effort has gone into the development of
+PoS-taggers, to the effect that this task is often considered
+more or less a solved task, considerably less effort has been
+devoted to the empirical evaluation of the PoS tag sets
+themselves.
+PoS tag sets are usually employed in a "one size fits
+all" fashion, regardless of the requirements posed by the
+downstream NLP tasks for which PoS-tagging is performed.
+
+It is well known that syntactic parsing often benefits from quite
+fine-grained morphological distinctions. However, in a realistic
+setting where the aim is to parse raw text, this will often
+require a separate step of morphological analysis in addition to
+PoS-tagging that can be quite expensive.
+
+In this paper, we optimize a Norwegian PoS tag set for the task of
+data-driven dependency parsing.  We show that the introduction of
+morphological distinctions not present in the original tag set, whilst
+compromising tagger accuracy, actually leads to significantly improved
+parsing accuracy.
+
+The article is structured as follows. We start out by reviewing
+previous work on tag set evaluation in Section \ref{sec:prev} and
 

### 2016-05-03 submission.tex
Additions to previous work, small fixes

@@
 all" fashion, regardless of the requirements posed by the
+devoted to empirical evaluation of the PoS tag sets
+themselves.
+PoS tag sets are usually employed in a ``one size fits
+all'' fashion, regardless of the requirements posed by the
 downstream NLP tasks for which PoS-tagging is performed.
@@
 It is well known that syntactic parsing often benefits from quite
-fine-grained morphological distinctions. However, in a realistic
-setting where the aim is to parse raw text, this will often
-require a separate step of morphological analysis in addition to
-PoS-tagging that can be quite expensive.
+fine-grained morphological distinctions since morphology interacts
+with syntax though phenomena such as agreement and case
+marking. However, in a realistic setting where the aim is to
+automatically parse raw text, the generation of morphological
+information will often require a separate step of morphological
+analysis that can be quite costly.
 
@@
 parsing accuracy.
+data-driven dependency parsing. In a set of experiments, various
+morphological distinctions are introduced and evaluated both
+intrinsically, i.e. in terms of PoS-tagging accuracy and
+extrinsically, in terms of parsing accuracy.  Our results show that
+the introduction of morphological distinctions not present in the
+original tag set, whilst compromising tagger accuracy, actually leads
+to significantly improved parsing accuracy. This optimization allows
+us to bypass the additional step of morphological analysis, framing
+the whole preprocessing problem as a simple tagging task.
 
@@
 previous work on tag set evaluation in Section \ref{sec:prev} and
+\ref{sec:data} provides an overview of the treebank used for
+experimentation. Section \ref{sec:setup} describes the experimental
+setup used in our work and Section \ref{sec:optimization} goes on to
+provide the results from our optimization experiments. Finally,
+Section \ref{sec:summary} summarizes our main findings and discusses
+some avenues for future work.
 

### 2016-05-03 submission.tex
Experimental setup nearly finished

@@
 \cite{Meg:01,Meg:02} trained and evaluated a range of PoS taggers on the
-Stockholm-Umeå Corpus (SUC) \cite{Gus:Har:06}, annotated with a Swedish version
-of PAROLE tags, with the tag set totaling 139 tags. Furthermore, they
+Stockholm-Umeå Corpus (SUC) \cite{Gus:Har:06}, annotated with a tag set based
+on a Swedish version of PAROLE tags totaling 139 tags. Furthermore, they
 investigated the effects of tag set size on tagging by mapping the original tag
@@
 tag set as well as the type of information in the tags are crucial factors for
-tagger performance. However/unfortunately, they do not report results of parsing
-with the various PoS tag sets.
+tagger performance. However/unfortunately, they do not report results of
+parsing with the various PoS tag sets.
 
@@
 Treebank to more fine-grained tag sets using linguistic insight to investigate
-whether the additional linguistic information included in the finer-grained
-tags could assist the tagger. Experimenting with both lexically and
-syntactically conditioned modifications such as distinguishing between count
-nouns and noncount nouns as well as between transitive and intransitive verbs,
-they found that more fine-grained tag sets rarely led to improvements in tagger
-accuracy; the most successful modification yielded an improvement in tagger
-accuracy of 0.05 percentage points.
+whether additional linguistic information included in finer-grained tags could
+assist the tagger. Experimenting with both lexically and syntactically
+conditioned modifications such as distinguishing between count nouns and
+noncount nouns and between transitive and intransitive verbs, they found that
+more fine-grained tag sets rarely led to improvements in tagger accuracy; the
+most successful modification yielded an improvement in tagger accuracy of 0.05
+percentage points.
 
@@
 We used the newly developed Norwegian Dependency Treebank (NDT)
-\cite{Sol:Skj:Ovr:14}, which is the first publicly available treebank for
-Norwegian. It was developed at the National Library of Norway in collaboration
-with the University of Oslo, and contains manual syntactic and morphological
-annotation. The treebank contains 311 000 tokens of Bokmål and 303 000 tokens
-of Nynorsk.  The annotated texts are mostly newspaper text, but also include
-government reports, parliament transcripts and excerpts from blogs. The
-annotation process of the treebank was supported by the Oslo-Bergen Tagger
-\cite{Hag:Joh:Nok:00} and then manually corrected by annotators.
+\cite{Sol:Skj:Ovr:14},  the first publicly available treebank for Norwegian. It
+was developed at the National Library of Norway in collaboration with the
+University of Oslo, and contains manual syntactic and morphological annotation.
+The treebank contains data from both varieties of Norwegian; 311 000 tokens of
+Bokmål and 303 000 tokens of Nynorsk. The annotated texts are mostly newspaper
+text, but also include government reports, parliament transcripts and excerpts
+from blogs. The annotation process of the treebank was supported by the
+Oslo-Bergen Tagger \cite{Hag:Joh:Nok:00} and then manually corrected by
+annotators.
 
@@
 modifications, where the coarse PoS tag of relevant tokens are concatenated
-with one or more of these features.  This allows us to include more linguistic
-information in the tags.
+with one or more of these features to include more linguistic information in
+the tags.
 
@@
 Reference Grammar \cite{Faa:Lie:Van:97}. The annotation choices are outlined in
-Table \ref{ndtannotation}, taken from \newcite{Sol:Skj:Ovr:14}, providing
-overview of the analyses of syntactic constructions that often distinguish
-dependency treebanks, such as coordination and the treatment of auxiliary and
-main verbs. The set of dependency relations contains 29 dependency relations,
-including \textsc{adv} (adverbial), \textsc{subj} (subject) and \textsc{koord}
-(coordination).
+Table \ref{ndtannotation}, taken from/courtesy of \newcite{Sol:Skj:Ovr:14},
+providing overview of the analyses of syntactic constructions that often
+distinguish dependency treebanks, such as coordination and the treatment of
+auxiliary and main verbs. The set of dependency relations comprises 29
+dependency relations, including \textsc{adv} (adverbial), \textsc{subj}
+(subject) and \textsc{koord} (coordination).
 
@@
 development, we needed to establish a data set split
-(training/development/test) in preparation to our experiments. Our data set
-split of the treebank follows the standard 80-10-10 (training/development/test)
-split and will be distributed with the treebank and proposed as the new
-standard(?). In creating the data set split, care has been taken to preserve
-contiguous texts in the various data sets while keeping the split balanced in
-terms of genre (and source). Our proposed data set split was used in the
-Norwegian contribution to the Universal Dependencies project \cite{Ovr:Hoh:16}.
-The split will be made available at a companion website.
+(training/development/test). Our data set split of the treebank follows the
+standard 80-10-10 (training/development/test) split and will be distributed
+with the treebank and proposed as the new standard(?). In creating the data set
+split, care has been taken to preserve contiguous texts in the various data
+sets while keeping the split balanced in terms of genre (and source). Our
+proposed data set split was used in the Norwegian contribution to the Universal
+Dependencies project \cite{Ovr:Hoh:16}.  The split will be made available at a
+companion website.
 
 \paragraph{Tagger}
-For our experiments with tag set modifications, we wanted a PoS tagger that is
-both fast and accurate. There is often a trade-off between the two, as the best
-taggers tend to suffer in terms of speed due to their complexity. However, a
-tagger that achieves both close to state-of-the-art accuracy as well as very
-high speed is TnT \cite{Bra:00}. The fact that TnT was used for evaluating the
-universal tag set \cite{Pet:Das:McD:12}, served as another good indication of
-TnT being appropriate for our task. The sum of these factors led to TnT being
-the tagger of choice for our experiments.
+For our experiments with tag set modifications, we wanted a PoS tagger both
+reasonably fast and accurate. There is often a trade-off between the two
+factors/considerations, as the most accurate taggers tend to suffer in terms of
+speed due to their complexity. However, a tagger that achieves both close to
+state-of-the-art accuracy as well as very high speed is TnT \cite{Bra:00}. The
+fact that TnT was used for evaluating the universal tag set
+\cite{Pet:Das:McD:12}, served as another good indication of TnT being
+appropriate for our task. The sum of these factors led to TnT being the tagger
+of choice for our experiments.
 
@@
 the parsing of NDT. Furthermore, recent dependency parser comparisons
-\cite{Cho:Tet:Ste:15} showed that Mate performed very well on parsing of the
-English portion of the OntoNotes 5 corpus, beating a range of contemporary
-state-of-the-art parsers.
+\cite{Cho:Tet:Ste:15} showed that Mate performed very well on parsing of
+English, beating a range of contemporary state-of-the-art parsers.
 
 \paragraph{Tag Set Mapping}
-In order to alter the tag set of NDT, we created a mapping for carrying out the
-tag set modifications. We created a mapping for carrying out the tag set
-modifications that maps the relevant existing tags to new, more fine-grained
-tags including more relevant morphological features for the applicable tokens.
+In order to carrying out the tag set modifications, we created a mapping that
+maps the relevant existing tags to new, more fine-grained tags including more
+relevant morphological features for the applicable tokens. A given tag set
+modification involves specifying a tag and one or more features to be
+concatenated with said tag for all applicable tokens, i.e., tokens that are
+assigned said tag and contains the set of features in its morphological
+features.
 
@@
 reveal that the combination of training and testing on automatic tags is
-superior to training on gold standard tags and testing on automatic tags,
-surprisingly. Consequently, the parser was both trained and tested on
-automatically assigned tags in our experiments.
+superior to training on gold standard tags and testing on automatic tags.
+Consequently, the parser was both trained and tested on automatically assigned
+tags in our experiments.
 
-Note that it is absolutely crucial that the morphological features in the
-treebank are removed when using automatic tags, as they are still gold
-standard. For instance, if a verb token is erroneously tagged as a noun, we
-could potentially have a noun token with verbal features such as tense, which
-markedly obfuscates the training and parsing. Another important factor is that
-we want to isolate the effect of PoS tags, necessitating the exclusion of
-morphological features.
+We removed any morphological features in order to simulate a realistic setting.
+Moreover, it is crucial that we remove these features when working with
+automatically assigned tags, as the features is still gold standard.  For
+instance, if a verb token is erroneously tagged as a noun, we could potentially
+have a noun token with verbal features such as tense, which markedly obfuscates
+the training and parsing. Another important aspect is that we want to isolate
+the effect of PoS tags, necessitating the exclusion of morphological features.
+
+%Note that it is absolutely crucial that the morphological features in the
+%treebank are removed when using automatic tags, as they are still gold
+%standard. For instance, if a verb token is erroneously tagged as a noun, we
+%could potentially have a noun token with verbal features such as tense, which
+%markedly obfuscates the training and parsing. Another important factor is that
+%we want to isolate the effect of PoS tags, necessitating the exclusion of
+%morphological features.
 

### 2016-05-04 submission.tex
Added some refs

@@
 \label{sec:intro}
-PoS tagging is an important preprocessing step for many NLP
-tasks, such as parsing, named entity recognition and sentiment
+PoS tagging is an important preprocessing step for many NLP tasks,
+such as parsing, named entity recognition and sentiment
 analysis. Whereas much effort has gone into the development of
-PoS-taggers, to the effect that this task is often considered
-more or less a solved task, considerably less effort has been
-devoted to the empirical evaluation of the PoS tag sets
-themselves.
-PoS tag sets are usually employed in a "one size fits
-all" fashion, regardless of the requirements posed by the
-devoted to empirical evaluation of the PoS tag sets
-themselves.
-PoS tag sets are usually employed in a ``one size fits
-all'' fashion, regardless of the requirements posed by the
-downstream NLP tasks for which PoS-tagging is performed.
+PoS-taggers, to the effect that this task is often considered more or
+less a solved task, considerably less effort has been devoted to the
+empirical evaluation of the PoS tag sets themselves. Error analysis of
+PoS-taggers indicate that, whereas tagger improvement though means of
+learning algorithm or feature engineering seems to have reached
+something of a plateau, linguistic assessment of the distinctions made
+by the PoS tag set may be a venue worth investigating further
+\cite{Man:11}.  Clearly, the utility of a PoS tag set is tightly
+coupled with the the downstream task for which it is performed. Even
+so, PoS tag sets are usually employed in a ``one size fits all''
+fashion, regardless of the requirements posed by the task which makes
+use of this information.
 
 It is well known that syntactic parsing often benefits from quite
-fine-grained morphological distinctions since morphology interacts
-with syntax though phenomena such as agreement and case
-marking. However, in a realistic setting where the aim is to
-automatically parse raw text, the generation of morphological
-information will often require a separate step of morphological
-analysis that can be quite costly.
+fine-grained morphological distinctions \cite{Zha:Niv:11}. Morphology
+interacts with syntax though phenomena such as agreement and case
+marking and incorporating information on morphological properties of
+words can therefore often improved parsing performance. However, in a
+realistic setting where the aim is to automatically parse raw text,
+the generation of morphological information will often require a
+separate step of morphological analysis that can be quite costly.
 
@@
 compromising tagger accuracy, actually leads to significantly improved
-parsing accuracy.
-data-driven dependency parsing. In a set of experiments, various
-morphological distinctions are introduced and evaluated both
-intrinsically, i.e. in terms of PoS-tagging accuracy and
-extrinsically, in terms of parsing accuracy.  Our results show that
-the introduction of morphological distinctions not present in the
-original tag set, whilst compromising tagger accuracy, actually leads
-to significantly improved parsing accuracy. This optimization allows
-us to bypass the additional step of morphological analysis, framing
-the whole preprocessing problem as a simple tagging task.
+parsing accuracy.  In a set of experiments, various morphological
+distinctions are introduced and evaluated both intrinsically, i.e. in
+terms of PoS-tagging accuracy, and extrinsically, in terms of parsing
+accuracy.  Our results show that the introduction of morphological
+distinctions not present in the original tag set, whilst compromising
+tagger accuracy, actually leads to significantly improved parsing
+accuracy. This optimization allows us to bypass the additional step of
+morphological analysis, framing the whole preprocessing problem as a
+simple tagging task.
 

### 2016-05-04 submission.tex
More refs to intro

@@
 PoS tagging is an important preprocessing step for many NLP tasks,
-such as parsing, named entity recognition and sentiment
-analysis. Whereas much effort has gone into the development of
-PoS-taggers, to the effect that this task is often considered more or
-less a solved task, considerably less effort has been devoted to the
-empirical evaluation of the PoS tag sets themselves. Error analysis of
-PoS-taggers indicate that, whereas tagger improvement though means of
-learning algorithm or feature engineering seems to have reached
-something of a plateau, linguistic assessment of the distinctions made
-by the PoS tag set may be a venue worth investigating further
-\cite{Man:11}.  Clearly, the utility of a PoS tag set is tightly
-coupled with the the downstream task for which it is performed. Even
-so, PoS tag sets are usually employed in a ``one size fits all''
-fashion, regardless of the requirements posed by the task which makes
-use of this information.
+such as dependency parsing \cite{Niv:Hal:Kub:07,Haj:Cia:Joh:09}, named
+entity recognition and sentiment analysis
+\cite{Wil:Wie:Hof:09}. Whereas much effort has gone into the
+development of PoS-taggers, to the effect that this task is often
+considered more or less a solved task, considerably less effort has
+been devoted to the empirical evaluation of the PoS tag sets
+themselves. Error analysis of PoS-taggers indicate that, whereas
+tagger improvement though means of learning algorithm or feature
+engineering seems to have reached something of a plateau, linguistic
+assessment of the distinctions made by the PoS tag set may be a venue
+worth investigating further \cite{Man:11}.  Clearly, the utility of a
+PoS tag set is tightly coupled with the the downstream task for which
+it is performed. Even so, PoS tag sets are usually employed in a ``one
+size fits all'' fashion, regardless of the requirements posed by the
+task which makes use of this information.
 

### 2016-05-04 submission.tex
More refs to intro

@@
 such as dependency parsing \cite{Niv:Hal:Kub:07,Haj:Cia:Joh:09}, named
-entity recognition and sentiment analysis
+entity recognition \cite{Tjo:DeM:03} and sentiment analysis
 \cite{Wil:Wie:Hof:09}. Whereas much effort has gone into the
@@
 It is well known that syntactic parsing often benefits from quite
-fine-grained morphological distinctions \cite{Zha:Niv:11}. Morphology
-interacts with syntax though phenomena such as agreement and case
-marking and incorporating information on morphological properties of
-words can therefore often improved parsing performance. However, in a
-realistic setting where the aim is to automatically parse raw text,
-the generation of morphological information will often require a
-separate step of morphological analysis that can be quite costly.
+fine-grained morphological distinctions
+\cite{Zha:Niv:11,See:Kuh:13}. Morphology interacts with syntax though
+phenomena such as agreement and case marking and incorporating
+information on morphological properties of words can therefore often
+improved parsing performance. However, in a realistic setting where
+the aim is to automatically parse raw text, the generation of
+morphological information will often require a separate step of
+morphological analysis that can be quite costly.
 

### 2016-05-04 submission.tex
Minor changes

@@
 previous work on tag set evaluation in Section \ref{sec:prev} and
-\ref{sec:data} provides an overview of the treebank used for
+Section \ref{sec:data} provides an overview of the treebank used for
 experimentation. Section \ref{sec:setup} describes the experimental
@@
 \label{sec:prev}
-There is a considerable lack of previous efforts on extrinsic evaluation of the
-effects of PoS tag sets on downstream applications. We will consider(/noe
-annet, considerable...consider) papers that evaluate the effects of PoS tag
-set granularity on PoS tagging. This includes investigation of the effects of
-PoS tag sets on tagging of Swedish \cite{Meg:01,Meg:02} and English
-\cite{Mac:05}.
+There has been little previous work dedicated specifically to
+extrinsic evaluation of the effects of PoS tag sets on downstream
+applications.  There has, however, been some work that evaluates the
+effects of PoS tag set granularity on PoS tagging. This includes
+investigation of the effects of PoS tag sets on tagging of Swedish
+\cite{Meg:01,Meg:02} and English \cite{Mac:05}.
 

### 2016-05-04 submission.tex
Minor changes

@@
 
-\section{Data}
+\section{The Norwegian Dependency Treebank}
 \label{sec:data}
-\subsection{The Norwegian Dependency Treebank}
 We used the newly developed Norwegian Dependency Treebank (NDT)
-\cite{Sol:Skj:Ovr:14},  the first publicly available treebank for Norwegian. It
-was developed at the National Library of Norway in collaboration with the
-University of Oslo, and contains manual syntactic and morphological annotation.
-The treebank contains data from both varieties of Norwegian; 311 000 tokens of
-Bokmål and 303 000 tokens of Nynorsk. The annotated texts are mostly newspaper
-text, but also include government reports, parliament transcripts and excerpts
-from blogs. The annotation process of the treebank was supported by the
-Oslo-Bergen Tagger \cite{Hag:Joh:Nok:00} and then manually corrected by
-annotators.
+\cite{Sol:Skj:Ovr:14}, the first publicly available treebank for
+Norwegian. It was developed at the National Library of Norway in
+collaboration with the University of Oslo, and contains manual
+syntactic and morphological annotation.  The treebank contains data
+from both varieties of Norwegian; 311 000 tokens of Bokmål and 303 000
+tokens of Nynorsk. We will in the following only be using the Bokmål
+portion of the treebank. The annotated texts are mostly newspaper
+text, but also include government reports, parliament transcripts and
+excerpts from blogs. The annotation process of the treebank was
+supported by the rule-based Oslo-Bergen Tagger \cite{Hag:Joh:Nok:00}
+and then manually corrected by annotators, adding syntactic dependency analysis
+to the morphosyntactic annotation.
 

### 2016-05-04 submission.tex
Minor changes

@@
     \end{tabular}
-    \caption{Annotation choices in NDT.}
+    \caption{Central head-dependent annotation choices in NDT.}
     \label{ndtannotation}
@@
 speed due to their complexity. However, a tagger that achieves both close to
-state-of-the-art accuracy as well as very high speed is TnT \cite{Bra:00}. The
-fact that TnT was used for evaluating the universal tag set
-\cite{Pet:Das:McD:12}, served as another good indication of TnT being
-appropriate for our task. The sum of these factors led to TnT being the tagger
+state-of-the-art accuracy as well as very high speed is TnT \cite{Bra:00}. TnT was furthermore recently used to evaluate the recently proposed universal tag set
+\cite{Pet:Das:McD:12}. The sum of these factors led to TnT being the tagger
 of choice for our experiments.
@@
 \paragraph{Tags \& Features}
-As we seek to quantify the effects of PoS tagging in a realistic setting, we
-want to run the parser on automatically assigned PoS tags. For the training of
-the parser, however, we have two options: using either gold standard or
+As we seek to quantify the effects of PoS tagging in a realistic
+setting, i.e. in application to raw text, we want to evaluate the parser on
+automatically assigned PoS tags. For the training of the parser,
+however, we have two options: using either gold standard or
 automatically assigned tags. In order to settle on a configuration, we
-conducted experiments with gold standard and automatically assigned tags to see
-how they differ with respect to performance. The results of our experiments
-reveal that the combination of training and testing on automatic tags is
-superior to training on gold standard tags and testing on automatic tags.
-Consequently, the parser was both trained and tested on automatically assigned
-tags in our experiments.
+conducted experiments with gold standard and automatically assigned
+tags to see how they differ with respect to performance. The results
+of our experiments reveal that the combination of training and testing
+on automatic tags is superior to training on gold standard tags and
+testing on automatic tags.  Consequently, the parser was both trained
+and tested on automatically assigned tags in our experiments.
 

### 2016-05-04 submission.tex
Transition to optimization
More on optimization

--- 2016-05-04 submission.tex
@@
 been devoted to the empirical evaluation of the PoS tag sets
-themselves. Error analysis of PoS-taggers indicate that, whereas
-tagger improvement though means of learning algorithm or feature
-engineering seems to have reached something of a plateau, linguistic
-assessment of the distinctions made by the PoS tag set may be a venue
-worth investigating further \cite{Man:11}.  Clearly, the utility of a
-PoS tag set is tightly coupled with the the downstream task for which
-it is performed. Even so, PoS tag sets are usually employed in a ``one
-size fits all'' fashion, regardless of the requirements posed by the
-task which makes use of this information.
+themselves. Error analysis of PoS taggers indicate that, whereas tagger
+improvement though means of learning algorithm or feature engineering seems to
+have reached something of a plateau, linguistic assessment of the distinctions
+made by the PoS tag set may be a venue worth investigating further
+\cite{Man:11}. Clearly, the utility of a PoS tag set is tightly coupled with
+the downstream task for which it is performed. Even so, PoS tag sets are
+usually employed in a ``one size fits all'' fashion, regardless of the
+requirements posed by the task which makes use of this information.
 
@@
 distinctions are introduced and evaluated both intrinsically, i.e. in
-terms of PoS-tagging accuracy, and extrinsically, in terms of parsing
+terms of PoS tagging accuracy, and extrinsically, in terms of parsing
 accuracy.  Our results show that the introduction of morphological
@@
 
-The article is structured as follows. We start out by reviewing
-previous work on tag set evaluation in Section \ref{sec:prev} and
-Section \ref{sec:data} provides an overview of the treebank used for
-experimentation. Section \ref{sec:setup} describes the experimental
-setup used in our work and Section \ref{sec:optimization} goes on to
-provide the results from our optimization experiments. Finally,
-Section \ref{sec:summary} summarizes our main findings and discusses
-some avenues for future work.
+The article is structured as follows. We start out by reviewing previous work
+on tag set evaluation in Section \ref{sec:prev} and Section \ref{sec:data}
+provides an overview of the treebank used for experimentation. Section
+\ref{sec:setup} describes the experimental setup used in our work and Section
+\ref{sec:optimization} goes on to provide the results from our optimization
+experiments. Finally, Section \ref{sec:summary} summarizes our main findings
+and discusses some avenues for future work.
 
@@
 \label{sec:prev}
-There has been little previous work dedicated specifically to
-extrinsic evaluation of the effects of PoS tag sets on downstream
-applications.  There has, however, been some work that evaluates the
-effects of PoS tag set granularity on PoS tagging. This includes
-investigation of the effects of PoS tag sets on tagging of Swedish
-\cite{Meg:01,Meg:02} and English \cite{Mac:05}.
+There is a considerable lack of previous efforts on extrinsic evaluation of the
+effects of PoS tag sets on downstream applications. We will consider(/noe
+annet, considerable...consider) papers that evaluate the effects of PoS tag
+set granularity on PoS tagging. This includes investigation of the effects of
+PoS tag sets on tagging of Swedish \cite{Meg:01,Meg:02} and English
+\cite{Mac:05}.
 
@@
     \end{tabular}
-    \caption{Annotation choices in NDT.}
+    \caption{Syntactic annotation choices in NDT.}
     \label{ndtannotation}
@@
 \paragraph{Tag Set Mapping}
-In order to carrying out the tag set modifications, we created a mapping that
-maps the relevant existing tags to new, more fine-grained tags including more
-relevant morphological features for the applicable tokens. A given tag set
-modification involves specifying a tag and one or more features to be
-concatenated with said tag for all applicable tokens, i.e., tokens that are
-assigned said tag and contains the set of features in its morphological
-features.
+In order to carry out the tag set modifications, we created a mapping that
+takes as input a set of tags and associated morphological features to be
+included in said tags. This mapping maps the relevant existing tags to new,
+more fine-grained tags including more relevant morphological features for the
+applicable tokens. This is done by concatenating the original tag with said
+features for all applicable tokens, i.e., tokens that are assigned said tag and
+contains the set of features in its morphological features.
 
@@
 \label{sec:optimization}
-With more fine-grained linguistically
-motivated distinctions, we increase the linguistic information represented in
-the tags, which may assist the tagger in disambiguating ambiguous and unknown
-words, which in turn may aid the parser in recognizing and generalizing
-syntactic patterns. However, the addition of more linguistic information to the
-tags and thus a more fine-grained tag set will most likely lead to a drop in
-tagger accuracy, due to the increase in complexity. The best tagging does not
+By introducing more fine-grained linguistically motivated distinctions in a tag
+set, we increase the linguistic information represented in the tags, which may
+assist the parser in recognizing and generalizing syntactic patterns. However,
+the addition of more linguistic information to the tags and thus a more
+fine-grained tag set will most likely lead to a drop in tagger accuracy/tagging
+quality due to the increase in complexity. The best tagging does not
 necessarily lead to the best parse, and it is therefore interesting to
 investigate how the tag set modifications may affect the interplay between
-tagging and parsing.
+tagging and parsing. As we want to investigate how we can improve the
+linguistic quality of a tag set, we will not introduce tag set modifications
+which are not linguistically motivated. We will only consider distinctions we
+deem linguistically sensible, even if we expect them to impair the
+computational tractability; investigating the interplay between these two
+considerations is ultimately our goal.
+
+\subsection{Initial/Baseline Experiments}
+In an initial round of experiments, we concatenated the tag of each token with
+its set of morphological features in order to map the original tag set to a
+new, more fine-grained tag set (hereafter referred to as the \emph{full} tag
+set). The result of this was a total of 368 tags, which is clearly very
+fine-grained. The two initial tag sets, i.e., the original tag set comprising
+19 tags and the full tag set comprising 368 tags, thus represent two extremes
+in terms of granularity.  To evaluate these tag sets and investigate how the tag
+set granularity would affect the performance of tagging and parsing, we trained
+and evaluated TnT and Mate on the training and development data, respectively,
+on the two tag sets. In Table \ref{inittagseteval}, we report the results of
+these experiments. We see that the tagger accuracy drastically drops when going
+from the original to the full tag set. The MFT baseline for the original tag
+set is 94.14\%, while it drops by almost 9 percentage points to 85.15\% for the
+full tag set. TnT reports an accuracy of 97.47\% on the original tag set, which
+is reduced to 93.48\% for the full tag set. These results confirm our
+hypothesis that the very high linguistic quality in the full, fine-grained tag
+set comes at the expense of drops in tagger performance. However, the
+additional linguistic information provided by the full tag set improves the
+parser performance. With the original tag set, Mate reports an LAS of 87.01\%
+and a UAS of 90.19\%, which increases to 87.15\% and 90.39\%, respectively,
+when using the full tag set. As we are looking for linguistically informed
+distinctions that improve the syntactic parsing, these results are promising
+and serve to indicate that additional morphological information assists the
+syntactic parsers, motivating the optimization of the existing PoS tag set.
+
+\begin{table}
+    \centering
+    \smaller[0.5]
+    \begin{tabular}{@{}lllll@{}}
+        \toprule
+        \textbf{Tag set} & \textbf{MFT} & \textbf{Accuracy} &
+        \textbf{LAS} & \textbf{UAS} \\
+        \midrule
+        Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
+        Full & 85.15\% & 93.48\% & \textbf{87.15\%} & \textbf{90.39\%} \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of tagging and parsing with the two initial tag sets.}
+    \label{inittagseteval}
+\end{table}
 
@@
         Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
-        Full & 85.12\% & 93.46\% & 87.13\% & 90.32\% \\
+        Full & 85.15\% & 93.48\% & 87.15\% & 90.39\% \\
         Optimized & 89.20\% & 96.85\% & \textbf{88.87\%} & \textbf{91.78\%} \\
@@
     \end{tabular}
-    \caption{Results of tagging and parsing with the optimized tag set, compared to
-        the initial tag sets.}
+    \caption{Results of tagging and parsing with the optimized tag set,
+        compared to the initial tag sets.}
     \label{finalresults}
@@
 
-\section{Optimized Pipeline}
-\label{sec:pipeline}
+%\section{Optimized Pipeline}
+%\label{sec:pipeline}
 

### 2016-05-04 submission.tex

@@
 and discusses some avenues for future work.
+The article is structured as follows. We start out by reviewing
+previous work on tag set evaluation in Section \ref{sec:prev} and
+Section \ref{sec:data} provides an overview of the treebank used for
+experimentation. Section \ref{sec:setup} describes the experimental
+setup used in our work and Section \ref{sec:optimization} goes on to
+provide the results from our optimization experiments. Finally,
+Section \ref{sec:summary} summarizes our main findings and discusses
+some avenues for future work.
 
@@
     \end{tabular}
-    \caption{Syntactic annotation choices in NDT.}
+    %\caption{Syntactic annotation choices in NDT.}
+    \caption{Central head-dependent annotation choices in NDT.}
     \label{ndtannotation}
@@
 speed due to their complexity. However, a tagger that achieves both close to
-state-of-the-art accuracy as well as very high speed is TnT \cite{Bra:00}. The
-fact that TnT was used for evaluating the universal tag set
-\cite{Pet:Das:McD:12}, served as another good indication of TnT being
-appropriate for our task. The sum of these factors led to TnT being the tagger
+state-of-the-art accuracy as well as very high speed is TnT \cite{Bra:00}. TnT
+was furthermore recently used to evaluate the recently proposed universal tag
+set \cite{Pet:Das:McD:12}. The sum of these factors led to TnT being the tagger
 of choice for our experiments.
@@
 \end{table}
+
+In the following, we report on a set of experiments which evaluate
+increasingly fine-grained PoS tag sets. The PoS tag sets are motivated
+by morphosyntactic properties of Norwegian and are evaluated both in
+terms of tagger and parser accuracy.
 

### 2016-05-04 submission.tex
mindre endringer i introduksjonen

@@
 \label{sec:intro}
-PoS tagging is an important preprocessing step for many NLP tasks,
+Part-of-speech (PoS) tagging is an important preprocessing step for many NLP tasks,
 such as dependency parsing \cite{Niv:Hal:Kub:07,Haj:Cia:Joh:09}, named
@@
 \cite{Wil:Wie:Hof:09}. Whereas much effort has gone into the
-development of PoS-taggers, to the effect that this task is often
-considered more or less a solved task, considerably less effort has
+development of PoS-taggers -- to the effect that this task is often
+considered more or less a solved task -- considerably less effort has
 been devoted to the empirical evaluation of the PoS tag sets
-themselves. Error analysis of PoS taggers indicate that, whereas tagger
-improvement though means of learning algorithm or feature engineering seems to
-have reached something of a plateau, linguistic assessment of the distinctions
-made by the PoS tag set may be a venue worth investigating further
-\cite{Man:11}. Clearly, the utility of a PoS tag set is tightly coupled with
-the downstream task for which it is performed. Even so, PoS tag sets are
-usually employed in a ``one size fits all'' fashion, regardless of the
-requirements posed by the task which makes use of this information.
+themselves. Error analysis of PoS-taggers indicate that, whereas
+tagging improvement through means of learning algorithm or feature
+engineering seems to have reached something of a plateau, linguistic
+and empirical assessment of the distinctions made in the PoS tag set may be a venue
+worth investigating further \cite{Man:11}.  Clearly, the utility of a
+PoS tag set is tightly coupled with the the downstream task for which
+it is performed. Even so, PoS tag sets are usually employed in a ``one
+size fits all'' fashion, regardless of the requirements posed by the
+task which makes use of this information.
 
@@
 \cite{Zha:Niv:11,See:Kuh:13}. Morphology interacts with syntax though
-phenomena such as agreement and case marking and incorporating
+phenomena such as agreement and case marking, and incorporating
 information on morphological properties of words can therefore often
@@
 
-In this paper, we optimize a Norwegian PoS tag set for the task of
-data-driven dependency parsing.  We show that the introduction of
-morphological distinctions not present in the original tag set, whilst
-compromising tagger accuracy, actually leads to significantly improved
-parsing accuracy.  In a set of experiments, various morphological
-distinctions are introduced and evaluated both intrinsically, i.e. in
-terms of PoS tagging accuracy, and extrinsically, in terms of parsing
+In this paper, we optimize a PoS tag set for the task of
+data-driven dependency parsing of Norwegian. We report on a set of experiments where various morphological
+distinctions are introduced to the PoS annotations and evaluated both intrinsically, i.e. in
+terms of PoS-tagging accuracy, and extrinsically, in terms of parsing
 accuracy.  Our results show that the introduction of morphological
@@
 
-The article is structured as follows. We start out by reviewing previous work
-on tag set evaluation in Section \ref{sec:prev} and Section \ref{sec:data}
-provides an overview of the treebank used for experimentation. Section
-\ref{sec:setup} describes the experimental setup used in our work and Section
-\ref{sec:optimization} goes on to provide the results from our optimization
-experiments. Finally, Section \ref{sec:summary} summarizes our main findings
-and discusses some avenues for future work.
 The article is structured as follows. We start out by reviewing
-previous work on tag set evaluation in Section \ref{sec:prev} and
-Section \ref{sec:data} provides an overview of the treebank used for
-experimentation. Section \ref{sec:setup} describes the experimental
+previous work on tag set evaluation in Section \ref{sec:prev}, while
+Section \ref{sec:data} details the treebank that provides the initial gold 
+annotations used for our experiments. Section \ref{sec:setup} describes the experimental
 setup used in our work and Section \ref{sec:optimization} goes on to
-provide the results from our optimization experiments. Finally,
+provide the results from our tag set optimization. Finally,
 Section \ref{sec:summary} summarizes our main findings and discusses

### 2016-05-04 submission.tex
More on optimization

--- 2016-05-04 submission.tex
@@
 text, but also include government reports, parliament transcripts and
-excerpts from blogs. The annotation process of the treebank was
-supported by the rule-based Oslo-Bergen Tagger \cite{Hag:Joh:Nok:00}
-and then manually corrected by annotators, adding syntactic dependency analysis
-to the morphosyntactic annotation.
-
-%Linguistic motivation, insight, utility...
+excerpts from blogs. The annotation process of the treebank was supported by
+the rule-based Oslo-Bergen Tagger \cite{Hag:Joh:Nok:00} and then manually
+corrected by annotators, adding syntactic dependency analysis to the
+morphosyntactic annotation.
 
@@
 definiteness, number and tense. These features are used in our tag set
-modifications, where the coarse PoS tag of relevant tokens are concatenated
-with one or more of these features to include more linguistic information in
-the tags.
+modifications, where the coarse PoS tag of relevant tokens is concatenated with
+one or more of these features to include more linguistic information in the
+tags.
 
@@
 
-\begin{table} \centering
+\begin{table}
+    \centering
     %\smaller[0.5]
@@
     \end{tabular}
-    %\caption{Syntactic annotation choices in NDT.}
     \caption{Central head-dependent annotation choices in NDT.}
@@
 \paragraph{Tags \& Features}
-As we seek to quantify the effects of PoS tagging in a realistic
-setting, i.e. in application to raw text, we want to evaluate the parser on
-automatically assigned PoS tags. For the training of the parser,
-however, we have two options: using either gold standard or
-automatically assigned tags. In order to settle on a configuration, we
-conducted experiments with gold standard and automatically assigned
-tags to see how they differ with respect to performance. The results
-of our experiments reveal that the combination of training and testing
-on automatic tags is superior to training on gold standard tags and
-testing on automatic tags.  Consequently, the parser was both trained
-and tested on automatically assigned tags in our experiments.
+As we seek to quantify the effects of PoS tagging in a realistic setting, i.e.,
+in application to raw text, we want to evaluate the parser on automatically
+assigned PoS tags. For the training of the parser, however, we have two
+options: using either gold standard or automatically assigned tags. In order to
+settle on a configuration, we conducted experiments with gold standard and
+automatically assigned tags to see how they differ with respect to performance.
+The results of our experiments reveal that the combination of training and
+testing on automatic tags is superior to training on gold standard tags and
+testing on automatic tags.  Consequently, the parser was both trained and
+tested on automatically assigned tags in our experiments.
 
@@
 
-%Note that it is absolutely crucial that the morphological features in the
-%treebank are removed when using automatic tags, as they are still gold
-%standard. For instance, if a verb token is erroneously tagged as a noun, we
-%could potentially have a noun token with verbal features such as tense, which
-%markedly obfuscates the training and parsing. Another important factor is that
-%we want to isolate the effect of PoS tags, necessitating the exclusion of
-%morphological features.
+In the following, we report on a set of experiments which evaluate
+increasingly fine-grained PoS tag sets. The PoS tag sets are motivated
+by morphosyntactic properties of Norwegian and are evaluated both in
+terms of tagger and parser accuracy.
 
@@
 these experiments. We see that the tagger accuracy drastically drops when going
-from the original to the full tag set. The MFT baseline for the original tag
-set is 94.14\%, while it drops by almost 9 percentage points to 85.15\% for the
-full tag set. TnT reports an accuracy of 97.47\% on the original tag set, which
-is reduced to 93.48\% for the full tag set. These results confirm our
-hypothesis that the very high linguistic quality in the full, fine-grained tag
-set comes at the expense of drops in tagger performance. However, the
-additional linguistic information provided by the full tag set improves the
-parser performance. With the original tag set, Mate reports an LAS of 87.01\%
-and a UAS of 90.19\%, which increases to 87.15\% and 90.39\%, respectively,
-when using the full tag set. As we are looking for linguistically informed
-distinctions that improve the syntactic parsing, these results are promising
-and serve to indicate that additional morphological information assists the
-syntactic parsers, motivating the optimization of the existing PoS tag set.
+from the original to the full tag set. TnT reports an accuracy of 97.47\% on
+the original tag set, which is reduced to 93.48\% for the full tag set. These
+results confirm our hypothesis that the very high linguistic quality in the
+full, fine-grained tag set comes at the expense of drops in tagger performance.
+However, the additional linguistic information provided by the full tag set
+improves the parser performance. With the original tag set, Mate reports an LAS
+of 87.01\% and a UAS of 90.19\%, which increases to 87.15\% and 90.39\%,
+respectively, when using the full tag set. As we are looking for linguistically
+informed distinctions that improve the syntactic parsing, these results are
+promising and serve to indicate that additional morphological information
+assists the syntactic parsers, motivating the optimization of the existing PoS
+tag set.
 
@@
 
-In the following, we report on a set of experiments which evaluate
-increasingly fine-grained PoS tag sets. The PoS tag sets are motivated
-by morphosyntactic properties of Norwegian and are evaluated both in
-terms of tagger and parser accuracy.
-
 \subsection{Tag Set Experiments}
@@
 
+The morphological properties of the various parts-of-speech is reflected in the
+morphological features associated with the respective PoS tags. In addition to
+morphological properties such as definiteness, tense and number, all classes
+except for verbs can be distinguished "on" type, e.g., common nouns and proper
+nouns, while pronouns can be reflexive, reciprocal, personal or interrogative.
+
+\subsection{Experiments}
 \paragraph{Nouns}
@@
 is agreement in gender, definiteness and number between nouns and their
-modifiers (adjectives and determiners).
+modifiers (adjectives and determiners). Additionally, NDT has a separate case
+feature to distinguishing nouns in genitive case.
+
+Experimenting with nouns, we found that all tag set modifications yielded large
+increases in LAS and UAS. The most informative features are definiteness, which
+leads to an increase in LAS by 1.26 percentage points, to 88.27\%, and type,
+yielding an LAS of 88.07\%. Turning to combinations of features, we found that
+the combination of type and case, as well as type and definiteness, were the
+most promising, which led us to combine type, case and definiteness in a final
+experiment, resulting in LAS of 88.81\% and UAS of 91.73\%, constituting large
+increases from parsing with the original tag set, 1.80 percentage points
+and 1.54 percentage points, respectively.
 
@@
 in Norwegian and can additionally take on mood (imperative, indicative) and
-voice (active or passive).
+voice (active or passive). Note that voice and mood have only a single value,
+\texttt{pass} (passive) and \texttt{imp} (imperative), respectively. Verbs
+which are not passive are implicitly active, and verbs which are not imperative
+are indicative.
+
+Introducing features in isolation, mood is the only feature leading to
+increases in LAS, with a reported LAS of 87.04\%. Imperative clauses are
+fundamentally different from indicative clauses, as they lack an overt subject.
+Tense yields an LAS of 86.97\% while voice yields an LAS of 86.96\%. Combining
+mood and tense resulted in an LAS of 87.12\% and UAS of 90.31\%.
+
+In an additional experiment, we mapped the verb tenses (mood, in the case of
+imperative) to finiteness.  All verbs have finiteness, hence this distinction
+has broad coverage.  This mapping is syntactically grounded as finite verbs and
+nonfinite verbs appear in completely different syntactic construction, and
+proved to greatly improve the parsing, as we saw the overall largest parser
+accuracy scores, with 87.30\% for LAS and 90.43\% for UAS, 0.29 and 0.24
+percentage points higher than the baseline, respectively.  This coincides with
+the observations seen for Swedish in \newcite{Ovr:08}, where finiteness was found
+to be a very beneficial linguistic feature for parsing.
 
@@
 Adjectives agree with the noun they modify in terms of gender, number and
-definiteness in Norwegian.
+definiteness in Norwegian. Adjectives are also inflected for degree, either
+positive, comparative or superlative.
 
@@
 Like adjectives, determiners in Norwegian agree with the noun they modify in
-terms of gender, number and definiteness.
+terms of gender, number and definiteness. Possessive pronouns are
+treated/analyzed as determiners in NDT.
 

### 2016-05-04 submission.tex
Added some results from tag set experiments

@@
 
-\cite{Meg:01,Meg:02} trained and evaluated a range of PoS taggers on the
-Stockholm-Umeå Corpus (SUC) \cite{Gus:Har:06}, annotated with a tag set based
-on a Swedish version of PAROLE tags totaling 139 tags. Furthermore, they
-investigated the effects of tag set size on tagging by mapping the original tag
-set into smaller subsets designed for parsing. They argue that a tag set with
-complete morphological tags may not be necessary for all NLP applications, for
-instance syntactic parsing. They found that the smallest tag set comprising 26
-tags yields the lowest tagger error rate. However, for some of the taggers,
-augmenting the tag set with more linguistically informative tags may actually
-lead to a drop in error rate. They argue that this shows that the size of the
-tag set as well as the type of information in the tags are crucial factors for
-tagger performance. However/unfortunately, they do not report results of
-parsing with the various PoS tag sets.
+\newcite{Meg:01} and \newcite{Meg:02} trained and evaluated a range of PoS
+taggers on the Stockholm-Umeå Corpus (SUC) \cite{Gus:Har:06}, annotated with a
+tag set based on a Swedish version of PAROLE tags totaling 139 tags.
+Furthermore, they investigated the effects of tag set size on tagging by
+mapping the original tag set into smaller subsets designed for parsing. They
+argue that a tag set with complete morphological tags may not be necessary for
+all NLP applications, for instance syntactic parsing. They found that the
+smallest tag set comprising 26 tags yields the lowest tagger error rate.
+However, for some of the taggers, augmenting the tag set with more
+linguistically informative tags may actually lead to a drop in error rate. They
+argue that this shows that the size of the tag set as well as the type of
+information in the tags are crucial factors for tagger performance.
+However/unfortunately, they do not report results of parsing with the various
+PoS tag sets.
 
@@
 expect them to have similar properties.
-%, there are two main
-%approaches: assign it the most frequent tag overall in the training data, or
-%the tag most frequently assigned to words seen only once in the training data.
-%We adopted the latter approach for our baseline tagger, as we believe that
-%unknown words may have much in common with words that occur only once in the
-%training, more so than simply words of the most frequent part-of-speech. The
-%reason for this is that unknown and infrequent words have in common that they
-%rarely occur, and we might therefore expect them to have similar properties.
 
@@
 The morphological properties of the various parts-of-speech is reflected in the
-morphological features associated with the respective PoS tags. In addition to
+morphological features associated with the respective PoS tags. For instance,
+as nouns can take on gender, definiteness and number, additionally "being of"
+genitive case, the treebank operates with features for gender, definiteness,
+number and case with "accompanying" values. In addition to
 morphological properties such as definiteness, tense and number, all classes
@@
 \subsection{Experiments}
+"Nearly" all tag set modifications leads to a drop in tagger accuracy, as
+expected. Few exceptions...
+
 \paragraph{Nouns}
@@
 and 1.54 percentage points, respectively.
+
+%\begin{table}
+    %\centering
+    %\smaller[0.5]
+    %\begin{tabular}{@{}lllll@{}}
+        %\toprule
+        %\textbf{Feature(s)} & \textbf{MFT} & \textbf{Accuracy} &
+        %\textbf{LAS} & \textbf{UAS} \\
+        %\midrule
+        %--- & \textbf{94.14\%} & 97.47\% & 87.01\% & 90.19\% \\
+        %Case & 93.77\% & \textbf{97.48\%} & 87.63\% & 90.72\% \\
+        %Definiteness & 89.67\% & 97.00\% & 88.27\% & 91.42\% \\
+        %Gender & 89.54\% & 96.09\% & 87.21\% & 90.36\% \\
+        %Number & 90.04\% & 96.37\% & 87.97\% & 91.00\% \\
+        %Type & 91.90\% & 96.92\% & 88.07\% & 91.11\% \\
+        %Case \& definiteness & 89.65\% & 97.03\% & 88.39\% & 91.44\% \\
+        %Type \& case & 91.73\% & 96.92\% & 88.46\% & 91.51\% \\
+        %Type \& definiteness & 89.65\% & 96.99\% & 88.44\% & 91.48\% \\
+        %Type \& number & 90.02\% & 96.37\% & 87.95\% & 90.95\% \\
+        %Type, case \& definiteness & 89.61\% & 97.05\% & \textbf{88.81\%} &
+        %\textbf{91.73\%} \\
+        %Type, definiteness \& number & 88.39\% & 96.46\% & 88.06\% & 91.12\%
+        %\\
+        %\bottomrule
+    %\end{tabular}
+    %\caption{Results of experiments with modified PoS tags for nouns.}
+    %\label{substresults}
+%\end{table}
 
@@
 
+%\begin{table}
+    %\centering
+    %\smaller[0.5]
+    %\begin{tabular}{@{}lllll@{}}
+        %\toprule
+        %\textbf{Feature(s)} & \textbf{MFT} & \textbf{Accuracy} &
+        %\textbf{LAS} & \textbf{UAS} \\
+        %\midrule
+        %--- & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
+        %Mood & 94.12\% & 97.43\% & 87.04\% & 90.19\% \\
+        %Tense & 93.74\% & 97.30\% & 86.97\% & 90.18\% \\
+        %Voice & 94.13\% & 97.45\% & 86.96\% & 90.09\% \\
+        %Mood \& tense & 93.74\% & 97.31\% & 87.12\% & 90.31\% \\
+        %Voice \& tense & 93.74\% & 97.28\% & 86.99\% & 90.15\% \\
+        %Mood, tense \& voice & 93.74\% & 97.27\% & 86.83\% & 90.05\% \\
+        %Finiteness & 93.72\% & 97.35\% & \textbf{87.30\%} &
+        %\textbf{90.43\%} \\
+        %\bottomrule
+    %\end{tabular}
+    %\caption{Results of experiments with modified PoS tags for verbs.}
+    %\label{verbresults}
+%\end{table}
+
 \paragraph{Adjectives}
@@
 
+All features except for number (LAS of 86.99\%) led to increases in parser
+accuracy scores. Degree being the most promising with a reported LAS of
+87.29\%, while distinguishing adjectives on definiteness yield an LAS of
+87.14\% and gender leads to LAS of 87.10\%.
+
+Combining the most promising features, we found that none of the combinations
+surpass the LAS with the distinction of degree alone. Definiteness and degree
+with an LAS of 87.23\% and definiteness and number with an LAS of 87.27\% and
+UAS identical to that of degree alone. The most fine-grained distinction,
+definiteness, degree and number, yields an LAS of 87.14\%.
+
+%\begin{table}
+    %\centering
+    %\smaller[0.5]
+    %\begin{tabular}{@{}lllll@{}}
+        %\toprule
+        %\textbf{Feature(s)} & \textbf{MFT} & \textbf{Accuracy} &
+        %\textbf{LAS} & \textbf{UAS} \\
+        %\midrule
+        %--- & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
+        %Definiteness & 93.45\% & 96.84\% & 87.14\% & 90.29\% \\
+        %Degree & 94.13\% & 97.41\% & \textbf{87.29\%} & \textbf{90.44\%} \\
+        %Gender & 93.56\% & 96.89\% & 87.10\% & 90.25\% \\
+        %Number & 93.51\% & 96.71\% & 86.99\% & 90.10\% \\
+        %Type & 94.12\% & 97.40\% & 87.11\% & 90.25\% \\
+        %Definiteness \& degree & 93.45\% & 96.81\% & 87.23\% & 90.39\% \\
+        %Definiteness \& gender & 92.94\% & 96.31\% & 87.18\% & 90.39\% \\
+        %Definiteness \& number & 93.48\% & 96.78\% & 87.27\% &
+        %\textbf{90.44\%} \\
+        %Degree \& gender & 93.56\% & 96.87\% & 87.00\% & 90.16\% \\
+        %Degree \& number & 93.49\% & 96.76\% & 87.13\% & 90.26\% \\
+        %Definiteness, degree \& number & 93.47\% & 96.81\% & 87.14\% & 90.30\% \\
+        %\bottomrule
+    %\end{tabular}
+    %\caption{Results of experiments with modified PoS tags for adjectives.}
+    %\label{adjresults}
+%\end{table}
+
 \paragraph{Determiners}
@@
 treated/analyzed as determiners in NDT.
+
+\begin{table}
+    \centering
+    \smaller[0.5]
+    \begin{tabular}{@{}lllll@{}}
+        \toprule
+        \textbf{Feature} & \textbf{MFT} & \textbf{Accuracy} &
+        \textbf{LAS} & \textbf{UAS} \\
+        \midrule
+        --- & \textbf{94.14\%} & 97.47\% & 87.01\% & 90.19\% \\
+        Definiteness & 94.13\% & 97.49\% & \textbf{87.30\%} &
+        \textbf{90.42\%} \\
+        Gender & 93.86\% & 97.28\% & 87.09\% & 90.31\% \\
+        Number & 94.06\% & 97.49\% & 87.04\% & 90.18\% \\
+        Type & 94.14\% & \textbf{97.61\%} & 87.00\% & 90.11\% \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of experiments with modified PoS tags for determiners.}
+    \label{detresults}
+\end{table}
 
@@
         %\toprule
-        %\textbf{Tag Set} & \textbf{MFT} & \textbf{Accuracy} &
+        %\textbf{Feature(s)} & \textbf{MFT} & \textbf{Accuracy} &
         %\textbf{LAS} & \textbf{UAS} \\
         %\midrule
-        %Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
-        %Full & 85.15\% & 93.48\% & \textbf{87.15\%} & \textbf{90.39\%} \\
+        %--- & \textbf{94.14\%} & 97.47\% & 87.01\% & 90.19\% \\
+        %Case & 94.12\% & 97.50\% & 87.08\% & 90.21\% \\
+        %Gender & 94.13\% & 97.48\% & 87.06\% & 90.23\% \\
+        %Number & 94.13\% & 97.49\% & 87.21\% & 90.33\% \\
+        %Person & 94.14\% & 97.49\% & 87.22\% & 90.32\% \\
+        %Type & 94.14\% & 97.48\% & 87.19\% & 90.40\% \\
+        %Number \& person & 94.13\% & 97.49\% & 96.98\% & 90.16\% \\
+        %Type \& case & 94.12\% & 97.51\% & \textbf{87.30\%} &
+        %\textbf{90.41\%} \\
+        %Type \& number & 94.13\% & 97.49\% & 87.27\% & \textbf{90.41\%} \\
+        %Type \& person & 94.14\% & 97.49\% & 87.00\% & 90.14\% \\
+        %Type, case \& number & 94.12\% & \textbf{97.52\%} & 87.11\% & 90.36\% \\
         %\bottomrule
     %\end{tabular}
-    %\caption{Evaluation of tagging and parsing the development data with the two
-        %initial tag sets.}
-    %\label{inittagseteval}
+    %\caption{Results of experiments with modified PoS tags for pronouns.}
+    %\label{pronresults}
 %\end{table}

### 2016-05-04 submission.tex
Added results from parsing with held-out

@@
 \usepackage{relsize}
-
+\usepackage{multirow}
 %\aclfinalcopy % Uncomment this line for the final submission
@@
 \label{sec:intro}
-Part-of-speech (PoS) tagging is an important preprocessing step for many NLP tasks,
-such as dependency parsing \cite{Niv:Hal:Kub:07,Haj:Cia:Joh:09}, named
+Part-of-speech (PoS) tagging is an important preprocessing step for many NLP
+tasks, such as dependency parsing \cite{Niv:Hal:Kub:07,Haj:Cia:Joh:09}, named
 entity recognition \cite{Tjo:DeM:03} and sentiment analysis
-\cite{Wil:Wie:Hof:09}. Whereas much effort has gone into the
-development of PoS-taggers -- to the effect that this task is often
-considered more or less a solved task -- considerably less effort has
-been devoted to the empirical evaluation of the PoS tag sets
-themselves. Error analysis of PoS-taggers indicate that, whereas
-tagging improvement through means of learning algorithm or feature
-engineering seems to have reached something of a plateau, linguistic
-and empirical assessment of the distinctions made in the PoS tag set may be a venue
-worth investigating further \cite{Man:11}.  Clearly, the utility of a
-PoS tag set is tightly coupled with the the downstream task for which
-it is performed. Even so, PoS tag sets are usually employed in a ``one
-size fits all'' fashion, regardless of the requirements posed by the
-task which makes use of this information.
+\cite{Wil:Wie:Hof:09}. Whereas much effort has gone into the development of
+PoS-taggers -- to the effect that this task is often considered more or less a
+solved task -- considerably less effort has been devoted to the empirical
+evaluation of the PoS tag sets themselves. Error analysis of PoS taggers
+indicate that, whereas tagging improvement through means of learning algorithm
+or feature engineering seems to have reached something of a plateau, linguistic
+and empirical assessment of the distinctions made in the PoS tag sets may be a
+venue worth investigating further \cite{Man:11}. Clearly, the utility of a PoS
+tag set is tightly coupled with the downstream task for which it is performed.
+Even so, PoS tag sets are usually employed in a ``one size fits all'' fashion,
+regardless of the requirements posed by the task which makes use of this
+information.
 
-It is well known that syntactic parsing often benefits from quite
-fine-grained morphological distinctions
-\cite{Zha:Niv:11,See:Kuh:13}. Morphology interacts with syntax though
-phenomena such as agreement and case marking, and incorporating
-information on morphological properties of words can therefore often
-improved parsing performance. However, in a realistic setting where
-the aim is to automatically parse raw text, the generation of
-morphological information will often require a separate step of
-morphological analysis that can be quite costly.
+It is well known that syntactic parsing often benefits from quite fine-grained
+morphological distinctions \cite{Zha:Niv:11,See:Kuh:13}. Morphology interacts
+with syntax though phenomena such as agreement and case marking, and
+incorporating information on morphological properties of words can therefore
+often improved parsing performance. However, in a realistic setting where the
+aim is to automatically parse raw text, the generation of morphological
+information will often require a separate step of morphological analysis that
+can be quite costly.
 
-In this paper, we optimize a PoS tag set for the task of
-data-driven dependency parsing of Norwegian. We report on a set of experiments where various morphological
-distinctions are introduced to the PoS annotations and evaluated both intrinsically, i.e. in
-terms of PoS-tagging accuracy, and extrinsically, in terms of parsing
-accuracy.  Our results show that the introduction of morphological
-distinctions not present in the original tag set, whilst compromising
-tagger accuracy, actually leads to significantly improved parsing
+In this paper, we optimize a PoS tag set for the task of data-driven dependency
+parsing of Norwegian. We report on a set of experiments where various
+morphological distinctions are introduced to the PoS annotations and evaluated
+both intrinsically, i.e., in terms of PoS tagging accuracy, and extrinsically,
+in terms of parsing accuracy. Our results show that the introduction of
+morphological distinctions not present in the original tag set, whilst
+compromising tagger accuracy, actually leads to significantly improved parsing
 accuracy. This optimization allows us to bypass the additional step of
-morphological analysis, framing the whole preprocessing problem as a
-simple tagging task.
+morphological analysis, framing the whole preprocessing problem as a simple
+tagging task.
 
-The article is structured as follows. We start out by reviewing
-previous work on tag set evaluation in Section \ref{sec:prev}, while
-Section \ref{sec:data} details the treebank that provides the initial gold
-annotations used for our experiments. Section \ref{sec:setup} describes the experimental
-setup used in our work and Section \ref{sec:optimization} goes on to
-provide the results from our tag set optimization. Finally,
-Section \ref{sec:summary} summarizes our main findings and discusses
-some avenues for future work.
+The article is structured as follows. We start out by reviewing previous work
+on tag set evaluation in Section \ref{sec:prev}, while Section \ref{sec:data}
+details the treebank that provides the initial gold annotations used for our
+experiments. Section \ref{sec:setup} describes the experimental setup used in
+our work and Section \ref{sec:optimization} goes on to provide the results from
+our tag set optimization. Finally, Section \ref{sec:summary} summarizes our
+main findings and discusses some avenues for future work.
 
-\section{Previous/Related work}
+\section{Previous Work}
 \label{sec:prev}
@@
 
-\newcite{Meg:01} and \newcite{Meg:02} trained and evaluated a range of PoS
-taggers on the Stockholm-Umeå Corpus (SUC) \cite{Gus:Har:06}, annotated with a
-tag set based on a Swedish version of PAROLE tags totaling 139 tags.
-Furthermore, they investigated the effects of tag set size on tagging by
-mapping the original tag set into smaller subsets designed for parsing. They
-argue that a tag set with complete morphological tags may not be necessary for
-all NLP applications, for instance syntactic parsing. They found that the
-smallest tag set comprising 26 tags yields the lowest tagger error rate.
-However, for some of the taggers, augmenting the tag set with more
-linguistically informative tags may actually lead to a drop in error rate. They
-argue that this shows that the size of the tag set as well as the type of
-information in the tags are crucial factors for tagger performance.
-However/unfortunately, they do not report results of parsing with the various
-PoS tag sets.
+\newcite{Meg:02} trained and evaluated a range of PoS taggers on the
+Stockholm-Umeå Corpus (SUC) \cite{Gus:Har:06}, annotated with a tag set based
+on a Swedish version of PAROLE tags totaling 139 tags.  Furthermore, they
+investigated the effects of tag set size on tagging by mapping the original tag
+set into smaller subsets designed for parsing. They argue that a tag set with
+complete morphological tags may not be necessary for all NLP applications, for
+instance syntactic parsing. They found that the smallest tag set comprising 26
+tags yields the lowest tagger error rate.  However, for some of the taggers,
+augmenting the tag set with more linguistically informative tags may actually
+lead to a drop in error rate. They argue that this shows that the size of the
+tag set as well as the type of information in the tags are crucial factors for
+tagger performance.  However/unfortunately, they do not report results of
+parsing with the various PoS tag sets.
 
@@
 
+Our approach is somewhat semi-automatic in that we use a "mix" of both manual
+annotation in our experiments. Our initial tags are gold standard, and using
+linguistic insight coupled with computational considerations, we introduce new
+features or feature combination to the tag set. These tags are run through a
+pipeline with automatically assigned tags in both training and testing, where
+the most promising modifications are used "further". Hence, the tag set
+modifications are not deterministic, nor do we try all possible combination; we
+only experiment with linguistically motivated distinctions.
+
 \paragraph{Data Set Split}
@@
 automatically assigned tags to see how they differ with respect to performance.
-The results of our experiments reveal that the combination of training and
-testing on automatic tags is superior to training on gold standard tags and
-testing on automatic tags.  Consequently, the parser was both trained and
-tested on automatically assigned tags in our experiments.
+The results of our experiments, shown in Table \ref{tagconfigeval}, reveal that
+the combination of training and testing on automatic tags is clearly superior
+to training on gold standard tags and testing on automatic tags. Consequently,
+the parser was both trained and tested on automatically assigned tags in our
+experiments.
+
+\begin{table}
+    \centering
+    \smaller[0.5]
+    \begin{tabular}{@{}llll@{}}
+        \toprule
+        \textbf{Training} & \textbf{Testing} & \textbf{LAS} & \textbf{UAS} \\
+        \midrule
+        Gold & Gold & 90.15\% & 92.51\% \\
+        Gold & Auto & 85.68\% & 88.98\% \\
+        Auto & Auto & 87.01\% & 90.19\% \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of parsing with Mate using the various tag configurations,
+        either using gold standard tags or automatically assigned from TnT.
+        \emph{Gold} denotes gold standard tags, \emph{Auto} denotes
+        automatically assigned tags from TnT.}
+    \label{tagconfigeval}
+\end{table}
 
@@
 
-In the following, we report on a set of experiments which evaluate
-increasingly fine-grained PoS tag sets. The PoS tag sets are motivated
-by morphosyntactic properties of Norwegian and are evaluated both in
-terms of tagger and parser accuracy.
+In the following, we report on a set of experiments which evaluate increasingly
+fine-grained PoS tag sets. The PoS tag sets are motivated by morphosyntactic
+properties of Norwegian and are evaluated in terms of both tagger and parser
+accuracy.
 
-\section{Tag Set Optimization/Experiments}
+\section{Tag Set Optimization}
 \label{sec:optimization}
@@
 investigate how the tag set modifications may affect the interplay between
-tagging and parsing. As we want to investigate how we can improve the
+tagging and parsing.
+
+As we want to investigate how we can improve the
 linguistic quality of a tag set, we will not introduce tag set modifications
@@
 
-\subsection{Initial/Baseline Experiments}
+\subsection{Baseline Experiments}
 In an initial round of experiments, we concatenated the tag of each token with
@@
 new, more fine-grained tag set (hereafter referred to as the \emph{full} tag
-set). The result of this was a total of 368 tags, which is clearly very
-fine-grained. The two initial tag sets, i.e., the original tag set comprising
-19 tags and the full tag set comprising 368 tags, thus represent two extremes
-in terms of granularity.  To evaluate these tag sets and investigate how the tag
-set granularity would affect the performance of tagging and parsing, we trained
-and evaluated TnT and Mate on the training and development data, respectively,
-on the two tag sets. In Table \ref{inittagseteval}, we report the results of
-these experiments. We see that the tagger accuracy drastically drops when going
-from the original to the full tag set. TnT reports an accuracy of 97.47\% on
-the original tag set, which is reduced to 93.48\% for the full tag set. These
+set). This resulted in a total of 368 tags, which is clearly very fine-grained.
+The two initial tag sets, i.e., the original tag set comprising 19 tags and the
+full tag set comprising 368 tags, thus represent two extremes in terms of
+granularity.  To evaluate these tag sets and investigate how the tag set
+granularity would affect the performance of tagging and parsing, we trained and
+evaluated TnT and Mate on the training and development data, respectively, on
+the two tag sets. In Table \ref{inittagseteval}, we report the results of these
+experiments. We see that the tagger accuracy drastically drops when going from
+the original to the full tag set. TnT reports an accuracy of 97.47\% on the
+original tag set, which is reduced to 93.48\% for the full tag set. These
 results confirm our hypothesis that the very high linguistic quality in the
@@
 "Nearly" all tag set modifications leads to a drop in tagger accuracy, as
-expected. Few exceptions...
+expected. Few exceptions... We observed increases in parser accuracy scores
+with linguistically motivated tag set modifications for all PoS tags we
+experimented with, the most improved being noun with an increase in LAS from
+the original tag set by 1.80 percentage points.
 
@@
 
-\begin{table}
-    \centering
-    \smaller[0.5]
-    \begin{tabular}{@{}lllll@{}}
-        \toprule
-        \textbf{Feature} & \textbf{MFT} & \textbf{Accuracy} &
-        \textbf{LAS} & \textbf{UAS} \\
-        \midrule
-        --- & \textbf{94.14\%} & 97.47\% & 87.01\% & 90.19\% \\
-        Definiteness & 94.13\% & 97.49\% & \textbf{87.30\%} &
-        \textbf{90.42\%} \\
-        Gender & 93.86\% & 97.28\% & 87.09\% & 90.31\% \\
-        Number & 94.06\% & 97.49\% & 87.04\% & 90.18\% \\
-        Type & 94.14\% & \textbf{97.61\%} & 87.00\% & 90.11\% \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results of experiments with modified PoS tags for determiners.}
-    \label{detresults}
-\end{table}
+For a number of reasons, we do not report results from combinations of features
+for determiners, as the features could not be combined in any meaningful way.
+For instance, determiners in plural never have marked definiteness, hence this
+combination would only apply to determiners in singular and roughly correspond
+to the distinction on definiteness alone. Moreover, determiners in plural are
+not distinguished in terms of gender (thus ruling out the combination of number
+and gender), neither are definite determiners. Another aspect taken into
+consideration is that the most promising distinction, definiteness, applies to
+such a small number of tokens that more fine-grained distinctions would be
+overly sparse.
+
+The results from the experiments with determiners are shown in Table
+\ref{detresults}.  Introducing the type led to an increase in tagger accuracy
+by 0.14 percentage points to 97.61\%, while marginally impacting the parsing,
+with LAS of 87.00\%, 0.01 percentage points below the baseline, and UAS of
+90.11\%, 0.08 percentage points below the baseline.
+%The fact that the baseline
+%remains the same is especially worth noting, as increased tag set granularity
+%almost inevitably leads to a drop in the MFT baseline.
+The increase in tagger
+accuracy when introducing the distinction of type is noteworthy, as we expected
+the finer granularity to lead to a decrease in accuracy. This serves to
+indicate that more fine-grained distinctions for determiners, which is a quite
+disparate category in the treebank, may be quite useful for tagging. However,
+as it has negative impact on the syntactic parsing, we can conclude that the
+type of a determiner does not assist in generalizing syntactic patterns, as
+most determiners, regardless of type, appear in the same syntactic
+constructions (i.e., before a noun or noun phrase).
+
+Gender, on the other hand, improved the parsing, but complicated the tagging,
+as the various genders are often difficult to differentiate, especially so in
+the case of masculine and feminine, which share many of the same determiners.
+The number of a determiner, i.e., singular or plural, led to a small increase
+in tagger accuracy and LAS, while marginally lower UAS, 90.18\%, 0.01
+percentage points lower than that of the original tag set. Almost all
+determiners have number, and the introduction of this distinction led to small
+increases in tagger accuracy and LAS, but marginally lower UAS. The
+introduction of definiteness to the determiners led to the best parsing
+results, LAS of 87.30\% and UAS of 90.42\%, while also increasing the tagger
+accuracy slightly. The increase in LAS and UAS is rather interesting, as there
+are only 121 "determiner tokens" with marked definiteness in the development
+data. As this accounts for a very small number of tokens, coupled with the
+previously noted considerations, we did not consider further fine-grained
+modifications with definiteness. This goes to show that tokens with overt
+definiteness have noticeable impact on the syntactic parsing, and that
+distinguishing on definiteness is very beneficial.
+
+%\begin{table}
+    %\centering
+    %\smaller[0.5]
+    %\begin{tabular}{@{}lllll@{}}
+        %\toprule
+        %\textbf{Feature} & \textbf{MFT} & \textbf{Accuracy} &
+        %\textbf{LAS} & \textbf{UAS} \\
+        %\midrule
+        %--- & \textbf{94.14\%} & 97.47\% & 87.01\% & 90.19\% \\
+        %Definiteness & 94.13\% & 97.49\% & \textbf{87.30\%} &
+        %\textbf{90.42\%} \\
+        %Gender & 93.86\% & 97.28\% & 87.09\% & 90.31\% \\
+        %Number & 94.06\% & 97.49\% & 87.04\% & 90.18\% \\
+        %Type & 94.14\% & \textbf{97.61\%} & 87.00\% & 90.11\% \\
+        %\bottomrule
+    %\end{tabular}
+    %\caption{Results of experiments with modified PoS tags for determiners.}
+    %\label{detresults}
+%\end{table}
 
@@
 have case (accusative or nominative).
+
+The results in Table \ref{pronresults} show that number, person and type are
+the most informative features for parsing, with LAS of 87.21\%, 87.22\% and
+87.19\%, respectively. However, when combining number and person, we observe a
+drop by more than 0.2 percentage points, indicating that these features do not
+interact in any syntactically distinctive way. The most interesting observation
+is that all experiments exceed the baseline tagger accuracy, the most improved
+being the most fine-grained distinction, namely type, case and number combined,
+obtaining a tagger accuracy of 97.52\%. This shows that the introduction of
+more fine-grained distinctions for pronouns is beneficial and aids the PoS
+tagger in disambiguating ambiguous words. While case alone yields an LAS of
+87.08\%, we found that the combination of type and case, which is the most
+successful experiment in terms of parser performance, yields the second highest
+tagging accuracy of 97.51\%. The reason for this is that pronouns of different
+type and personal pronouns of different case exhibit quite different properties
+and appear in different constructions. Pronouns in nominative case (i.e.,
+subjects) primarily occur before the main verb, while pronouns in accusative
+case (i.e., objects) occur after the main verb, as Norwegian exhibits so-called
+V2 word order, requiring that the finite verb of a declarative clause appears
+in the second position, hence its name. The combination of type and number
+comes in close to the performance of type and case, with LAS of 87.27\% and UAS
+identical to that of type and case.
 
@@
 
+\subsection{Optimized Tag Set}
+The most successful tag set modifications and their results are seen in Table
+\ref{respectiveresults}.
+
 \begin{table*}
@@
 \end{table*}
+
+An overview of the tags comprising the final, optimized tag set can be seen in
+Table \ref{optimizedtagset}.
 
@@
 
+Results from parsing with the optimized tag are shown in Table
+\ref{finalresults}, compared to the results obtained with the initial tag sets.
+
 \begin{table}
@@
     \smaller[0.5]
-    \begin{tabular}{@{}lllll@{}}
+    \begin{tabular}{@{}llllll@{}}
         \toprule
-        \textbf{Tag set} & \textbf{MFT} & \textbf{Accuracy} &
-        \textbf{LAS} & \textbf{UAS} \\
+        \textbf{Data Set} & \textbf{Tag Set} & \textbf{MFT} & \textbf{Accuracy}
+        & \textbf{LAS} & \textbf{UAS} \\
         \midrule
-        Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
-        Full & 85.15\% & 93.48\% & 87.15\% & 90.39\% \\
-        Optimized & 89.20\% & 96.85\% & \textbf{88.87\%} & \textbf{91.78\%} \\
+        \multirow{2}{*}{Dev}
+        & Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
+        %Full & 85.15\% & 93.48\% & 87.15\% & 90.39\% \\
+        & Optimized & 89.20\% & 96.85\% & \textbf{88.87\%} & \textbf{91.78\%} \\
+        \midrule
+        \multirow{2}{*}{Test}
+        & Original & 94.22\% & 97.38\% & 86.64\% & 90.07\% \\
+        %& Full & 84.27\% & 92.79\% & 87.03\% & 90.16\% \\
+        & Optimized & 88.08\% & 96.35\% & 88.55\% & 91.41\% \\
         \bottomrule
@@
     \caption{Results of tagging and parsing with the optimized tag set,
-        compared to the initial tag sets.}
+        compared to the initial tag sets. Trained and tested using
+        automatically assigned tags from TnT.}
     \label{finalresults}
@@
 
-\section{Summary/Conclusion and Future Work}
+\section{Summary and Future Work}
 \label{sec:summary}
+The improvements in parser performance with our optimized tag set indicate that
+a more fine-grained PoS tag set may assist syntactic parsers in recognizing and
+generalizing syntactic patterns, while potentially compromising the performance
+of PoS taggers.
+
+There are several aspects of this thesis that can be further explored in future
+work, including extrinsic evaluation of the effects of PoS tag sets on other
+downstream NLP applications besides parsing, such as sentiment analysis and
+named entity recognition. These applications often require tagged data, but are
+markedly different from syntactic parsing, hence the evaluation would involve
+investigating an entirely different aspect of the effects of tag set
+granularity.
 

### 2016-05-05 submission.tex
Fixes to experimental setup

@@
 Treebank \cite{Mar:San:Mar:93}. They mapped the original tag set of Penn
-Treebank to more fine-grained tag sets using linguistic insight to investigate
-whether additional linguistic information included in finer-grained tags could
-assist the tagger. Experimenting with both lexically and syntactically
-conditioned modifications such as distinguishing between count nouns and
-noncount nouns and between transitive and intransitive verbs, they found that
-more fine-grained tag sets rarely led to improvements in tagger accuracy; the
-most successful modification yielded an improvement in tagger accuracy of 0.05
-percentage points.
+Treebank comprising 45 tags to more fine-grained tag sets using linguistic
+insight to investigate whether additional linguistic information included in
+finer-grained tags could assist the tagger. Experimenting with both lexically
+and syntactically conditioned modifications such as distinguishing between
+count nouns and noncount nouns and between transitive and intransitive verbs,
+they found that more fine-grained tag sets rarely led to improvements in tagger
+accuracy; the most successful modification yielded an improvement in tagger
+accuracy of 0.05 percentage points.  They did not find any statistically
+significant improvements using linguistically informed distinctions, arguing
+that their results do not support the hypothesis that it is possible to achieve
+significant performance improvements in PoS tagging by utilizing a
+finer-grained tag set.
 
@@
 of \newcite{Faa:Lie:Van:97}. The tag set consists of 12 morphosyntactic PoS
-tags, with 7 additional tags for punctuation and symbols. The tag set is thus
-rather coarse-grained, with broad categories such as \texttt{subst} (noun) and
-\texttt{verb} (verb). The PoS tags are complemented by a large set of
-morphological features, providing information about morphological properties such as
-definiteness, number and tense. These features are used in our tag set
-modifications, where the coarse PoS tag of relevant tokens is concatenated with
-one or more of these features to include more linguistic information in the
-tags.
+tags outlined in Table \ref{ndttagset}, with 7 additional tags for punctuation
+and symbols. The tag set is thus rather coarse-grained, with broad categories
+such as \texttt{subst} (noun) and \texttt{verb} (verb). The PoS tags are
+complemented by a large set of morphological features, providing information
+about morphological properties such as definiteness, number and tense. These
+features are used in our tag set modifications, where the coarse PoS tag of
+relevant tokens is concatenated with one or more of these features to include
+more linguistic information in the tags.
 
@@
     \caption{Overview of the PoS tag set of NDT.}
-    \label{ndttagest}
+    \label{ndttagset}
 \end{table}
@@
 
-Our approach is somewhat semi-automatic in that we use a "mix" of both manual
-annotation in our experiments. Our initial tags are gold standard, and using
-linguistic insight coupled with computational considerations, we introduce new
-features or feature combination to the tag set. These tags are run through a
-pipeline with automatically assigned tags in both training and testing, where
-the most promising modifications are used "further". Hence, the tag set
-modifications are not deterministic, nor do we try all possible combination; we
-only experiment with linguistically motivated distinctions.
+Our approach is somewhat semi-automatic in that we employ a "hybrid"
+combination of manual annotation and automatic labeling in our experiments. Our
+initial tags are gold standard, and using linguistic insight coupled with
+computational considerations, we introduce new features or feature combination
+to the tag set. These tags are run through a pipeline with automatically
+assigned tags in both training and testing, where the most promising
+modifications are used "further". Hence, the tag set modifications are not
+deterministic, nor do we try all possible combination; we only experiment with
+linguistically motivated distinctions. We are in reality altering the gold
+standard tags by the addition of more linguistic information.
 
@@
 split, care has been taken to preserve contiguous texts in the various data
-sets while keeping the split balanced in terms of genre (and source). Our
+sets while keeping the split balanced in terms of genre. Our
 proposed data set split was used in the Norwegian contribution to the Universal
@@
 \paragraph{Tagger}
-For our experiments with tag set modifications, we wanted a PoS tagger both
-reasonably fast and accurate. There is often a trade-off between the two
-factors/considerations, as the most accurate taggers tend to suffer in terms of
-speed due to their complexity. However, a tagger that achieves both close to
+For our experiments with tag set modifications, we sought a PoS tagger both
+reasonably fast and accurate. There is a often a considerable trade-off between
+the two factors, as the most accurate taggers tend to suffer in terms of speed
+due to their complexity. However, a tagger that achieves both close to
 state-of-the-art accuracy as well as very high speed is TnT \cite{Bra:00}. TnT
@@
 In choosing a syntactic parser for our experiments, we considered previous work
-on dependency parsing of Norwegian, specifically that of \cite{Sol:Skj:Ovr:14}.
-They found the Mate parser \cite{Boh:10} to be the most successful parser for
-the parsing of NDT. Furthermore, recent dependency parser comparisons
-\cite{Cho:Tet:Ste:15} showed that Mate performed very well on parsing of
-English, beating a range of contemporary state-of-the-art parsers.
+on dependency parsing of Norwegian, specifically that of
+\newcite{Sol:Skj:Ovr:14}, who found the Mate parser \cite{Boh:10} to be the
+most successful parser for parsing of NDT. Furthermore, recent dependency
+parser comparisons \cite{Cho:Tet:Ste:15} showed that Mate performed very well
+on parsing of English, beating a range of contemporary state-of-the-art
+parsers. Mate was consequently selected to evaluate the tag set modifications in
+our experiments.
 
 \paragraph{Tag Set Mapping}
-In order to carry out the tag set modifications, we created a mapping that
-takes as input a set of tags and associated morphological features to be
-included in said tags. This mapping maps the relevant existing tags to new,
-more fine-grained tags including more relevant morphological features for the
-applicable tokens. This is done by concatenating the original tag with said
-features for all applicable tokens, i.e., tokens that are assigned said tag and
-contains the set of features in its morphological features.
+The tag set modifications are realized by mapping the existing tags to new,
+more fine-grained tags "based" on associated morphological features to be
+included in specified tags. The procedure maps the relevant existing tags to
+new, more fine-grained tags including more relevant morphological features for
+the applicable tokens, i.e., tokens that are assigned said tag and feature(s)
+in the gold standard data. Hence, we are in actuality altering the gold
+standard tags by the addition of more relevant linguistic information,
+producing new gold standard annotations of higher quality.
+
+%This is done by concatenating the original tag with the
+%specified features for all applicable tokens
+
+%contains the set of features in its morphological features.
 
@@
 It is common practice to compare the performance of PoS taggers to a
-pre-computed baseline for an initial point of comparison.
-For PoS tagging, a commonly used baseline is the Most Frequent Tag (MFT)
-baseline, which we use in our experiments. This involves labeling each word
-with the tag it was assigned most frequently in the training. All unknown
-words, i.e., words not seen in the training data, are assigned the tag most
-frequently assigned to words seen only once in the training. Unknown and
-infrequent words have in common that they rarely occur, and we might therefore
-expect them to have similar properties.
+pre-computed baseline for an initial point of comparison.  For PoS tagging, a
+commonly used baseline is the Most Frequent Tag (MFT) baseline, which we used
+in our experiments. This involves labeling each word with the tag it was
+assigned most frequently in the training. All unknown words, i.e., words not
+seen in the training data, are assigned the tag most frequently assigned to
+words seen only once in the training. Unknown and infrequent words have in
+common that they rarely occur, and we might therefore expect them to have
+similar properties.
 
@@
 automatically assigned tags to see how they differ with respect to performance.
-The results of our experiments, shown in Table \ref{tagconfigeval}, reveal that
+The results of these experiments, shown in Table \ref{tagconfigeval}, reveal that
 the combination of training and testing on automatic tags is clearly superior
@@
 Moreover, it is crucial that we remove these features when working with
-automatically assigned tags, as the features is still gold standard.  For
+automatically assigned tags, as the features are still gold standard. For
 instance, if a verb token is erroneously tagged as a noun, we could potentially
@@
 the addition of more linguistic information to the tags and thus a more
-fine-grained tag set will most likely lead to a drop in tagger accuracy/tagging
-quality due to the increase in complexity. The best tagging does not
-necessarily lead to the best parse, and it is therefore interesting to
-investigate how the tag set modifications may affect the interplay between
-tagging and parsing.
-
-As we want to investigate how we can improve the
-linguistic quality of a tag set, we will not introduce tag set modifications
-which are not linguistically motivated. We will only consider distinctions we
-deem linguistically sensible, even if we expect them to impair the
-computational tractability; investigating the interplay between these two
-considerations is ultimately our goal.
+fine-grained tag set will most likely lead to a drop in tagger accuracy due to
+the increase in complexity. The best tagging does not necessarily lead to the
+best parse, hence it is interesting to investigate how linguistically informed
+tag set modifications may affect the interplay between tagging and parsing.
+%As
+%we want to investigate how we can improve the linguistic quality of a tag set,
+%we will
+%only consider tag set modifications we deem linguistically sensible.
+%we will not introduce tag set modifications which are not linguistically
+%motivated. We will only consider distinctions we deem linguistically sensible.
 
@@
 full tag set comprising 368 tags, thus represent two extremes in terms of
-granularity.  To evaluate these tag sets and investigate how the tag set
+granularity. To evaluate the tag sets and investigate how the tag set
 granularity would affect the performance of tagging and parsing, we trained and
@@
 and 1.54 percentage points, respectively.
+
+The results from tagging and parsing with modifications to nouns are reported
+in Table \ref{substresults}. As nouns constitute the largest class in the
+treebank by far (as shown in Table \ref{baselinetagerror}), the effects are
+greatest for noun tokens in terms of overall change in performance. Apart from
+case, none of the tag set modifications improves the tagging. However, they all
+give rise to increases in parser accuracy scores. Genitive case marks
+possession, hence nouns marked with genitive case are quite different from
+other nouns, taking a noun phrase as complement. Distinguishing on type is
+useful and informative, as evident by the presence of separate tags for proper
+and common nouns in many tag sets, such as those of PTB, SUC and UD (described
+in Section \ref{sec:extagsets}). When introducing the distinction of type, we
+see a large increase in parser accuracy scores, with an LAS of 88.07\% and UAS
+of 91.11\%, both exceeding the baseline by more than a percentage point.
+Definiteness is the most informative feature for parsing, achieving LAS of
+88.27\% and UAS of 91.42\%.
 
@@
         \multirow{2}{*}{Test}
-        & Original & 94.22\% & 97.38\% & 86.64\% & 90.07\% \\
+        & Original & \textbf{94.22\%} & \textbf{97.38\%} & 86.64\% & 90.07\% \\
         %& Full & 84.27\% & 92.79\% & 87.03\% & 90.16\% \\
-        & Optimized & 88.08\% & 96.35\% & 88.55\% & 91.41\% \\
+        & Optimized & 88.08\% & 96.35\% & \textbf{88.55\%} & \textbf{91.41\%} \\
         \bottomrule

### 2016-05-05 submission.tex
Minor changes, fixed references

@@
     \centering
-    %\smaller[0.5]
+    \smaller[0.5]
     \begin{tabular}{@{}ll@{}}
@@
 Dependencies project \cite{Ovr:Hoh:16}. The split will be made available at a
-companion website.
+companion website. In our experiments, we train on the training data and
+evaluate on the development data, before finally evaluating our optimized tag
+set on the held-out test data set.
 
@@
 %contains the set of features in its morphological features.
+
+\paragraph{Evaluation}
+To evaluate tagging and parsing with the various tag set modifications in our
+experiments, we employed commonly used evaluation metrics. Tagging is evaluated
+in terms of accuracy (reported by the TnT-included \texttt{tnt-diff} script),
+while labeled attachment score (LAS) and unlabeled attachment score (UAS) are
+computed by the
+\texttt{eval.pl}\footnote{\url{http://ilk.uvt.nl/conll/software.html}} script
+(used in the CoNLL shared tasks) to evaluate the parsing.
 
@@
 full tag set comprising 368 tags, thus represent two extremes in terms of
-granularity. To evaluate the tag sets and investigate how the tag set
-granularity would affect the performance of tagging and parsing, we trained and
-evaluated TnT and Mate on the training and development data, respectively, on
-the two tag sets. In Table \ref{inittagseteval}, we report the results of these
-experiments. We see that the tagger accuracy drastically drops when going from
-the original to the full tag set. TnT reports an accuracy of 97.47\% on the
-original tag set, which is reduced to 93.48\% for the full tag set. These
-results confirm our hypothesis that the very high linguistic quality in the
-full, fine-grained tag set comes at the expense of drops in tagger performance.
-However, the additional linguistic information provided by the full tag set
-improves the parser performance. With the original tag set, Mate reports an LAS
-of 87.01\% and a UAS of 90.19\%, which increases to 87.15\% and 90.39\%,
-respectively, when using the full tag set. As we are looking for linguistically
-informed distinctions that improve the syntactic parsing, these results are
-promising and serve to indicate that additional morphological information
-assists the syntactic parsers, motivating the optimization of the existing PoS
-tag set.
+granularity. To evaluate the initial tag sets and investigate how the tag set
+granularity would affect the performance of tagging and parsing, we ran
+baseline experiments with the two tag sets. In Table \ref{inittagseteval}, we
+report the results of these experiments. We see that the tagger accuracy
+drastically drops when going from the original to the full tag set. TnT reports
+an accuracy of 97.47\% on the original tag set, which is reduced to 93.48\% for
+the full tag set. These results confirm our hypothesis that the very high
+linguistic quality in the full, fine-grained tag set comes at the expense of
+drops in tagger performance.  However, the additional linguistic information
+provided by the full tag set improves the parser performance. With the original
+tag set, Mate reports an LAS of 87.01\% and a UAS of 90.19\%, which increases
+to 87.15\% and 90.39\%, respectively, using the full tag set. As we are looking
+for linguistically informed distinctions that improve the syntactic parsing,
+these results are promising and serve to indicate that additional linguistic
+information assists the syntactic parsers, motivating the optimization of the
+existing PoS tag set.
 
@@
 \subsection{Tag Set Experiments}
-We modified the tags for nouns, verbs, adjectives, determiners and pronouns in
-NDT by appending selected sets of morphological features to each tag in order
-to increase the linguistic information expressed by the tags.  For each tag, we
-first experimented with each of the features in isolation before employing
-various combinations of them. We based our choices of combinations on how
-promising the features are and what we deem worth investigating in terms of
-linguistic utility, in order to see how the features might interact.
+We modified the tags for nouns (\texttt{subst}), verbs (\texttt{verb}),
+adjectives (\texttt{adj}), determiners (\texttt{det}) and pronouns
+(\texttt{pron}) in NDT by appending selected sets of morphological features to
+each tag in order to increase the linguistic information expressed by the tags.
+For each tag, we first experimented with each of the available features in
+isolation before employing various combinations of them. We based our choices
+of combinations on how promising the features are and what we deem worth
+investigating in terms of linguistic utility, in order to see how the features
+might interact.
 
-The morphological properties of the various parts-of-speech is reflected in the
-morphological features associated with the respective PoS tags. For instance,
-as nouns can take on gender, definiteness and number, additionally "being of"
-genitive case, the treebank operates with features for gender, definiteness,
-number and case with "accompanying" values. In addition to
-morphological properties such as definiteness, tense and number, all classes
-except for verbs can be distinguished "on" type, e.g., common nouns and proper
-nouns, while pronouns can be reflexive, reciprocal, personal or interrogative.
+The morphological properties of the various parts-of-speech are reflected in
+the morphological features associated with the respective PoS tags. For
+instance, as nouns in Norwegian can take on gender, definiteness and number,
+additionally marked with genitive case, the treebank operates with features for
+gender, definiteness, number and case.
+%with "accompanying" values
+In addition to morphological properties such as definiteness, tense and number,
+all classes except for verbs can be distinguished "on" type, e.g., common nouns
+and proper nouns, while pronouns can be reflexive, reciprocal, personal or
+interrogative.
 
-\subsection{Experiments}
-"Nearly" all tag set modifications leads to a drop in tagger accuracy, as
-expected. Few exceptions... We observed increases in parser accuracy scores
-with linguistically motivated tag set modifications for all PoS tags we
-experimented with, the most improved being noun with an increase in LAS from
-the original tag set by 1.80 percentage points.
+%"Nearly" all tag set modifications leads to a drop in tagger accuracy, as
+%expected. Few exceptions... We observed increases in parser accuracy scores
+%with linguistically motivated tag set modifications for all PoS tags we
+%experimented with, the most improved being noun with an increase in LAS from
+%the original tag set by 1.80 percentage points.
 
 \paragraph{Nouns}
-In Norwegian, nouns are assigned gender (feminine, masculine, neuter),
-definiteness (indefinite or definite) and number (singular or plural). There
+In Norwegian, nouns are assigned gender (feminine, masculine or neuter),
+definiteness (definite or indefinite) and number (singular or plural). There
 is agreement in gender, definiteness and number between nouns and their
 modifiers (adjectives and determiners). Additionally, NDT has a separate case
-feature to distinguishing nouns in genitive case.
+feature for distinguishing nouns in genitive case.
 
-Experimenting with nouns, we found that all tag set modifications yielded large
-increases in LAS and UAS. The most informative features are definiteness, which
-leads to an increase in LAS by 1.26 percentage points, to 88.27\%, and type,
+As nouns constitute the largest class in the treebank by far, the effects are
+greatest for noun tokens in terms of overall change in performance. The results
+of experimenting with tag set modifications for nouns are shown in Table
+\ref{substresults}, and reveal that, apart from case, none of the tag set
+modifications improves the tagging. However, they all give rise to increases in
+parser accuracy scores. Genitive case marks possession, hence nouns marked with
+genitive case are quite different from other nouns, taking a noun phrase as
+complement. Distinguishing on type is useful and informative, as evident by the
+presence of separate tags for proper and common nouns in many tag sets, such as
+those of Penn Treebank and Stockholm-Umeå corpus.
+
+The most informative features are definiteness, which
+leads to an increase in LAS by 1.26 percentage points to 88.27\%, and type,
 yielding an LAS of 88.07\%. Turning to combinations of features, we found that
@@
 experiment, resulting in LAS of 88.81\% and UAS of 91.73\%, constituting large
-increases from parsing with the original tag set, 1.80 percentage points
+increases from parsing with the original tag set, by 1.80 percentage points
 and 1.54 percentage points, respectively.
 
-The results from tagging and parsing with modifications to nouns are reported
-in Table \ref{substresults}. As nouns constitute the largest class in the
-treebank by far (as shown in Table \ref{baselinetagerror}), the effects are
-greatest for noun tokens in terms of overall change in performance. Apart from
-case, none of the tag set modifications improves the tagging. However, they all
-give rise to increases in parser accuracy scores. Genitive case marks
-possession, hence nouns marked with genitive case are quite different from
-other nouns, taking a noun phrase as complement. Distinguishing on type is
-useful and informative, as evident by the presence of separate tags for proper
-and common nouns in many tag sets, such as those of PTB, SUC and UD (described
-in Section \ref{sec:extagsets}). When introducing the distinction of type, we
-see a large increase in parser accuracy scores, with an LAS of 88.07\% and UAS
-of 91.11\%, both exceeding the baseline by more than a percentage point.
-Definiteness is the most informative feature for parsing, achieving LAS of
-88.27\% and UAS of 91.42\%.
-
-%\begin{table}
-    %\centering
-    %\smaller[0.5]
-    %\begin{tabular}{@{}lllll@{}}
-        %\toprule
-        %\textbf{Feature(s)} & \textbf{MFT} & \textbf{Accuracy} &
-        %\textbf{LAS} & \textbf{UAS} \\
-        %\midrule
-        %--- & \textbf{94.14\%} & 97.47\% & 87.01\% & 90.19\% \\
-        %Case & 93.77\% & \textbf{97.48\%} & 87.63\% & 90.72\% \\
-        %Definiteness & 89.67\% & 97.00\% & 88.27\% & 91.42\% \\
-        %Gender & 89.54\% & 96.09\% & 87.21\% & 90.36\% \\
-        %Number & 90.04\% & 96.37\% & 87.97\% & 91.00\% \\
-        %Type & 91.90\% & 96.92\% & 88.07\% & 91.11\% \\
-        %Case \& definiteness & 89.65\% & 97.03\% & 88.39\% & 91.44\% \\
-        %Type \& case & 91.73\% & 96.92\% & 88.46\% & 91.51\% \\
-        %Type \& definiteness & 89.65\% & 96.99\% & 88.44\% & 91.48\% \\
-        %Type \& number & 90.02\% & 96.37\% & 87.95\% & 90.95\% \\
-        %Type, case \& definiteness & 89.61\% & 97.05\% & \textbf{88.81\%} &
-        %\textbf{91.73\%} \\
-        %Type, definiteness \& number & 88.39\% & 96.46\% & 88.06\% & 91.12\%
-        %\\
-        %\bottomrule
-    %\end{tabular}
-    %\caption{Results of experiments with modified PoS tags for nouns.}
-    %\label{substresults}
-%\end{table}
+\begin{table}
+    \centering
+    \smaller[1]
+    \begin{tabular}{@{}llll@{}}
+        \toprule
+        \textbf{Feature(s)} & \textbf{Acc} & \textbf{LAS} & \textbf{UAS}
+        \\
+        \midrule
+        --- & 97.47\% & 87.01\% & 90.19\% \\
+        Case & \textbf{97.48\%} & 87.63\% & 90.72\% \\
+        Definiteness & 97.00\% & 88.27\% & 91.42\% \\
+        Gender & 96.09\% & 87.21\% & 90.36\% \\
+        Number & 96.37\% & 87.97\% & 91.00\% \\
+        Type & 96.92\% & 88.07\% & 91.11\% \\
+        Case \& definiteness & 97.03\% & 88.39\% & 91.44\% \\
+        Type \& case & 96.92\% & 88.46\% & 91.51\% \\
+        Type \& definiteness & 96.99\% & 88.44\% & 91.48\% \\
+        Type \& number & 96.37\% & 87.95\% & 90.95\% \\
+        Type, case \& definiteness & 97.05\% & \textbf{88.81\%} &
+        \textbf{91.73\%} \\
+        %Type, definiteness \& number & 96.46\% & 88.06\% & 91.12\% \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of experiments with modified PoS tags for nouns.}
+    \label{substresults}
+\end{table}
 
 \paragraph{Verbs}
-Verbs are inflected for tense (infinitive, present, preterite, past perfect)
-in Norwegian and can additionally take on mood (imperative, indicative) and
-voice (active or passive). Note that voice and mood have only a single value,
-\texttt{pass} (passive) and \texttt{imp} (imperative), respectively. Verbs
-which are not passive are implicitly active, and verbs which are not imperative
-are indicative.
+Verbs are inflected for tense (infinitive, present, preterite, past perfect) in
+Norwegian and additionally exhibit mood (imperative, indicative) and voice
+(active or passive). Note that both voice and mood have only a single value in
+the treebank, \texttt{pass} (passive) and \texttt{imp} (imperative),
+respectively.  Verbs which are not passive are implicitly active, and verbs
+which are not imperative are indicative.
 
-Introducing features in isolation, mood is the only feature leading to
-increases in LAS, with a reported LAS of 87.04\%. Imperative clauses are
-fundamentally different from indicative clauses, as they lack an overt subject.
-Tense yields an LAS of 86.97\% while voice yields an LAS of 86.96\%. Combining
-mood and tense resulted in an LAS of 87.12\% and UAS of 90.31\%.
+Table \ref{verbresults} presents the results from tagging and parsing with
+modified verb tags. Introducing features in isolation, mood is the only feature
+leading to an increase in LAS, with a reported LAS of 87.04\%. Imperative
+clauses are fundamentally different from indicative clauses, as they lack an
+overt subject.  Although voice is a very distinguishing property for verbs, and
+passive verbs appear in clauses very different from those with active verbs,
+introducing this distinction in the tag set leads to a drop in LAS of 0.05
+percentage points.  Distinguishing between the various tenses yields an LAS of
+86.97\%. Combining the two most promising features, mood and tense, resulted in
+an LAS of 87.12\% and UAS of 90.31\%.
 
-In an additional experiment, we mapped the verb tenses (mood, in the case of
-imperative) to finiteness.  All verbs have finiteness, hence this distinction
+In an additional experiment, we mapped the verb tenses (and mood, in the case
+of imperative) to finiteness. All verbs have finiteness, hence this distinction
 has broad coverage. This mapping is syntactically grounded as finite verbs and
-nonfinite verbs appear in completely different syntactic construction, and
+nonfinite verbs appear in completely different syntactic constructions, and
 proved to greatly improve the parsing, as we saw the overall largest parser
@@
 percentage points higher than the baseline, respectively. This coincides with
-the observations seen for Swedish in \newcite{Ovr:08}, where finiteness was found
-to be a very beneficial linguistic feature for parsing.
+the observations seen for Swedish in \newcite{Ovr:08}, where finiteness was
+found to be a very beneficial linguistic feature for parsing.
 
-%\begin{table}
-    %\centering
-    %\smaller[0.5]
-    %\begin{tabular}{@{}lllll@{}}
-        %\toprule
-        %\textbf{Feature(s)} & \textbf{MFT} & \textbf{Accuracy} &
-        %\textbf{LAS} & \textbf{UAS} \\
-        %\midrule
-        %--- & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
-        %Mood & 94.12\% & 97.43\% & 87.04\% & 90.19\% \\
-        %Tense & 93.74\% & 97.30\% & 86.97\% & 90.18\% \\
-        %Voice & 94.13\% & 97.45\% & 86.96\% & 90.09\% \\
-        %Mood \& tense & 93.74\% & 97.31\% & 87.12\% & 90.31\% \\
-        %Voice \& tense & 93.74\% & 97.28\% & 86.99\% & 90.15\% \\
-        %Mood, tense \& voice & 93.74\% & 97.27\% & 86.83\% & 90.05\% \\
-        %Finiteness & 93.72\% & 97.35\% & \textbf{87.30\%} &
-        %\textbf{90.43\%} \\
-        %\bottomrule
-    %\end{tabular}
-    %\caption{Results of experiments with modified PoS tags for verbs.}
-    %\label{verbresults}
-%\end{table}
+\begin{table}
+    \centering
+    \smaller[0.5]
+    \begin{tabular}{@{}llll@{}}
+        \toprule
+        \textbf{Feature(s)} & \textbf{Acc} & \textbf{LAS} & \textbf{UAS} \\
+        \midrule
+        --- & \textbf{97.47\%} & 87.01\% & 90.19\% \\
+        Mood & 97.43\% & 87.04\% & 90.19\% \\
+        Tense & 97.30\% & 86.97\% & 90.18\% \\
+        Voice & 97.45\% & 86.96\% & 90.09\% \\
+        Mood \& tense & 97.31\% & 87.12\% & 90.31\% \\
+        Voice \& tense & 97.28\% & 86.99\% & 90.15\% \\
+        Mood, tense \& voice & 97.27\% & 86.83\% & 90.05\% \\
+        Finiteness & 97.35\% & \textbf{87.30\%} & \textbf{90.43\%} \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of experiments with modified PoS tags for verbs.}
+    \label{verbresults}
+\end{table}
 
@@
 Adjectives agree with the noun they modify in terms of gender, number and
-definiteness in Norwegian. Adjectives are also inflected for degree, either
-positive, comparative or superlative.
+definiteness in Norwegian. Adjectives are also inflected for degree (positive,
+comparative or superlative).
 
-All features except for number (LAS of 86.99\%) led to increases in parser
-accuracy scores. Degree being the most promising with a reported LAS of
-87.29\%, while distinguishing adjectives on definiteness yield an LAS of
-87.14\% and gender leads to LAS of 87.10\%.
+Table \ref{adjresults}
+
+All features except for number led to increases in parser accuracy scores.
+Degree being the most promising with a reported LAS of 87.29\%, while
+distinguishing adjectives on definiteness yield an LAS of 87.14\% and gender
+leads to LAS of 87.10\%.
 
@@
 with an LAS of 87.23\% and definiteness and number with an LAS of 87.27\% and
-UAS identical to that of degree alone. The most fine-grained distinction,
-definiteness, degree and number, yields an LAS of 87.14\%.
+UAS identical to that of degree alone.
 
-%\begin{table}
-    %\centering
-    %\smaller[0.5]
-    %\begin{tabular}{@{}lllll@{}}
-        %\toprule
-        %\textbf{Feature(s)} & \textbf{MFT} & \textbf{Accuracy} &
-        %\textbf{LAS} & \textbf{UAS} \\
-        %\midrule
-        %--- & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
-        %Definiteness & 93.45\% & 96.84\% & 87.14\% & 90.29\% \\
-        %Degree & 94.13\% & 97.41\% & \textbf{87.29\%} & \textbf{90.44\%} \\
-        %Gender & 93.56\% & 96.89\% & 87.10\% & 90.25\% \\
-        %Number & 93.51\% & 96.71\% & 86.99\% & 90.10\% \\
-        %Type & 94.12\% & 97.40\% & 87.11\% & 90.25\% \\
-        %Definiteness \& degree & 93.45\% & 96.81\% & 87.23\% & 90.39\% \\
-        %Definiteness \& gender & 92.94\% & 96.31\% & 87.18\% & 90.39\% \\
-        %Definiteness \& number & 93.48\% & 96.78\% & 87.27\% &
-        %\textbf{90.44\%} \\
-        %Degree \& gender & 93.56\% & 96.87\% & 87.00\% & 90.16\% \\
-        %Degree \& number & 93.49\% & 96.76\% & 87.13\% & 90.26\% \\
-        %Definiteness, degree \& number & 93.47\% & 96.81\% & 87.14\% & 90.30\% \\
-        %\bottomrule
-    %\end{tabular}
-    %\caption{Results of experiments with modified PoS tags for adjectives.}
-    %\label{adjresults}
-%\end{table}
+Turning to combinations of features, definiteness and number achieve the best
+results, very close to that of degree, with 0.02 percentage points lower LAS
+and identical UAS. Adjectives agree with their head noun and determiner in
+definiteness and number, making this an expected improvement. The combination
+of definiteness and degree is also quite promising, obtaining LAS of 87.23\%
+and UAS of 90.39\%. It is interesting that none of the combinations surpass the
+experiment with degree alone, which indicates that degree does not interact
+with the other features in any syntactically significant way.
+
+\begin{table}
+    \centering
+    \smaller[0.5]
+    \begin{tabular}{@{}llll@{}}
+        \toprule
+        \textbf{Feature(s)} & \textbf{Acc} & \textbf{LAS} & \textbf{UAS}
+        \\
+        \midrule
+        --- & \textbf{97.47\%} & 87.01\% & 90.19\% \\
+        Definiteness & 96.84\% & 87.14\% & 90.29\% \\
+        Degree & 97.41\% & \textbf{87.29\%} & \textbf{90.44\%} \\
+        Gender & 96.89\% & 87.10\% & 90.25\% \\
+        Number & 96.71\% & 86.99\% & 90.10\% \\
+        Type & 97.40\% & 87.11\% & 90.25\% \\
+        Definiteness \& degree & 96.81\% & 87.23\% & 90.39\% \\
+        Definiteness \& gender & 96.31\% & 87.18\% & 90.39\% \\
+        Definiteness \& number & 96.78\% & 87.27\% & \textbf{90.44\%} \\
+        Degree \& gender & 96.87\% & 87.00\% & 90.16\% \\
+        Degree \& number & 96.76\% & 87.13\% & 90.26\% \\
+        %Definiteness, degree \& number & 96.81\% & 87.14\% & 90.30\% \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of experiments with modified PoS tags for adjectives.}
+    \label{adjresults}
+\end{table}
 
@@
 The results from the experiments with determiners are shown in Table
-\ref{detresults}.  Introducing the type led to an increase in tagger accuracy
-by 0.14 percentage points to 97.61\%, while marginally impacting the parsing,
-with LAS of 87.00\%, 0.01 percentage points below the baseline, and UAS of
-90.11\%, 0.08 percentage points below the baseline.
-%The fact that the baseline
-%remains the same is especially worth noting, as increased tag set granularity
-%almost inevitably leads to a drop in the MFT baseline.
-The increase in tagger
-accuracy when introducing the distinction of type is noteworthy, as we expected
-the finer granularity to lead to a decrease in accuracy. This serves to
-indicate that more fine-grained distinctions for determiners, which is a quite
-disparate category in the treebank, may be quite useful for tagging. However,
-as it has negative impact on the syntactic parsing, we can conclude that the
-type of a determiner does not assist in generalizing syntactic patterns, as
-most determiners, regardless of type, appear in the same syntactic
-constructions (i.e., before a noun or noun phrase).
+\ref{detresults}.  Introducing the distinction of type led to an increase in
+tagger accuracy by 0.14 percentage points to 97.61\%, while marginally
+impacting the parsing, with LAS of 87.00\%, 0.01 percentage points below that
+of the original tag set.  The increase in tagger accuracy when introducing the
+distinction of type is noteworthy, as we expected the finer granularity to lead
+to a decrease in accuracy. This serves to indicate that more fine-grained
+distinctions for determiners, which is a quite disparate category in the
+treebank, may be quite useful for tagging. However, as it has negative impact
+on the syntactic parsing, we can conclude that the type of a determiner does
+not assist in generalizing syntactic patterns, as most determiners, regardless
+of type, appear in the same syntactic constructions (i.e., before a noun or
+noun phrase).
 
@@
 in tagger accuracy and LAS, while marginally lower UAS, 90.18\%, 0.01
-percentage points lower than that of the original tag set. Almost all
-determiners have number, and the introduction of this distinction led to small
-increases in tagger accuracy and LAS, but marginally lower UAS. The
+percentage points lower than that of the original tag set. The
 introduction of definiteness to the determiners led to the best parsing
@@
 
-%\begin{table}
-    %\centering
-    %\smaller[0.5]
-    %\begin{tabular}{@{}lllll@{}}
-        %\toprule
-        %\textbf{Feature} & \textbf{MFT} & \textbf{Accuracy} &
-        %\textbf{LAS} & \textbf{UAS} \\
-        %\midrule
-        %--- & \textbf{94.14\%} & 97.47\% & 87.01\% & 90.19\% \\
-        %Definiteness & 94.13\% & 97.49\% & \textbf{87.30\%} &
-        %\textbf{90.42\%} \\
-        %Gender & 93.86\% & 97.28\% & 87.09\% & 90.31\% \\
-        %Number & 94.06\% & 97.49\% & 87.04\% & 90.18\% \\
-        %Type & 94.14\% & \textbf{97.61\%} & 87.00\% & 90.11\% \\
-        %\bottomrule
-    %\end{tabular}
-    %\caption{Results of experiments with modified PoS tags for determiners.}
-    %\label{detresults}
-%\end{table}
+\begin{table}
+    \centering
+    \smaller[0.5]
+    \begin{tabular}{@{}llll@{}}
+        \toprule
+        \textbf{Feature} & \textbf{Acc} & \textbf{LAS} & \textbf{UAS} \\
+        \midrule
+        --- & 97.47\% & 87.01\% & 90.19\% \\
+        Definiteness & 97.49\% & \textbf{87.30\%} & \textbf{90.42\%} \\
+        Gender & 97.28\% & 87.09\% & 90.31\% \\
+        Number & 97.49\% & 87.04\% & 90.18\% \\
+        Type & \textbf{97.61\%} & 87.00\% & 90.11\% \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of experiments with modified PoS tags for determiners.}
+    \label{detresults}
+\end{table}
 
@@
 
-%\begin{table}
-    %\centering
-    %\smaller[0.5]
-    %\begin{tabular}{@{}lllll@{}}
-        %\toprule
-        %\textbf{Feature(s)} & \textbf{MFT} & \textbf{Accuracy} &
-        %\textbf{LAS} & \textbf{UAS} \\
-        %\midrule
-        %--- & \textbf{94.14\%} & 97.47\% & 87.01\% & 90.19\% \\
-        %Case & 94.12\% & 97.50\% & 87.08\% & 90.21\% \\
-        %Gender & 94.13\% & 97.48\% & 87.06\% & 90.23\% \\
-        %Number & 94.13\% & 97.49\% & 87.21\% & 90.33\% \\
-        %Person & 94.14\% & 97.49\% & 87.22\% & 90.32\% \\
-        %Type & 94.14\% & 97.48\% & 87.19\% & 90.40\% \\
-        %Number \& person & 94.13\% & 97.49\% & 96.98\% & 90.16\% \\
-        %Type \& case & 94.12\% & 97.51\% & \textbf{87.30\%} &
-        %\textbf{90.41\%} \\
-        %Type \& number & 94.13\% & 97.49\% & 87.27\% & \textbf{90.41\%} \\
-        %Type \& person & 94.14\% & 97.49\% & 87.00\% & 90.14\% \\
-        %Type, case \& number & 94.12\% & \textbf{97.52\%} & 87.11\% & 90.36\% \\
-        %\bottomrule
-    %\end{tabular}
-    %\caption{Results of experiments with modified PoS tags for pronouns.}
-    %\label{pronresults}
-%\end{table}
+\begin{table}
+    \centering
+    \smaller[0.5]
+    \begin{tabular}{@{}llll@{}}
+        \toprule
+        \textbf{Feature(s)} & \textbf{Acc} & \textbf{LAS} & \textbf{UAS} \\
+        \midrule
+        --- & 97.47\% & 87.01\% & 90.19\% \\
+        Case & 97.50\% & 87.08\% & 90.21\% \\
+        Gender & 97.48\% & 87.06\% & 90.23\% \\
+        Number & 97.49\% & 87.21\% & 90.33\% \\
+        Person & 97.49\% & 87.22\% & 90.32\% \\
+        Type & 97.48\% & 87.19\% & 90.40\% \\
+        Number \& person & 97.49\% & 96.98\% & 90.16\% \\
+        Type \& case & 97.51\% & \textbf{87.30\%} & \textbf{90.41\%} \\
+        Type \& number & 97.49\% & 87.27\% & \textbf{90.41\%} \\
+        Type \& person & 97.49\% & 87.00\% & 90.14\% \\
+        Type, case \& number & \textbf{97.52\%} & 87.11\% & 90.36\% \\
+        \bottomrule
+    \end{tabular}
+    \caption{Results of experiments with modified PoS tags for pronouns.}
+    \label{pronresults}
+\end{table}
 
@@
     \centering
-    \smaller[1]
+    \smaller[0.5]
     \begin{tabular}{@{}ll@{}}
@@
 
-\begin{table}
+\begin{table*}
     \centering
@@
     \label{finalresults}
-\end{table}
+\end{table*}
 

### 2016-05-05 submission.tex
Added results to each PoS. Unminor changes to section 5 and 6, tables and all

@@
 initial tags are gold standard, and using linguistic insight coupled with
-computational considerations, we introduce new features or feature combination
-to the tag set. These tags are run through a pipeline with automatically
-assigned tags in both training and testing, where the most promising
-modifications are used "further". Hence, the tag set modifications are not
-deterministic, nor do we try all possible combination; we only experiment with
-linguistically motivated distinctions. We are in reality altering the gold
-standard tags by the addition of more linguistic information.
+computational considerations, we introduce new gold standard distinctions to
+the tag set. The data sets with these tags are run through a pipeline with
+automatically assigned tags in both training and testing, where the most
+promising modifications are used "further". Hence, the tag set modifications
+are not deterministic, nor do we try all possible combination; we only
+experiment with linguistically motivated distinctions. We are in reality
+altering the gold standard tags by the addition of more linguistic information.
+
+As we want to investigate how we can improve the linguistic quality of a tag
+set, we will only consider tag set modifications we deem linguistically
+sensible.
 
@@
 tag set modifications may affect the interplay between tagging and parsing.
-%As
+%as
 %we want to investigate how we can improve the linguistic quality of a tag set,
@@
 %we will not introduce tag set modifications which are not linguistically
-%motivated. We will only consider distinctions we deem linguistically sensible.
+%motivated. we will only consider distinctions we deem linguistically sensible.
 
@@
     \smaller[0.5]
-    \begin{tabular}{@{}lllll@{}}
+    \begin{tabular}{@{}llll@{}}
         \toprule
-        \textbf{Tag set} & \textbf{MFT} & \textbf{Accuracy} &
+        \textbf{Tag Set} & \textbf{Acc} &
         \textbf{LAS} & \textbf{UAS} \\
         \midrule
-        Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
-        Full & 85.15\% & 93.48\% & \textbf{87.15\%} & \textbf{90.39\%} \\
+        Original & \textbf{97.47\%} & 87.01\% & 90.19\% \\
+        Full & 93.48\% & \textbf{87.15\%} & \textbf{90.39\%} \\
         \bottomrule
@@
 In addition to morphological properties such as definiteness, tense and number,
-all classes except for verbs can be distinguished "on" type, e.g., common nouns
-and proper nouns, while pronouns can be reflexive, reciprocal, personal or
-interrogative.
+all classes except for verbs have a \emph{type} feature that provides
+information about the subtype, e.g., common nouns and proper nouns, while
+pronouns can be reflexive, reciprocal, personal or interrogative.
 
@@
         Type \& definiteness & 96.99\% & 88.44\% & 91.48\% \\
-        Type \& number & 96.37\% & 87.95\% & 90.95\% \\
+        %Type \& number & 96.37\% & 87.95\% & 90.95\% \\
         Type, case \& definiteness & 97.05\% & \textbf{88.81\%} &
@@
 
-Table \ref{adjresults}
+In Table \ref{adjresults}, we report the results of modifying the \texttt{pron}
+tag in NDT. All features except for number led to increases in parser accuracy
+scores, the most successful of which is degree with a reported LAS of 87.29\%,
+while distinguishing adjectives on definiteness yield an LAS of 87.14\% and
+gender leads to LAS of 87.10\%.
 
-All features except for number led to increases in parser accuracy scores.
-Degree being the most promising with a reported LAS of 87.29\%, while
-distinguishing adjectives on definiteness yield an LAS of 87.14\% and gender
-leads to LAS of 87.10\%.
-
-Combining the most promising features, we found that none of the combinations
-surpass the LAS with the distinction of degree alone. Definiteness and degree
-with an LAS of 87.23\% and definiteness and number with an LAS of 87.27\% and
-UAS identical to that of degree alone.
+%Combining the most promising features, we
+%found that none of the combinations surpass the LAS with the distinction of
+%degree alone. Definiteness and degree with an LAS of 87.23\% and definiteness
+%and number with an LAS of 87.27\% and UAS identical to that of degree alone.
 
@@
         Definiteness \& number & 96.78\% & 87.27\% & \textbf{90.44\%} \\
-        Degree \& gender & 96.87\% & 87.00\% & 90.16\% \\
-        Degree \& number & 96.76\% & 87.13\% & 90.26\% \\
+        %Degree \& gender & 96.87\% & 87.00\% & 90.16\% \\
+        %Degree \& number & 96.76\% & 87.13\% & 90.26\% \\
         %Definiteness, degree \& number & 96.81\% & 87.14\% & 90.30\% \\
@@
 
-For a number of reasons, we do not report results from combinations of features
-for determiners, as the features could not be combined in any meaningful way.
-For instance, determiners in plural never have marked definiteness, hence this
-combination would only apply to determiners in singular and roughly correspond
-to the distinction on definiteness alone. Moreover, determiners in plural are
-not distinguished in terms of gender (thus ruling out the combination of number
-and gender), neither are definite determiners. Another aspect taken into
-consideration is that the most promising distinction, definiteness, applies to
-such a small number of tokens that more fine-grained distinctions would be
-overly sparse.
+%For a number of reasons, we do not report results from combinations of features
+%for determiners, as the features could not be combined in any meaningful way.
+%For instance, determiners in plural never have marked definiteness, hence this
+%combination would only apply to determiners in singular and roughly correspond
+%to the distinction on definiteness alone. Moreover, determiners in plural are
+%not distinguished in terms of gender (thus ruling out the combination of number
+%and gender), neither are definite determiners. Another aspect taken into
+%consideration is that the most promising distinction, definiteness, applies to
+%such a small number of tokens that more fine-grained distinctions would be
+%overly sparse.
 
@@
 accuracy slightly. The increase in LAS and UAS is rather interesting, as there
-are only 121 "determiner tokens" with marked definiteness in the development
-data. As this accounts for a very small number of tokens, coupled with the
-previously noted considerations, we did not consider further fine-grained
-modifications with definiteness. This goes to show that tokens with overt
-definiteness have noticeable impact on the syntactic parsing, and that
-distinguishing on definiteness is very beneficial.
+are only 121 determiner tokens with marked definiteness in the development
+data. As this accounts for a very small number of tokens, we did not consider
+further fine-grained modifications with definiteness. This goes to show that
+tokens with overt definiteness have noticeable impact on the syntactic parsing,
+and that distinguishing on definiteness is very beneficial.
 
@@
 interact in any syntactically distinctive way. The most interesting observation
-is that all experiments exceed the baseline tagger accuracy, the most improved
-being the most fine-grained distinction, namely type, case and number combined,
-obtaining a tagger accuracy of 97.52\%. This shows that the introduction of
-more fine-grained distinctions for pronouns is beneficial and aids the PoS
-tagger in disambiguating ambiguous words. While case alone yields an LAS of
-87.08\%, we found that the combination of type and case, which is the most
-successful experiment in terms of parser performance, yields the second highest
-tagging accuracy of 97.51\%. The reason for this is that pronouns of different
-type and personal pronouns of different case exhibit quite different properties
-and appear in different constructions. Pronouns in nominative case (i.e.,
-subjects) primarily occur before the main verb, while pronouns in accusative
-case (i.e., objects) occur after the main verb, as Norwegian exhibits so-called
-V2 word order, requiring that the finite verb of a declarative clause appears
-in the second position, hence its name. The combination of type and number
-comes in close to the performance of type and case, with LAS of 87.27\% and UAS
-identical to that of type and case.
+is that all experiments exceed the tagging accuracy of the original tag set,
+the most improved being the most fine-grained distinction, namely type, case
+and number combined, obtaining a tagger accuracy of 97.52\%. This shows that
+the introduction of more fine-grained distinctions for pronouns is beneficial
+and aids the PoS tagger in disambiguating ambiguous words. While case alone
+yields an LAS of 87.08\%, we found that the combination of type and case, which
+is the most successful experiment in terms of parser performance, yields the
+second highest tagging accuracy of 97.51\%. The reason for this is that
+pronouns of different type and personal pronouns of different case exhibit
+quite different properties and appear in different constructions. Pronouns in
+nominative case (i.e., subjects) primarily occur before the main verb, while
+pronouns in accusative case (i.e., objects) occur after the main verb, as
+Norwegian exhibits so-called V2 word order, requiring that the finite verb of a
+declarative clause appears in the second position, hence its name. The
+combination of type and number comes in close to the performance of type and
+case, with LAS of 87.27\% and UAS identical to that of type and case.
 
@@
 \subsection{Optimized Tag Set}
-The most successful tag set modifications and their results are seen in Table
-\ref{respectiveresults}.
-
-\begin{table*}
+\begin{table*}[ht]
     \centering
     \smaller[0.5]
-    \begin{tabular}{@{}llllll@{}}
+    \begin{tabular}{@{}lllll@{}}
         \toprule
-        \textbf{Category} & \textbf{Feature(s)} & \textbf{MFT} &
-        \textbf{Accuracy} & \textbf{LAS} & \textbf{UAS} \\
+        \textbf{Category} & \textbf{Feature(s)} & \textbf{Acc} & \textbf{LAS} &
+        \textbf{UAS} \\
         \midrule
-        \emph{Baseline} & --- & 94.14\% & 97.47\% & 87.01\% & 90.19\% \\
-        Noun & Type, case \& definiteness & 89.61\% & 97.05\% & 88.81\% &
-        91.73\% \\
-        Verb & Finiteness & 93.72\% & 97.35\% & 87.30\% & 90.43\% \\
-        Adjective & Degree & 94.13\% & 97.41\% & 87.29\% & 90.44\% \\
-        Determiner & Definiteness & 94.13\% & 97.49\% & 87.30\% & 90.42\% \\
-        Pronoun & Type \& case & 94.12\% & 97.51\% & 87.30\% & 90.41\% \\
+        \emph{Baseline} & --- & 97.47\% & 87.01\% & 90.19\% \\
+        Noun & Type, case \& definiteness & 97.05\% & 88.81\% & 91.73\% \\
+        Verb & Finiteness & 97.35\% & 87.30\% & 90.43\% \\
+        Adjective & Degree & 97.41\% & 87.29\% & 90.44\% \\
+        Determiner & Definiteness & 97.49\% & 87.30\% & 90.42\% \\
+        Pronoun & Type \& case & 97.51\% & 87.30\% & 90.41\% \\
         \bottomrule
@@
 
-An overview of the tags comprising the final, optimized tag set can be seen in
-Table \ref{optimizedtagset}.
+The most successful tag set modification for each PoS and their results are
+seen in Table \ref{respectiveresults}. Nouns benefit by far the most from the
+introduction of more fine-grained linguistically motivated distinctions, with
+an LAS of 88.81\% and UAS of 91.73\%. We observe that the most promising tag
+set modifications for verbs, adjectives, determiners and pronouns all reach LAS
+of \textasciitilde 87.30\% and UAS of \textasciitilde 90.40\%. To investigate
+the overall effect of these tag set modifications, we tested each of the
+improvements in parser accuracy scores from baseline for statistical
+significance using Dan Bikel's randomized parsing evaluation comparator
+script\footnote{Available as \texttt{compare.pl} at
+    \url{http://ilk.uvt.nl/conll/software.html}}, as used in the CoNLL shared
+tasks. For the most successful tag set modification for each of the categories
+seen in Table \ref{respectiveresults}, the difference in LAS from the original
+tag set is statistically significant at significance level 0.05 (\emph{p}-value
+< 0.05), as are all differences in UAS, except for verbs with finiteness
+(\emph{p}-value 0.15) and pronouns with type and case (\emph{p}-value 0.06).
+
+An overview of the tags in the optimized tag set can be seen in Table
+\ref{optimizedtagset}, comprising three new tags for adjectives, two for
+determiners, six for pronouns, seven for nouns and two for verbs, totaling 20
+tags. Appending these to the original tag set comprising 19 tags, we reach a
+total of 39 tags for NDT.
 
@@
         \texttt{subst|appell} & Common noun \\
-        \texttt{subst|appell|be} & Common noun, definite \\
-        \texttt{subst|appell|be|gen} & Common noun, definite, genitive \\
-        \texttt{subst|appell|ub} & Common noun, indefinite \\
-        \texttt{subst|appell|ub|gen} & Common noun, indefinite, genitive \\
+        \texttt{subst|appell|be} & Common noun, def. \\
+        \texttt{subst|appell|be|gen} & Common noun, def., genitive \\
+        \texttt{subst|appell|ub} & Common noun, indef. \\
+        \texttt{subst|appell|ub|gen} & Common noun, indef., genitive \\
         \texttt{subst|prop} & Proper noun \\
@@
 
-Results from parsing with the optimized tag are shown in Table
-\ref{finalresults}, compared to the results obtained with the initial tag sets.
+\paragraph{Final Evaluation}
+In Table \ref{finalresults}, we show the results of parsing with the optimized
+tag set on the held-out test set, compared to the results obtained with the
+original tag set. We see significant improvements from the original tag set on
+both the development data and the held-out test data set. The improvement in
+LAS on the development data is 1.86 percentage points, while 1.91 percentage
+points on the held-out test data. These results indicate that the additional
+linguistic information in the tags of our optimized tag set greatly aids
+syntactic parsing and that optimizing an existing PoS tag set for a downstream
+application can be useful and beneficial.
 
-\begin{table*}
+\begin{table}
     \centering
     \smaller[0.5]
-    \begin{tabular}{@{}llllll@{}}
+    \begin{tabular}{@{}lllll@{}}
         \toprule
-        \textbf{Data Set} & \textbf{Tag Set} & \textbf{MFT} & \textbf{Accuracy}
+        \textbf{Data} & \textbf{Tag Set} & \textbf{Acc}
         & \textbf{LAS} & \textbf{UAS} \\
@@
         \multirow{2}{*}{Dev}
-        & Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
+        & Original & \textbf{97.47\%} & 87.01\% & 90.19\% \\
         %Full & 85.15\% & 93.48\% & 87.15\% & 90.39\% \\
-        & Optimized & 89.20\% & 96.85\% & \textbf{88.87\%} & \textbf{91.78\%} \\
+        & Optimized & 96.85\% & \textbf{88.87\%} & \textbf{91.78\%} \\
         \midrule
         \multirow{2}{*}{Test}
-        & Original & \textbf{94.22\%} & \textbf{97.38\%} & 86.64\% & 90.07\% \\
+        & Original & \textbf{97.30\%} & 86.64\% & 90.07\% \\
         %& Full & 84.27\% & 92.79\% & 87.03\% & 90.16\% \\
-        & Optimized & 88.08\% & 96.35\% & \textbf{88.55\%} & \textbf{91.41\%} \\
+        & Optimized & 96.35\% & \textbf{88.55\%} & \textbf{91.41\%} \\
         \bottomrule
@@
     \label{finalresults}
-\end{table*}
+\end{table}
 

### 2016-05-05 submission.tex
endringer i sec:prev

@@
 applications.  There has, however, been some work that evaluates the
-effects of PoS tag set granularity on PoS tagging. This includes
-investigation of the effects of PoS tag sets on tagging of Swedish
+effects of tag set granularity on PoS tagging itself. This includes
+investigation of the effects of tag sets on PoS tagging for Swedish
 \cite{Meg:01,Meg:02} and English \cite{Mac:05}.
@@
 Stockholm-Umeå Corpus (SUC) \cite{Gus:Har:06}, annotated with a tag set based
-on a Swedish version of PAROLE tags totaling 139 tags. Furthermore, they
-investigated the effects of tag set size on tagging by mapping the original tag
-set into smaller subsets designed for parsing. They argue that a tag set with
+on a Swedish version of PAROLE tags totaling 139 tags. Furthermore, the effects 
+of tag set size on tagging was investigated by mapping the original tag
+set into smaller subsets designed for parsing. \newcite{Meg:02} argues that a tag set with
 complete morphological tags may not be necessary for all NLP applications, for
-instance syntactic parsing. They found that the smallest tag set comprising 26
+instance syntactic parsing. The study found that the smallest tag set comprising 26
 tags yields the lowest tagger error rate.  However, for some of the taggers,
 augmenting the tag set with more linguistically informative tags may actually
-lead to a drop in error rate. They argue that this shows that the size of the
+lead to a drop in error rate. \newcite{Meg:02} argues that this shows that the size of the
 tag set as well as the type of information in the tags are crucial factors for
-tagger performance.  However/unfortunately, they do not report results of
-parsing with the various PoS tag sets.
+tagger performance.  Unfortunately, results for parsing with the various PoS
+tag sets are not reported.
 
 Similarly, \newcite{Mac:05} investigated the effects of PoS tag sets on tagger
-performance in English, specifically the Wall Street Journal portion of Penn
-Treebank \cite{Mar:San:Mar:93}. They mapped the original tag set of Penn
-Treebank comprising 45 tags to more fine-grained tag sets using linguistic
-insight to investigate whether additional linguistic information included in
-finer-grained tags could assist the tagger. Experimenting with both lexically
+performance in English, specifically the Wall Street Journal portion of the Penn
+Treebank (PTB) \cite{Mar:San:Mar:93}. Based on linguistic
+considerations \newcite{Mac:05} mapped the 45 tags of the original PTB tag set
+ to more fine-grained tag sets to investigate whether additional linguistic information 
+could assist the tagger. Experimenting with both lexically
 and syntactically conditioned modifications such as distinguishing between
@@
 
+% fixme:
 Transition/overgang her ...

### 2016-05-05 submission.tex
endringer i sec:data

@@
 \label{sec:data}
-We used the newly developed Norwegian Dependency Treebank (NDT)
+Our experiments our based on the newly developed Norwegian Dependency Treebank (NDT)
 \cite{Sol:Skj:Ovr:14}, the first publicly available treebank for
 Norwegian. It was developed at the National Library of Norway in
-collaboration with the University of Oslo, and contains manual
-syntactic and morphological annotation. The treebank contains data
-from both varieties of Norwegian; 311 000 tokens of Bokmål and 303 000
-tokens of Nynorsk. We will in the following only be using the Bokmål
-portion of the treebank. The annotated texts are mostly newspaper
+collaboration with the University of Oslo, and contains manually coded 
+syntactic and morphological annotation for 
+both varieties of Norwegian; 311 000 tokens of Bokmål and 303 000
+tokens of Nynorsk. This paper only report results for Bokmål, the main variant.
+The treebanked material mostly comprise newspaper
 text, but also include government reports, parliament transcripts and
-excerpts from blogs. The annotation process of the treebank was supported by
+blog excerpts. The annotation process was supported by
 the rule-based Oslo-Bergen Tagger \cite{Hag:Joh:Nok:00} and then manually
@@
 \paragraph{Morphological Annotation}
-The morphological annotation and PoS tag set of NDT follows the Oslo-Bergen
+The morphological annotation and PoS tag set of NDT is based on the same
+inventory as used by the Oslo-Bergen
 Tagger \cite{Hag:Joh:Nok:00,Sol:13}, which in turn is largely based on the work
@@
 complemented by a large set of morphological features, providing information
-about morphological properties such as definiteness, number and tense. These
+about morphological properties such as definiteness, number and tense. 
+Selected subsets of these
 features are used in our tag set modifications, where the coarse PoS tag of
-relevant tokens is concatenated with one or more of these features to include
+relevant tokens is concatenated with one or more features to include 
 more linguistic information in the tags.
@@
     \end{tabular}
-    \caption{Overview of the PoS tag set of NDT.}
+    \caption{Overview of the original PoS tag set of NDT (excluding punctuation tags).}
     \label{ndttagset}
@@
 The syntactic annotation choices in NDT are largely based on the Norwegian
-Reference Grammar \cite{Faa:Lie:Van:97}. The annotation choices are outlined in
-Table \ref{ndtannotation}, taken from/courtesy of \newcite{Sol:Skj:Ovr:14},
+Reference Grammar \cite{Faa:Lie:Van:97}. Some central annotation choices are outlined in
+Table \ref{ndtannotation}, taken from of \newcite{Sol:Skj:Ovr:14},
 providing overview of the analyses of syntactic constructions that often
 distinguish dependency treebanks, such as coordination and the treatment of
-auxiliary and main verbs. The set of dependency relations comprises 29
+auxiliary and main verbs. The annotations comprise 29
 dependency relations, including \textsc{adv} (adverbial), \textsc{subj}

### 2016-05-05 submission.tex
la til abstract

@@
 \begin{abstract}
-Abstraction ensues.
+
+This paper reports on a suite of experiments that evaluates how the linguistic
+granularity of part-of-speech tag sets impacts the performance of tagging and
+syntactic dependency parsing. Our results show that parsing accuracy can be
+significantly improved by introducing more fine-grained morphological
+information in the tag set, even when tagger accuracy goes down as a result of
+the increased complexity of the tagging task. Our taggers and parsers are
+trained and tested using the annotations of the Norwegian Dependency Treebank.
+
 \end{abstract}

### 2016-05-05 submission.tex
added mft to results

--- 2016-05-05 submission.tex
@@
 applications.  There has, however, been some work that evaluates the
-effects of tag set granularity on PoS tagging itself. This includes
-investigation of the effects of tag sets on PoS tagging for Swedish
+effects of PoS tag set granularity on PoS tagging. This includes
+investigation of the effects of PoS tag sets on tagging of Swedish
 \cite{Meg:01,Meg:02} and English \cite{Mac:05}.
@@
 Stockholm-Umeå Corpus (SUC) \cite{Gus:Har:06}, annotated with a tag set based
-on a Swedish version of PAROLE tags totaling 139 tags. Furthermore, the effects 
-of tag set size on tagging was investigated by mapping the original tag
-set into smaller subsets designed for parsing. \newcite{Meg:02} argues that a tag set with
+on a Swedish version of PAROLE tags totaling 139 tags. Furthermore, they
+investigated the effects of tag set size on tagging by mapping the original tag
+set into smaller subsets designed for parsing. They argue that a tag set with
 complete morphological tags may not be necessary for all NLP applications, for
-instance syntactic parsing. The study found that the smallest tag set comprising 26
+instance syntactic parsing. They found that the smallest tag set comprising 26
 tags yields the lowest tagger error rate.  However, for some of the taggers,
 augmenting the tag set with more linguistically informative tags may actually
-lead to a drop in error rate. \newcite{Meg:02} argues that this shows that the size of the
+lead to a drop in error rate. They argue that this shows that the size of the
 tag set as well as the type of information in the tags are crucial factors for
-tagger performance.  Unfortunately, results for parsing with the various PoS
-tag sets are not reported.
+tagger performance.  However/unfortunately, they do not report results of
+parsing with the various PoS tag sets.
 
 Similarly, \newcite{Mac:05} investigated the effects of PoS tag sets on tagger
-performance in English, specifically the Wall Street Journal portion of the Penn
-Treebank (PTB) \cite{Mar:San:Mar:93}. Based on linguistic
-considerations \newcite{Mac:05} mapped the 45 tags of the original PTB tag set
- to more fine-grained tag sets to investigate whether additional linguistic information 
-could assist the tagger. Experimenting with both lexically
+performance in English, specifically the Wall Street Journal portion of Penn
+Treebank \cite{Mar:San:Mar:93}. They mapped the original tag set of Penn
+Treebank comprising 45 tags to more fine-grained tag sets using linguistic
+insight to investigate whether additional linguistic information included in
+finer-grained tags could assist the tagger. Experimenting with both lexically
 and syntactically conditioned modifications such as distinguishing between
@@
 
-% fixme:
 Transition/overgang her ...
@@
 report the results of these experiments. We see that the tagger accuracy
-drastically drops when going from the original to the full tag set. TnT reports
-an accuracy of 97.47\% on the original tag set, which is reduced to 93.48\% for
-the full tag set. These results confirm our hypothesis that the very high
-linguistic quality in the full, fine-grained tag set comes at the expense of
-drops in tagger performance.  However, the additional linguistic information
-provided by the full tag set improves the parser performance. With the original
-tag set, Mate reports an LAS of 87.01\% and a UAS of 90.19\%, which increases
-to 87.15\% and 90.39\%, respectively, using the full tag set. As we are looking
-for linguistically informed distinctions that improve the syntactic parsing,
-these results are promising and serve to indicate that additional linguistic
+drastically drops when going from the original to the full tag set. The MFT
+baseline for the original tag set is 94.14\%, while it drops by almost 9
+percentage points to 85.15\% for the full tag set. TnT reports an accuracy of
+97.47\% on the original tag set, which is reduced to 93.48\% for the full tag
+set. These results confirm our hypothesis that the very high linguistic quality
+in the full, fine-grained tag set comes at the expense of drops in tagger
+performance.  However, the additional linguistic information provided by the
+full tag set improves the parser performance. With the original tag set, Mate
+reports an LAS of 87.01\% and a UAS of 90.19\%, which increases to 87.15\% and
+90.39\%, respectively, using the full tag set. As we are looking for
+linguistically informed distinctions that improve the syntactic parsing, these
+results are promising and serve to indicate that additional linguistic
 information assists the syntactic parsers, motivating the optimization of the
@@
     \smaller[0.5]
-    \begin{tabular}{@{}llll@{}}
+    \begin{tabular}{@{}lllll@{}}
         \toprule
-        \textbf{Tag Set} & \textbf{Acc} &
+        \textbf{Tag Set} & \textbf{MFT} & \textbf{Acc} &
         \textbf{LAS} & \textbf{UAS} \\
         \midrule
-        Original & \textbf{97.47\%} & 87.01\% & 90.19\% \\
-        Full & 93.48\% & \textbf{87.15\%} & \textbf{90.39\%} \\
+        Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
+        Full & 85.15\% & 93.48\% & \textbf{87.15\%} & \textbf{90.39\%} \\
         \bottomrule
@@
     \smaller[0.5]
-    \begin{tabular}{@{}lllll@{}}
+    \begin{tabular}{@{}llllll@{}}
         \toprule
-        \textbf{Category} & \textbf{Feature(s)} & \textbf{Acc} & \textbf{LAS} &
-        \textbf{UAS} \\
+        \textbf{Category} & \textbf{Feature(s)} & \textbf{MFT} &
+        \textbf{Acc} & \textbf{LAS} & \textbf{UAS} \\
         \midrule
-        \emph{Baseline} & --- & 97.47\% & 87.01\% & 90.19\% \\
-        Noun & Type, case \& definiteness & 97.05\% & 88.81\% & 91.73\% \\
-        Verb & Finiteness & 97.35\% & 87.30\% & 90.43\% \\
-        Adjective & Degree & 97.41\% & 87.29\% & 90.44\% \\
-        Determiner & Definiteness & 97.49\% & 87.30\% & 90.42\% \\
-        Pronoun & Type \& case & 97.51\% & 87.30\% & 90.41\% \\
+        \emph{Original} & --- & 94.14\% & 97.47\% & 87.01\% & 90.19\% \\
+        Noun & Type, case \& definiteness & 89.61\% & 97.05\% & 88.81\% &
+        91.73\% \\
+        Verb & Finiteness & 93.72\% & 97.35\% & 87.30\% & 90.43\% \\
+        Adjective & Degree & 94.13\% & 97.41\% & 87.29\% & 90.44\% \\
+        Determiner & Definiteness & 94.13\% & 97.49\% & 87.30\% & 90.42\% \\
+        Pronoun & Type \& case & 94.12\% & 97.51\% & 87.30\% & 90.41\% \\
         \bottomrule
@@
     \smaller[0.5]
-    \begin{tabular}{@{}lllll@{}}
+    \begin{tabular}{@{}llllll@{}}
         \toprule
-        \textbf{Data} & \textbf{Tag Set} & \textbf{Acc}
-        & \textbf{LAS} & \textbf{UAS} \\
+        \textbf{Data} & \textbf{Tag Set} & \textbf{MFT} & \textbf{Acc} &
+        \textbf{LAS} & \textbf{UAS} \\
         \midrule
         \multirow{2}{*}{Dev}
-        & Original & \textbf{97.47\%} & 87.01\% & 90.19\% \\
+        & Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
         %Full & 85.15\% & 93.48\% & 87.15\% & 90.39\% \\
-        & Optimized & 96.85\% & \textbf{88.87\%} & \textbf{91.78\%} \\
+        & Optimized & 85.15\% & 96.85\% & \textbf{88.87\%} & \textbf{91.78\%} \\
         \midrule
         \multirow{2}{*}{Test}
-        & Original & \textbf{97.30\%} & 86.64\% & 90.07\% \\
+        & Original & \textbf{94.22\%} & \textbf{97.30\%} & 86.64\% & 90.07\% \\
         %& Full & 84.27\% & 92.79\% & 87.03\% & 90.16\% \\
-        & Optimized & 96.35\% & \textbf{88.55\%} & \textbf{91.41\%} \\
+        & Optimized & 88.08\% & 96.35\% & \textbf{88.55\%} & \textbf{91.41\%} \\
         \bottomrule

### 2016-05-05 submission.tex
added clarifications on acc

@@
 experiments, we employed commonly used evaluation metrics. Tagging is evaluated
-in terms of accuracy (reported by the TnT-included \texttt{tnt-diff} script),
-while labeled attachment score (LAS) and unlabeled attachment score (UAS) are
-computed by the
+in terms of accuracy (reported by the TnT-included \texttt{tnt-diff} script;
+denoted \emph{Acc} in the following tables), while labeled attachment score
+(LAS) and unlabeled attachment score (UAS) are computed by the
 \texttt{eval.pl}\footnote{\url{http://ilk.uvt.nl/conll/software.html}} script
@@
     \end{tabular}
-    \caption{Results of tagging and parsing with the two initial tag sets.}
+    \caption{Results of tagging and parsing with the two initial tag sets.
+        \emph{Acc} denotes tagger accuracy.}
     \label{inittagseteval}

### 2016-05-05 submission.tex
endinger i data split og tagger i sec:setup

--- 2016-05-05 submission.tex
@@
 \paragraph{Data Set Split}
-As there was no standardized data set split of NDT due to its very recent
-development, we needed to establish a data set split
-(training/development/test). Our data set split of the treebank follows the
-standard 80-10-10 (training/development/test) split and will be distributed
-with the treebank and proposed as the new standard(?). In creating the data set
-split, care has been taken to preserve contiguous texts in the various data
-sets while keeping the split balanced in terms of genre. Our
-proposed data set split was used in the Norwegian contribution to the Universal
-Dependencies project \cite{Ovr:Hoh:16}. The split will be made available at a
-companion website. In our experiments, we train on the training data and
-evaluate on the development data, before finally evaluating our optimized tag
-set on the held-out test data set.
+As there was no existing standardized data set split of NDT due to its recent
+development, we first needed to define separate sections for training, development and
+testing. Our proposed sectioning of the treebank follows a 
+standard 80-10-10 split. In establishing the  
+split, care has been taken to preserve contiguous texts in the various 
+sections while also keeping them balanced in terms of genre.
+
+Our defined split into sections for training, development testing, and held-out
+testing will be distributed with the future releases of the treebank, but 
+details for replicating the split will also be made available at a companion website
+for the final version of the paper. 
+% Our proposed split has already been used in the Norwegian contribution to the
+% Universal Dependencies project \cite{Ovr:Hoh:16}.
 
 \paragraph{Tagger}
-For our experiments with tag set modifications, we sought a PoS tagger both
+As our experiments during development required many repeated cycles of training and
+testing for the different modified tag sets, we sought a PoS tagger that is both
 reasonably fast and accurate. There is a often a considerable trade-off between
@@
 report the results of these experiments. We see that the tagger accuracy
-drastically drops when going from the original to the full tag set. The MFT
-baseline for the original tag set is 94.14\%, while it drops by almost 9
-percentage points to 85.15\% for the full tag set. TnT reports an accuracy of
-97.47\% on the original tag set, which is reduced to 93.48\% for the full tag
-set. These results confirm our hypothesis that the very high linguistic quality
-in the full, fine-grained tag set comes at the expense of drops in tagger
-performance.  However, the additional linguistic information provided by the
-full tag set improves the parser performance. With the original tag set, Mate
-reports an LAS of 87.01\% and a UAS of 90.19\%, which increases to 87.15\% and
-90.39\%, respectively, using the full tag set. As we are looking for
-linguistically informed distinctions that improve the syntactic parsing, these
-results are promising and serve to indicate that additional linguistic
+drastically drops when going from the original to the full tag set. TnT reports
+an accuracy of 97.47\% on the original tag set, which is reduced to 93.48\% for
+the full tag set. These results confirm our hypothesis that the very high
+linguistic quality in the full, fine-grained tag set comes at the expense of
+drops in tagger performance.  However, the additional linguistic information
+provided by the full tag set improves the parser performance. With the original
+tag set, Mate reports an LAS of 87.01\% and a UAS of 90.19\%, which increases
+to 87.15\% and 90.39\%, respectively, using the full tag set. As we are looking
+for linguistically informed distinctions that improve the syntactic parsing,
+these results are promising and serve to indicate that additional linguistic
 information assists the syntactic parsers, motivating the optimization of the
@@
     \smaller[0.5]
-    \begin{tabular}{@{}lllll@{}}
+    \begin{tabular}{@{}llll@{}}
         \toprule
-        \textbf{Tag Set} & \textbf{MFT} & \textbf{Acc} &
+        \textbf{Tag Set} & \textbf{Acc} &
         \textbf{LAS} & \textbf{UAS} \\
         \midrule
-        Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
-        Full & 85.15\% & 93.48\% & \textbf{87.15\%} & \textbf{90.39\%} \\
+        Original & \textbf{97.47\%} & 87.01\% & 90.19\% \\
+        Full & 93.48\% & \textbf{87.15\%} & \textbf{90.39\%} \\
         \bottomrule
@@
     \smaller[0.5]
-    \begin{tabular}{@{}llllll@{}}
+    \begin{tabular}{@{}lllll@{}}
         \toprule
-        \textbf{Category} & \textbf{Feature(s)} & \textbf{MFT} &
-        \textbf{Acc} & \textbf{LAS} & \textbf{UAS} \\
+        \textbf{Category} & \textbf{Feature(s)} & \textbf{Acc} & \textbf{LAS} &
+        \textbf{UAS} \\
         \midrule
-        \emph{Original} & --- & 94.14\% & 97.47\% & 87.01\% & 90.19\% \\
-        Noun & Type, case \& definiteness & 89.61\% & 97.05\% & 88.81\% &
-        91.73\% \\
-        Verb & Finiteness & 93.72\% & 97.35\% & 87.30\% & 90.43\% \\
-        Adjective & Degree & 94.13\% & 97.41\% & 87.29\% & 90.44\% \\
-        Determiner & Definiteness & 94.13\% & 97.49\% & 87.30\% & 90.42\% \\
-        Pronoun & Type \& case & 94.12\% & 97.51\% & 87.30\% & 90.41\% \\
+        \emph{Baseline} & --- & 97.47\% & 87.01\% & 90.19\% \\
+        Noun & Type, case \& definiteness & 97.05\% & 88.81\% & 91.73\% \\
+        Verb & Finiteness & 97.35\% & 87.30\% & 90.43\% \\
+        Adjective & Degree & 97.41\% & 87.29\% & 90.44\% \\
+        Determiner & Definiteness & 97.49\% & 87.30\% & 90.42\% \\
+        Pronoun & Type \& case & 97.51\% & 87.30\% & 90.41\% \\
         \bottomrule
@@
     \smaller[0.5]
-    \begin{tabular}{@{}llllll@{}}
+    \begin{tabular}{@{}lllll@{}}
         \toprule
-        \textbf{Data} & \textbf{Tag Set} & \textbf{MFT} & \textbf{Acc} &
-        \textbf{LAS} & \textbf{UAS} \\
+        \textbf{Data} & \textbf{Tag Set} & \textbf{Acc}
+        & \textbf{LAS} & \textbf{UAS} \\
         \midrule
         \multirow{2}{*}{Dev}
-        & Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
+        & Original & \textbf{97.47\%} & 87.01\% & 90.19\% \\
         %Full & 85.15\% & 93.48\% & 87.15\% & 90.39\% \\
-        & Optimized & 85.15\% & 96.85\% & \textbf{88.87\%} & \textbf{91.78\%} \\
+        & Optimized & 96.85\% & \textbf{88.87\%} & \textbf{91.78\%} \\
         \midrule
         \multirow{2}{*}{Test}
-        & Original & \textbf{94.22\%} & \textbf{97.30\%} & 86.64\% & 90.07\% \\
+        & Original & \textbf{97.30\%} & 86.64\% & 90.07\% \\
         %& Full & 84.27\% & 92.79\% & 87.03\% & 90.16\% \\
-        & Optimized & 88.08\% & 96.35\% & \textbf{88.55\%} & \textbf{91.41\%} \\
+        & Optimized & 96.35\% & \textbf{88.55\%} & \textbf{91.41\%} \\
         \bottomrule

### 2016-05-05 submission.tex
endringer i abstract

@@
 significantly improved by introducing more fine-grained morphological
-information in the tag set, even when tagger accuracy goes down as a result of
-the increased complexity of the tagging task. Our taggers and parsers are
-trained and tested using the annotations of the Norwegian Dependency Treebank.
+information in the tag set, even if tagger accuracy is compromised. Our taggers
+and parsers are  trained and tested using the annotations of the Norwegian
+Dependency Treebank.
 

### 2016-05-05 submission.tex
small changes to sec:data

--- 2016-05-05 submission.tex
@@
 \label{sec:prev}
-There has been little previous work dedicated specifically to
-extrinsic evaluation of the effects of PoS tag sets on downstream
-applications.  There has, however, been some work that evaluates the
-effects of tag set granularity on PoS tagging itself. This includes
-investigation of the effects of tag sets on PoS tagging for Swedish
-\cite{Meg:01,Meg:02} and English \cite{Mac:05}.
-
-%Increases in tagger accuracy does not readily translate to
-%improved parsing.
+There has been little previous work dedicated specifically to extrinsic
+evaluation of the effects of PoS tag sets on downstream applications.  There
+has, however, been some work that evaluates the effects of tag set granularity
+on PoS tagging itself. This includes investigation of the effects of tag sets
+on PoS tagging for Swedish \cite{Meg:01,Meg:02} and English \cite{Mac:05}.
 
@@
 on a Swedish version of PAROLE tags totaling 139 tags. Furthermore, the effects
-of tag set size on tagging was investigated by mapping the original tag
-set into smaller subsets designed for parsing. \newcite{Meg:02} argues that a tag set with
-complete morphological tags may not be necessary for all NLP applications, for
-instance syntactic parsing. The study found that the smallest tag set comprising 26
-tags yields the lowest tagger error rate.  However, for some of the taggers,
-augmenting the tag set with more linguistically informative tags may actually
-lead to a drop in error rate. \newcite{Meg:02} argues that this shows that the size of the
-tag set as well as the type of information in the tags are crucial factors for
-tagger performance.  Unfortunately, results for parsing with the various PoS
-tag sets are not reported.
+of tag set size on tagging was investigated by mapping the original tag set
+into smaller subsets designed for parsing. \newcite{Meg:02} argues that a tag
+set with complete morphological tags may not be necessary for all NLP
+applications, for instance syntactic parsing. The study found that the smallest
+tag set comprising 26 tags yields the lowest tagger error rate. However, for
+some of the taggers, augmenting the tag set with more linguistically
+informative tags may actually lead to a drop in error rate. \newcite{Meg:02}
+argues that this shows that the size of the tag set as well as the type of
+information in the tags are crucial factors for tagger performance.
+Unfortunately, results for parsing with the various PoS tag sets are not
+reported.
 
 Similarly, \newcite{Mac:05} investigated the effects of PoS tag sets on tagger
-performance in English, specifically the Wall Street Journal portion of the Penn
-Treebank (PTB) \cite{Mar:San:Mar:93}. Based on linguistic
-considerations \newcite{Mac:05} mapped the 45 tags of the original PTB tag set
- to more fine-grained tag sets to investigate whether additional linguistic information
-could assist the tagger. Experimenting with both lexically
-and syntactically conditioned modifications such as distinguishing between
-count nouns and noncount nouns and between transitive and intransitive verbs,
-they found that more fine-grained tag sets rarely led to improvements in tagger
-accuracy; the most successful modification yielded an improvement in tagger
-accuracy of 0.05 percentage points.  They did not find any statistically
-significant improvements using linguistically informed distinctions, arguing
-that their results do not support the hypothesis that it is possible to achieve
+performance in English, specifically the Wall Street Journal portion of the
+Penn Treebank (PTB) \cite{Mar:San:Mar:93}. Based on linguistic considerations,
+\newcite{Mac:05} mapped the 45 tags of the original PTB tag set to more
+fine-grained tag sets to investigate whether additional linguistic information
+could assist the tagger. Experimenting with both lexically and syntactically
+conditioned modifications such as distinguishing between count nouns and
+noncount nouns and between transitive and intransitive verbs, they found that
+more fine-grained tag sets rarely led to improvements in tagger accuracy; the
+most successful modification yielded an improvement in tagger accuracy of 0.05
+percentage points.  They did not find any statistically significant
+improvements using linguistically informed distinctions, arguing that their
+results do not support the hypothesis that it is possible to achieve
 significant performance improvements in PoS tagging by utilizing a
@@
 % fixme:
-Transition/overgang her ...
+%Transition/overgang her ...
 
@@
 \label{sec:data}
-Our experiments our based on the newly developed Norwegian Dependency Treebank (NDT)
-\cite{Sol:Skj:Ovr:14}, the first publicly available treebank for
-Norwegian. It was developed at the National Library of Norway in
-collaboration with the University of Oslo, and contains manually coded
-syntactic and morphological annotation for
-both varieties of Norwegian; 311 000 tokens of Bokmål and 303 000
-tokens of Nynorsk. This paper only report results for Bokmål, the main variant.
-The treebanked material mostly comprise newspaper
-text, but also include government reports, parliament transcripts and
-blog excerpts. The annotation process was supported by
-the rule-based Oslo-Bergen Tagger \cite{Hag:Joh:Nok:00} and then manually
-corrected by annotators, adding syntactic dependency analysis to the
-morphosyntactic annotation.
+Our experiments are based on the newly developed Norwegian Dependency Treebank
+(NDT) \cite{Sol:Skj:Ovr:14}, the first publicly available treebank for
+Norwegian. It was developed at the National Library of Norway in collaboration
+with the University of Oslo, and contains manually coded syntactic and
+morphological annotation for both varieties of Norwegian; 311 000 tokens of
+Bokmål and 303 000 tokens of Nynorsk. This paper only reports results for
+Bokmål, the main variety. The treebanked material mostly comprise newspaper
+text, but also include government reports, parliament transcripts and blog
+excerpts. The annotation process was supported by the rule-based Oslo-Bergen
+Tagger \cite{Hag:Joh:Nok:00} and then manually corrected by annotators, adding
+syntactic dependency analysis to the morphosyntactic annotation.
 
 \paragraph{Morphological Annotation}
-The morphological annotation and PoS tag set of NDT is based on the same
-inventory as used by the Oslo-Bergen
-Tagger \cite{Hag:Joh:Nok:00,Sol:13}, which in turn is largely based on the work
-of \newcite{Faa:Lie:Van:97}. The tag set consists of 12 morphosyntactic PoS
-tags outlined in Table \ref{ndttagset}, with 7 additional tags for punctuation
-and symbols. The tag set is thus rather coarse-grained, with broad categories
-such as \texttt{subst} (noun) and \texttt{verb} (verb). The PoS tags are
-complemented by a large set of morphological features, providing information
-about morphological properties such as definiteness, number and tense.
-Selected subsets of these
-features are used in our tag set modifications, where the coarse PoS tag of
-relevant tokens is concatenated with one or more features to include
-more linguistic information in the tags.
-
 \begin{table}
@@
     \end{tabular}
-    \caption{Overview of the original PoS tag set of NDT (excluding punctuation tags).}
+    \caption{Overview of the original PoS tag set of NDT (excluding punctuation
+        tags).}
     \label{ndttagset}
@@
 
+The morphological annotation and PoS tag set of NDT is based on the same
+inventory as used by the Oslo-Bergen Tagger \cite{Hag:Joh:Nok:00,Sol:13}, which
+in turn is largely based on the work of \newcite{Faa:Lie:Van:97}. The tag set
+consists of 12 morphosyntactic PoS tags outlined in Table \ref{ndttagset}, with
+7 additional tags for punctuation and symbols. The tag set is thus rather
+coarse-grained, with broad categories such as \texttt{subst} (noun) and
+\texttt{verb} (verb). The PoS tags are complemented by a large set of
+morphological features, providing information about morphological properties
+such as definiteness, number and tense.  Selected subsets of these features are
+used in our tag set modifications, where the coarse PoS tag of relevant tokens
+is concatenated with one or more features to include more linguistic
+information in the tags.
+
 \paragraph{Syntactic Annotation}
-The syntactic annotation choices in NDT are largely based on the Norwegian
-Reference Grammar \cite{Faa:Lie:Van:97}. Some central annotation choices are outlined in
-Table \ref{ndtannotation}, taken from of \newcite{Sol:Skj:Ovr:14},
-providing overview of the analyses of syntactic constructions that often
-distinguish dependency treebanks, such as coordination and the treatment of
-auxiliary and main verbs. The annotations comprise 29
-dependency relations, including \textsc{adv} (adverbial), \textsc{subj}
-(subject) and \textsc{koord} (coordination).
-
 \begin{table}
@@
 \end{table}
+
+The syntactic annotation choices in NDT are largely based on the Norwegian
+Reference Grammar \cite{Faa:Lie:Van:97}. Some central annotation choices are outlined in
+Table \ref{ndtannotation}, taken from of \newcite{Sol:Skj:Ovr:14},
+providing overview of the analyses of syntactic constructions that often
+distinguish dependency treebanks, such as coordination and the treatment of
+auxiliary and main verbs. The annotations comprise 29
+dependency relations, including \textsc{adv} (adverbial), \textsc{subj}
+(subject) and \textsc{koord} (coordination).
 
@@
 \paragraph{Data Set Split}
-As there was no existing standardized data set split of NDT due to its recent
-development, we first needed to define separate sections for training, development and
-testing. Our proposed sectioning of the treebank follows a 
-standard 80-10-10 split. In establishing the  
-split, care has been taken to preserve contiguous texts in the various 
-sections while also keeping them balanced in terms of genre.
-
-Our defined split into sections for training, development testing, and held-out
-testing will be distributed with the future releases of the treebank, but 
-details for replicating the split will also be made available at a companion website
-for the final version of the paper. 
-% Our proposed split has already been used in the Norwegian contribution to the
-% Universal Dependencies project \cite{Ovr:Hoh:16}.
+As there was no standardized data set split of NDT due to its very recent
+development, we needed to establish a data set split
+(training/development/test). Our data set split of the treebank follows the
+standard 80-10-10 (training/development/test) split and will be distributed
+with the treebank and proposed as the new standard(?). In creating the data set
+split, care has been taken to preserve contiguous texts in the various data
+sets while keeping the split balanced in terms of genre. Our
+proposed data set split was used in the Norwegian contribution to the Universal
+Dependencies project \cite{Ovr:Hoh:16}. The split will be made available at a
+companion website. In our experiments, we train on the training data and
+evaluate on the development data, before finally evaluating our optimized tag
+set on the held-out test data set.
 
 \paragraph{Tagger}
-As our experiments during development required many repeated cycles of training and
-testing for the different modified tag sets, we sought a PoS tagger that is both
+For our experiments with tag set modifications, we sought a PoS tagger both
 reasonably fast and accurate. There is a often a considerable trade-off between

### 2016-05-05 submission.tex
endringer i abstract

--- 2016-05-05 submission.tex
@@
 
+% fixme: tror nok dette bør flyttes og kortes ned. kan bakes inn med det som
+% står under "Tag Set Mapping".
 Our approach is somewhat semi-automatic in that we employ a "hybrid"
@@
 on dependency parsing of Norwegian, specifically that of
-\newcite{Sol:Skj:Ovr:14}, who found the Mate parser \cite{Boh:10} to be the
-most successful parser for parsing of NDT. Furthermore, recent dependency
+\newcite{Sol:Skj:Ovr:14}, who found the Mate parser \cite{Boh:10} to have  
+the best performance for parsing of NDT. Furthermore, recent dependency
 parser comparisons \cite{Cho:Tet:Ste:15} showed that Mate performed very well
-on parsing of English, beating a range of contemporary state-of-the-art
-parsers. Mate was consequently selected to evaluate the tag set modifications in
+on parsing of English, outperforming a range of contemporary state-of-the-art
+parsers. We will be using Mate for gauging the effect of the tag set modifications in
 our experiments.
@@
 \paragraph{Baseline}
-It is common practice to compare the performance of PoS taggers to a
-pre-computed baseline for an initial point of comparison.  For PoS tagging, a
-commonly used baseline is the Most Frequent Tag (MFT) baseline, which we used
-in our experiments. This involves labeling each word with the tag it was
-assigned most frequently in the training. All unknown words, i.e., words not
+As a point of reference for the PoS tagging we will be reporting the commonly used
+most-frequent-tag baseline, henceforth MFT.
+This involves labeling each word with the tag it was
+assigned most frequently in the training data. All unknown words, i.e., words not
 seen in the training data, are assigned the tag most frequently assigned to
-words seen only once in the training. Unknown and infrequent words have in
-common that they rarely occur, and we might therefore expect them to have
-similar properties.
+words seen only once.  The rationale is that we might expect unknown
+and infrequent words to have similar properties given that they both occur rarely.
 
-\paragraph{Tags \& Features}
+\paragraph{Tags \& Features} % <-- fixme: rename? 
 As we seek to quantify the effects of PoS tagging in a realistic setting, i.e.,
-in application to raw text, we want to evaluate the parser on automatically
-assigned PoS tags. For the training of the parser, however, we have two
+in application to raw text, we choose to evaluate the parser on automatically
+assigned PoS tags. For training, however, we have two
 options: using either gold standard or automatically assigned tags. In order to
-settle on a configuration, we conducted experiments with gold standard and
-automatically assigned tags to see how they differ with respect to performance.
+settle on a configuration, we conducted experiments with training on both gold
+standard tags and
+automatically assigned tags to see how the performance differs.
 The results of these experiments, shown in Table \ref{tagconfigeval}, reveal that
@@
 to training on gold standard tags and testing on automatic tags. Consequently,
-the parser was both trained and tested on automatically assigned tags in our
-experiments.
+the parser is both trained and tested on automatically assigned tags in the
+remainder of our experiments.
 
@@
 
-We removed any morphological features in order to simulate a realistic setting.
+Though some parsers make use of morphological features, % fixme: status for mate?
+we removed all morphological features beyond the PoS tags in order to simulate a realistic setting.
 Moreover, it is crucial that we remove these features when working with
-automatically assigned tags, as the features are still gold standard. For
-instance, if a verb token is erroneously tagged as a noun, we could potentially
-have a noun token with verbal features such as tense, which markedly obfuscates
-the training and parsing. Another important aspect is that we want to isolate
-the effect of PoS tags, necessitating the exclusion of morphological features.
+automatically assigned tags, as such features would be provided by the gold
+standard.
+
+%For instance, if a verb token is erroneously tagged as a noun, we could potentially
+%have a noun token with verbal features such as tense, which markedly obfuscates
+%the training and parsing. Another important aspect is that we want to isolate
+%the effect of PoS tags, necessitating the exclusion of morphological features.
 

### 2016-05-06 submission.tex
Moving around, fixes to sec:setup
added transition from sec:prev to sec:data

@@
 \cite{Wil:Wie:Hof:09}. Whereas much effort has gone into the development of
-PoS-taggers -- to the effect that this task is often considered more or less a
+PoS taggers -- to the effect that this task is often considered more or less a
 solved task -- considerably less effort has been devoted to the empirical
@@
 
-% fixme:
-%Transition/overgang her ...
+We argue, however, that it is possible to significantly improve the performance
+of the downstream application of syntactic parsing by introducing
+linguistically informed, syntactically informative distinctions in a tag set.
+We will now turn to the data that provided the basis for our experiments and
+outline its morphological and syntactic annotation choices.
 
@@
 The syntactic annotation choices in NDT are largely based on the Norwegian
-Reference Grammar \cite{Faa:Lie:Van:97}. Some central annotation choices are outlined in
-Table \ref{ndtannotation}, taken from of \newcite{Sol:Skj:Ovr:14},
+Reference Grammar \cite{Faa:Lie:Van:97}. Some central annotation choices are
+outlined in Table \ref{ndtannotation}, taken from \newcite{Sol:Skj:Ovr:14},
 providing overview of the analyses of syntactic constructions that often
 distinguish dependency treebanks, such as coordination and the treatment of
-auxiliary and main verbs. The annotations comprise 29
-dependency relations, including \textsc{adv} (adverbial), \textsc{subj}
-(subject) and \textsc{koord} (coordination).
+auxiliary and main verbs. The annotations comprise 29 dependency relations,
+including \textsc{adv} (adverbial), \textsc{subj} (subject) and \textsc{koord}
+(coordination).
 
@@
 
-% fixme: tror nok dette bør flyttes og kortes ned. kan bakes inn med det som
-% står under "Tag Set Mapping".
-Our approach is somewhat semi-automatic in that we employ a "hybrid"
-combination of manual annotation and automatic labeling in our experiments. Our
-initial tags are gold standard, and using linguistic insight coupled with
-computational considerations, we introduce new gold standard distinctions to
-the tag set. The data sets with these tags are run through a pipeline with
-automatically assigned tags in both training and testing, where the most
-promising modifications are used "further". Hence, the tag set modifications
-are not deterministic, nor do we try all possible combination; we only
-experiment with linguistically motivated distinctions. We are in reality
-altering the gold standard tags by the addition of more linguistic information.
-
-As we want to investigate how we can improve the linguistic quality of a tag
-set, we will only consider tag set modifications we deem linguistically
-sensible.
-
 \paragraph{Data Set Split}
 As there was no existing standardized data set split of NDT due to its recent
-development, we first needed to define separate sections for training, development and
-testing. Our proposed sectioning of the treebank follows a 
-standard 80-10-10 split. In establishing the  
-split, care has been taken to preserve contiguous texts in the various 
-sections while also keeping them balanced in terms of genre.
+development, we first needed to define separate sections for training,
+development and testing. Our proposed sectioning of the treebank follows a
+standard 80-10-10 split. In establishing the split, care has been taken to
+preserve contiguous texts in the various sections while also keeping them
+balanced in terms of genre.
 
@@
 
+% fixme: tror nok dette bør flyttes og kortes ned. kan bakes inn med det som
+% står under "Tag Set Mapping".
+Our approach is somewhat semi-automatic in that we employ a "hybrid"
+combination of manual annotation and automatic labeling in our experiments. Our
+initial tags are gold standard, and using linguistic insight coupled with
+computational considerations, we introduce new gold standard distinctions to
+the tag set. The data sets with these tags are run through a pipeline with
+automatically assigned tags in both training and testing, where the most
+promising modifications are used "further". Hence, the tag set modifications
+are not deterministic, nor do we try all possible combination; we only
+experiment with linguistically motivated distinctions. We are in reality
+altering the gold standard tags by the addition of more linguistic information.
+
+As we want to investigate how we can improve the linguistic quality of a tag
+set, we will only consider tag set modifications we deem linguistically
+sensible.
 %This is done by concatenating the original tag with the
@@
 \paragraph{Baseline}
-As a point of reference for the PoS tagging we will be reporting the commonly used
-most-frequent-tag baseline, henceforth MFT.
-This involves labeling each word with the tag it was
-assigned most frequently in the training data. All unknown words, i.e., words not
-seen in the training data, are assigned the tag most frequently assigned to
-words seen only once.  The rationale is that we might expect unknown
-and infrequent words to have similar properties given that they both occur rarely.
+As a point of reference for the PoS tagging, we will be reporting the commonly
+used most-frequent-tag baseline, henceforth MFT. This involves labeling each
+word with the tag it was assigned most frequently in the training data. All
+unknown words, i.e., words not seen in the training data, are assigned the tag
+most frequently assigned to words seen only once. The rationale is that we
+might expect unknown and infrequent words to have similar properties given that
+they both occur rarely.
 
@@
 in application to raw text, we choose to evaluate the parser on automatically
-assigned PoS tags. For training, however, we have two
-options: using either gold standard or automatically assigned tags. In order to
-settle on a configuration, we conducted experiments with training on both gold
-standard tags and
-automatically assigned tags to see how the performance differs.
-The results of these experiments, shown in Table \ref{tagconfigeval}, reveal that
+assigned PoS tags. For training, however, we have two options: using either
+gold standard or automatically assigned tags. In order to settle on a
+configuration, we conducted experiments with training on both gold standard
+tags and automatically assigned tags to see how the performance differs. The
+results of these experiments, shown in Table \ref{tagconfigeval}, reveal that
 the combination of training and testing on automatic tags is clearly superior
@@
 Though some parsers make use of morphological features, % fixme: status for mate?
-we removed all morphological features beyond the PoS tags in order to simulate a realistic setting.
-Moreover, it is crucial that we remove these features when working with
-automatically assigned tags, as such features would be provided by the gold
-standard.
+we removed all morphological features beyond the PoS tags in order to simulate
+a realistic setting.  Moreover, it is crucial that we remove these features
+when working with automatically assigned tags, as such features would be
+provided by the gold standard.
 
@@
 Table \ref{verbresults} presents the results from tagging and parsing with
-modified verb tags. Introducing features in isolation, mood is the only feature
-leading to an increase in LAS, with a reported LAS of 87.04\%. Imperative
-clauses are fundamentally different from indicative clauses, as they lack an
-overt subject.  Although voice is a very distinguishing property for verbs, and
-passive verbs appear in clauses very different from those with active verbs,
-introducing this distinction in the tag set leads to a drop in LAS of 0.05
-percentage points.  Distinguishing between the various tenses yields an LAS of
-86.97\%. Combining the two most promising features, mood and tense, resulted in
-an LAS of 87.12\% and UAS of 90.31\%.
+modified verb tags. Imperative clauses are fundamentally different from
+indicative clauses, as they lack an overt subject, which is reflected in the
+fact that mood is the only feature leading to an increase in LAS, with a
+reported LAS of 87.04\% Although voice is a very distinguishing property for
+verbs, and passive verbs appear in clauses very different from those with
+active verbs, introducing this distinction in the tag set leads to a drop in
+LAS of 0.05 percentage points.  Distinguishing between the various tenses
+yields an LAS of 86.97\%. Combining the two most promising features, mood and
+tense, resulted in an LAS of 87.12\% and UAS of 90.31\%.
 
@@
 treebank, may be quite useful for tagging. However, as it has negative impact
-on the syntactic parsing, we can conclude that the type of a determiner does
-not assist in generalizing syntactic patterns, as most determiners, regardless
-of type, appear in the same syntactic constructions (i.e., before a noun or
-noun phrase).
+on parsing, we can conclude that the type of a determiner does not assist in
+generalizing syntactic patterns. Most determiners, regardless of type, appear
+in the same syntactic constructions (i.e., before a noun or noun phrase).
 

### 2016-05-06 submission.tex
added transition from sec:prev to sec:data

@@
 of the downstream application of syntactic parsing by introducing
-linguistically informed, syntactically informative distinctions in a tag set.
+linguistically informed, syntactically informative distinctions to a tag set.
 We will now turn to the data that provided the basis for our experiments and

### 2016-05-06 submission.tex
minor changes to sec:setup

@@
 on dependency parsing of Norwegian, specifically that of
-\newcite{Sol:Skj:Ovr:14}, who found the Mate parser \cite{Boh:10} to have
-the best performance for parsing of NDT. Furthermore, recent dependency
-parser comparisons \cite{Cho:Tet:Ste:15} showed that Mate performed very well
-on parsing of English, outperforming a range of contemporary state-of-the-art
-parsers. We will be using Mate for gauging the effect of the tag set modifications in
-our experiments.
+\newcite{Sol:Skj:Ovr:14}, who found the Mate parser \cite{Boh:10} to have the
+best performance for parsing of NDT. Furthermore, recent dependency parser
+comparisons \cite{Cho:Tet:Ste:15} showed that Mate performed very well on
+parsing of English, outperforming a range of contemporary state-of-the-art
+parsers. We will be using Mate for gauging the effects of the tag set
+modifications in our experiments.
 
@@
 are not deterministic, nor do we try all possible combination; we only
-experiment with linguistically motivated distinctions. We are in reality
-altering the gold standard tags by the addition of more linguistic information.
+experiment with linguistically motivated distinctions.
 
@@
 For each tag, we first experimented with each of the available features in
-isolation before employing various combinations of them. We based our choices
+isolation before employing various combinations of them. We base our choices
 of combinations on how promising the features are and what we deem worth

### 2016-05-06 submission.tex
changes to tag set experiments on nouns

@@
 \subsection{Tag Set Experiments}
-We modified the tags for nouns (\texttt{subst}), verbs (\texttt{verb}),
+We modify the tags for nouns (\texttt{subst}), verbs (\texttt{verb}),
 adjectives (\texttt{adj}), determiners (\texttt{det}) and pronouns
@@
 each tag in order to increase the linguistic information expressed by the tags.
-For each tag, we first experimented with each of the available features in
-isolation before employing various combinations of them. We base our choices
-of combinations on how promising the features are and what we deem worth
+For each tag, we first experiment with each of the available features in
+isolation before employing various combinations of them. We base our choices of
+combinations on how promising the features are and what we deem worth
 investigating in terms of linguistic utility, in order to see how the features
@@
 instance, as nouns in Norwegian can take on gender, definiteness and number,
-additionally marked with genitive case, the treebank operates with features for
-gender, definiteness, number and case.
-%with "accompanying" values
-In addition to morphological properties such as definiteness, tense and number,
-all classes except for verbs have a \emph{type} feature that provides
-information about the subtype, e.g., common nouns and proper nouns, while
-pronouns can be reflexive, reciprocal, personal or interrogative.
-
-%"Nearly" all tag set modifications leads to a drop in tagger accuracy, as
-%expected. Few exceptions... We observed increases in parser accuracy scores
-%with linguistically motivated tag set modifications for all PoS tags we
-%experimented with, the most improved being noun with an increase in LAS from
-%the original tag set by 1.80 percentage points.
+the treebank operates with features for gender, definiteness and number
+complementing the \texttt{subst} tag. In addition to morphological properties
+such as definiteness, tense and number, all classes except for verbs have a
+\emph{type} feature that provides information about the subtype of the PoS,
+e.g., whether a noun is common or proper.
 
@@
 In Norwegian, nouns are assigned gender (feminine, masculine or neuter),
-definiteness (definite or indefinite) and number (singular or plural). There
-is agreement in gender, definiteness and number between nouns and their
-modifiers (adjectives and determiners). Additionally, NDT has a separate case
-feature for distinguishing nouns in genitive case.
+definiteness (definite or indefinite) and number (singular or plural). There is
+agreement in gender, definiteness and number between nouns and their modifiers
+(adjectives and determiners). Additionally, NDT has a separate \emph{case}
+feature for distinguishing nouns in genitive case. Genitive case marks
+possession, hence nouns marked with genitive case are quite different from
+other nouns, taking a noun phrase as complement. Distinguishing on type is
+useful and informative, as evident by the presence of separate tags for proper
+and common nouns in many tag sets, such as those of Penn Treebank and
+Stockholm-Umeå corpus.
 
-As nouns constitute the largest class in the treebank by far, the effects are
-greatest for noun tokens in terms of overall change in performance. The results
-of experimenting with tag set modifications for nouns are shown in Table
-\ref{substresults}, and reveal that, apart from case, none of the tag set
-modifications improves the tagging. However, they all give rise to increases in
-parser accuracy scores. Genitive case marks possession, hence nouns marked with
-genitive case are quite different from other nouns, taking a noun phrase as
-complement. Distinguishing on type is useful and informative, as evident by the
-presence of separate tags for proper and common nouns in many tag sets, such as
-those of Penn Treebank and Stockholm-Umeå corpus.
-
+%As nouns constitute the largest class in the treebank by far, the effects of
+%the tag set modifications are greatest for noun tokens in terms of overall
+%change in performance.
+The results of experimenting with tag set modifications
+for nouns are shown in Table \ref{substresults} and reveal that, apart from
+case, none of the tag set modifications improve the tagging. However, they all
+give rise to increases in parser accuracy scores.
 The most informative features are definiteness, which
@@
 experiment, resulting in LAS of 88.81\% and UAS of 91.73\%, constituting large
-increases from parsing with the original tag set, by 1.80 percentage points
+improvements from parsing with the original tag set, of 1.80 percentage points
 and 1.54 percentage points, respectively.
@@
 \paragraph{Verbs}
-Verbs are inflected for tense (infinitive, present, preterite, past perfect) in
-Norwegian and additionally exhibit mood (imperative, indicative) and voice
-(active or passive). Note that both voice and mood have only a single value in
-the treebank, \texttt{pass} (passive) and \texttt{imp} (imperative),
+Verbs are inflected for tense (infinitive, present, preterite or past perfect)
+in Norwegian and additionally exhibit mood (imperative or indicative; also
+conjunctive, however, it is very uncommon and does not occur in the treebank)
+and voice (active or passive). Note that both voice and mood have only a single
+value in the treebank; \texttt{pass} (passive) and \texttt{imp} (imperative),
 respectively.  Verbs which are not passive are implicitly active, and verbs

### 2016-05-06 submission.tex
changes to verbs paragraph

@@
 respectively. Verbs which are not passive are implicitly active, and verbs
-which are not imperative are indicative.
+which are not imperative are in indicative mood.
 
@@
 modified verb tags. Imperative clauses are fundamentally different from
-indicative clauses, as they lack an overt subject, which is reflected in the
+indicative clauses as they lack an overt subject, which is reflected in the
 fact that mood is the only feature leading to an increase in LAS, with a
-reported LAS of 87.04\% Although voice is a very distinguishing property for
-verbs, and passive verbs appear in clauses very different from those with
-active verbs, introducing this distinction in the tag set leads to a drop in
-LAS of 0.05 percentage points.  Distinguishing between the various tenses
-yields an LAS of 86.97\%. Combining the two most promising features, mood and
-tense, resulted in an LAS of 87.12\% and UAS of 90.31\%.
+reported LAS of 87.04\%. Although voice is a very distinguishing property for
+verbs, and passive clauses are very different from active clauses, introducing
+this distinction in the tag set leads to a drop in LAS of 0.05 percentage
+points, while distinguishing between the various tenses yields an LAS of
+86.97\%. Combining the two most promising features of mood and tense resulted
+in an LAS of 87.12\% and UAS of 90.31\%.
 
@@
 nonfinite verbs appear in completely different syntactic constructions, and
-proved to greatly improve the parsing, as we saw the overall largest parser
-accuracy scores, with 87.30\% for LAS and 90.43\% for UAS, 0.29 and 0.24
-percentage points higher than the baseline, respectively. This coincides with
-the observations seen for Swedish in \newcite{Ovr:08}, where finiteness was
-found to be a very beneficial linguistic feature for parsing.
+proved to greatly improve the parsing, as we saw the highest parser accuracy
+scores of 87.30\% for LAS and 90.43\% for UAS, 0.29 and 0.24 percentage points
+higher than the baseline, respectively. This coincides with the observations
+seen for Swedish in \newcite{Ovr:08}, where finiteness was found
+to be a very beneficial linguistic feature for parsing.
 

### 2016-05-06 submission.tex
changes to paragraph on determiners in tag set experiments

@@
 Adjectives agree with the noun they modify in terms of gender, number and
-definiteness in Norwegian. Adjectives are also inflected for degree (positive,
-comparative or superlative).
+definiteness in Norwegian. Furthermore, adjectives are inflected for
+degree (positive, comparative or superlative).
 
 In Table \ref{adjresults}, we report the results of modifying the \texttt{pron}
-tag in NDT. All features except for number led to increases in parser accuracy
+tag in NDT. All features except for number lead to increases in parser accuracy
 scores, the most successful of which is degree with a reported LAS of 87.29\%,
-while distinguishing adjectives on definiteness yield an LAS of 87.14\% and
-gender leads to LAS of 87.10\%.
+while distinguishing adjectives on definiteness yields an LAS of 87.14\% and
+introducing the distinction of gender leads to LAS of 87.10\%.
 
@@
 Turning to combinations of features, definiteness and number achieve the best
-results, very close to that of degree, with 0.02 percentage points lower LAS
-and identical UAS. Adjectives agree with their head noun and determiner in
-definiteness and number, making this an expected improvement. The combination
-of definiteness and degree is also quite promising, obtaining LAS of 87.23\%
-and UAS of 90.39\%. It is interesting that none of the combinations surpass the
-experiment with degree alone, which indicates that degree does not interact
-with the other features in any syntactically significant way.
+parser accuracy scores, very close to those of degree, with 0.02 percentage
+points lower LAS and identical UAS. Adjectives agree with their head noun and
+determiner in definiteness and number, making this an expected improvement. The
+combination of definiteness and degree is also quite promising, obtaining LAS
+of 87.23\% and UAS of 90.39\%. It is interesting that none of the combinations
+surpass the experiment with degree alone, which indicates that degree does not
+interact with the other features in any syntactically significant way.
 
@@
 Like adjectives, determiners in Norwegian agree with the noun they modify in
-terms of gender, number and definiteness. Possessive pronouns are
-treated/analyzed as determiners in NDT.
+terms of gender, number and definiteness. We do not report results from
+combinations of features for determiners, as the features could not be combined
+in any meaningful way.
+%VPossessive pronouns are
+%treated/analyzed as determiners in NDT.
 
-%For a number of reasons, we do not report results from combinations of features
-%for determiners, as the features could not be combined in any meaningful way.
 %For instance, determiners in plural never have marked definiteness, hence this
@@
 The results from the experiments with determiners are shown in Table
-\ref{detresults}.  Introducing the distinction of type led to an increase in
-tagger accuracy by 0.14 percentage points to 97.61\%, while marginally
+\ref{detresults}. Introducing the distinction of type (demonstrative,
+amplifier, quantifier, possessive or interrogative) led to an increase in
+tagger accuracy of 0.14 percentage points to 97.61\%, while marginally
 impacting the parsing, with LAS of 87.00\%, 0.01 percentage points below that
@@
 distinctions for determiners, which is a quite disparate category in the
-treebank, may be quite useful for tagging. However, as it has negative impact
-on parsing, we can conclude that the type of a determiner does not assist in
-generalizing syntactic patterns. Most determiners, regardless of type, appear
-in the same syntactic constructions (i.e., before a noun or noun phrase).
+treebank, may be quite useful for tagging.
 
-Gender, on the other hand, improved the parsing, but complicated the tagging,
-as the various genders are often difficult to differentiate, especially so in
-the case of masculine and feminine, which share many of the same determiners.
-The number of a determiner, i.e., singular or plural, led to a small increase
-in tagger accuracy and LAS, while marginally lower UAS, 90.18\%, 0.01
-percentage points lower than that of the original tag set. The
+%However, as it has negative impact
+%on parsing, we can conclude that the type of a determiner does not assist in
+%generalizing syntactic patterns. Most determiners, regardless of type, appear
+%in the same syntactic constructions (i.e., before a noun or noun phrase).
+
+Gender, on the other hand, improved the parsing (87.09\% LAS), but complicated
+the tagging, as the various genders are often difficult to differentiate,
+especially so in the case of masculine and feminine, which share many of the
+same determiners. The number of a determiner, i.e., singular or plural, led to
+a small increase in tagger accuracy and LAS, while marginally lower UAS of
+90.18\%, 0.01 percentage points lower than that of the original tag set. The
 introduction of definiteness to the determiners led to the best parsing
@@
 data. As this accounts for a very small number of tokens, we did not consider
-further fine-grained modifications with definiteness. This goes to show that
-tokens with overt definiteness have noticeable impact on the syntactic parsing,
-and that distinguishing on definiteness is very beneficial.
+further fine-grained modifications with definiteness. This serves to indicate
+that tokens with overt definiteness have noticeable impact on the syntactic
+parsing and that distinguishing on definiteness is very useful.
 
@@
 \paragraph{Pronouns}
-Pronouns in Norwegian include personal, reciprocal, reflexive and
-interrogative. They can exhibit gender, number and person. Personal pronouns
-have case (accusative or nominative).
+Pronouns in Norwegian are personal, reciprocal, reflexive and interrogative.
+They can furthermore exhibit gender, number and person, while personal pronouns
+can be distinguished case (accusative or nominative).
 

### 2016-05-06 submission.tex
changes to paragraph on pronouns in tag set experiments

@@
 They can furthermore exhibit gender, number and person, while personal pronouns
-can be distinguished case (accusative or nominative).
+can be distinguished on case (either accusative or nominative).
 
@@
 and number combined, obtaining a tagger accuracy of 97.52\%. This shows that
-the introduction of more fine-grained distinctions for pronouns is beneficial
-and aids the PoS tagger in disambiguating ambiguous words. While case alone
-yields an LAS of 87.08\%, we found that the combination of type and case, which
-is the most successful experiment in terms of parser performance, yields the
-second highest tagging accuracy of 97.51\%. The reason for this is that
-pronouns of different type and personal pronouns of different case exhibit
-quite different properties and appear in different constructions. Pronouns in
-nominative case (i.e., subjects) primarily occur before the main verb, while
-pronouns in accusative case (i.e., objects) occur after the main verb, as
-Norwegian exhibits so-called V2 word order, requiring that the finite verb of a
-declarative clause appears in the second position, hence its name. The
-combination of type and number comes in close to the performance of type and
-case, with LAS of 87.27\% and UAS identical to that of type and case.
+the introduction of more fine-grained distinctions for pronouns aids the PoS
+tagger in disambiguating ambiguous words. While case alone yields an LAS of
+87.08\%, we found that the combination of type and case, which is the most
+successful experiment in terms of parser performance, yields the second highest
+tagging accuracy of 97.51\%. Pronouns of different type and personal pronouns
+of different case exhibit quite different properties and appear in different
+constructions. Pronouns in nominative case (i.e., subjects) primarily occur
+before the main verb, while pronouns in accusative case (i.e., objects) occur
+after the main verb, as Norwegian exhibits so-called V2 word order, requiring
+that the finite verb of a declarative clause appears in the second position,
+hence its name. The combination of type and number comes in close to the
+performance of type and case, with LAS of 87.27\% and UAS identical to that of
+type and case.
 

### 2016-05-06 submission.tex
changes to section optimized tag set

@@
 
-The most successful tag set modification for each PoS and their results are
-seen in Table \ref{respectiveresults}. Nouns benefit by far the most from the
-introduction of more fine-grained linguistically motivated distinctions, with
-an LAS of 88.81\% and UAS of 91.73\%. We observe that the most promising tag
-set modifications for verbs, adjectives, determiners and pronouns all reach LAS
-of \textasciitilde 87.30\% and UAS of \textasciitilde 90.40\%. To investigate
-the overall effect of these tag set modifications, we tested each of the
-improvements in parser accuracy scores from baseline for statistical
-significance using Dan Bikel's randomized parsing evaluation comparator
-script\footnote{Available as \texttt{compare.pl} at
-    \url{http://ilk.uvt.nl/conll/software.html}}, as used in the CoNLL shared
-tasks. For the most successful tag set modification for each of the categories
-seen in Table \ref{respectiveresults}, the difference in LAS from the original
-tag set is statistically significant at significance level 0.05 (\emph{p}-value
-< 0.05), as are all differences in UAS, except for verbs with finiteness
-(\emph{p}-value 0.15) and pronouns with type and case (\emph{p}-value 0.06).
+The most successful tag set modification for each PoS and the results from
+tagging and parsing with the respective modifications are seen in Table
+\ref{respectiveresults}. Nouns benefit by far the most from the introduction of
+more fine-grained linguistically motivated distinctions, with an LAS of 88.81\%
+and UAS of 91.73\% when distinguishing on type, case and definiteness. We
+observe that the most promising tag set modifications for verbs, adjectives,
+determiners and pronouns all reach LAS of \textasciitilde 87.30\% and UAS of
+\textasciitilde 90.40\%. To investigate the overall effect of these tag set
+modifications, we tested each of the improvements in parser accuracy scores
+from that of the original tag set for statistical significance using Dan
+Bikel's randomized parsing evaluation comparator script\footnote{Available as
+    \texttt{compare.pl} at \url{http://ilk.uvt.nl/conll/software.html}}, as
+used in the CoNLL shared tasks. For the most successful tag set modification
+for each of the categories seen in Table \ref{respectiveresults}, the
+difference in LAS from the original tag set is statistically significant at
+significance level 0.05 (\emph{p}-value < 0.05), as are all differences in UAS,
+except for verbs with finiteness (\emph{p}-value 0.15) and pronouns with type
+and case (\emph{p}-value 0.06).
 
@@
 In Table \ref{finalresults}, we show the results of parsing with the optimized
-tag set on the held-out test set, compared to the results obtained with the
-original tag set. We see significant improvements from the original tag set on
-both the development data and the held-out test data set. The improvement in
-LAS on the development data is 1.86 percentage points, while 1.91 percentage
-points on the held-out test data. These results indicate that the additional
-linguistic information in the tags of our optimized tag set greatly aids
-syntactic parsing and that optimizing an existing PoS tag set for a downstream
-application can be useful and beneficial.
+tag set on the held-out test data and the development data, compared to the
+results obtained with the original tag set. We see significant improvements
+from the original tag set on both the development data and the held-out test
+data set. The improvement in LAS on the development data is 1.86 percentage
+points, while 1.91 percentage points on the held-out test data. These results
+indicate that the additional linguistic information in the tags of our
+optimized tag set greatly aids syntactic parsing and that optimizing an
+existing PoS tag set for a downstream application can be useful and beneficial.
 

### 2016-05-06 submission.tex
endringer i sec:optimization

@@
 set, we increase the linguistic information represented in the tags, which may
-assist the parser in recognizing and generalizing syntactic patterns. However,
-the addition of more linguistic information to the tags and thus a more
-fine-grained tag set will most likely lead to a drop in tagger accuracy due to
-the increase in complexity. The best tagging does not necessarily lead to the
-best parse, hence it is interesting to investigate how linguistically informed
-tag set modifications may affect the interplay between tagging and parsing.
-%as
-%we want to investigate how we can improve the linguistic quality of a tag set,
-%we will
-%only consider tag set modifications we deem linguistically sensible.
-%we will not introduce tag set modifications which are not linguistically
-%motivated. we will only consider distinctions we deem linguistically sensible.
+assist the parser in recognizing and generalizing syntactic patterns. 
+We would then also increase the complexity of the tagging task though,  
+which can be expected to lead to a drop in tagger accuracy. 
+However, the best tagging, evaluated in isolation, does not necessarily lead to the
+best parse, hence it is interesting to investigate how tag set modifications
+may affect this interplay.
 
@@
 In an initial round of experiments, we concatenated the tag of each token with
-its set of morphological features in order to map the original tag set to a
-new, more fine-grained tag set (hereafter referred to as the \emph{full} tag
-set). This resulted in a total of 368 tags, which is clearly very fine-grained.
+its full set of morphological features, thereby mapping the original tag set to a
+new maximally fine-grained tag set given the annotations available in the
+treebank. This resulted in a total of 368 tags, 
+hereafter referred to as the \emph{full} tag set.  
 The two initial tag sets, i.e., the original tag set comprising 19 tags and the
 full tag set comprising 368 tags, thus represent two extremes in terms of
-granularity. To evaluate the initial tag sets and investigate how the tag set
-granularity would affect the performance of tagging and parsing, we ran
-baseline experiments with the two tag sets. In Table \ref{inittagseteval}, we
-report the results of these experiments. We see that the tagger accuracy
-drastically drops when going from the original to the full tag set. The MFT
+granularity. 
+To establish some initial points of reference for how tag set granularity affects the 
+performance of tagging and parsing on NDT, we trained and tested a full pipeline with both of
+these intital tags sets. The results are reported in Table \ref{inittagseteval}.
+Unsurprisingly, we see that the tagger accuracy
+plummets when we move from the original to the full tag set. The MFT
 baseline for the original tag set is 94.14\%, while it drops by almost 9
@@
 97.47\% on the original tag set, which is reduced to 93.48\% for the full tag
-set. These results confirm our hypothesis that the very high linguistic quality
-in the full, fine-grained tag set comes at the expense of drops in tagger
-performance.  However, the additional linguistic information provided by the
-full tag set improves the parser performance. With the original tag set, Mate
-reports an LAS of 87.01\% and a UAS of 90.19\%, which increases to 87.15\% and
+set. These results confirm our hypothesis that the very high level linguistic
+information 
+in the full, fine-grained tag set comes at the expense of reduced tagger
+performance.  In spite of this, however, we see that the additional information
+in the full tag set still improves the parser performance. With the original tag set, Mate
+yields 87.01\% LAS and 90.19\% UAS, which increases to 87.15\% and
 90.39\%, respectively, using the full tag set. As we are looking for
 linguistically informed distinctions that improve the syntactic parsing, these
-results are promising and serve to indicate that additional linguistic
-information assists the syntactic parsers, motivating the optimization of the
+results are promising and indicate that additional linguistic
+information assists syntactic parsing, motivating the optimization of the
 existing PoS tag set.
@@
     \end{tabular}
-    \caption{Results of tagging and parsing with the two initial tag sets.
-        \emph{Acc} denotes tagger accuracy.}
+    \caption{Results of tagging and parsing our development section of NDT with
+      the two initial tag sets. From left to right we report the tagger
+      accuracy of the most-frequent-tag baseline, the tagger accuracy of TnT,
+      and labeled / unlabeled attachment scores for the Mate parser.} 
     \label{inittagseteval}
@@
 \subsection{Tag Set Experiments}
-We modify the tags for nouns (\texttt{subst}), verbs (\texttt{verb}),
+We modified the tags for nouns (\texttt{subst}), verbs (\texttt{verb}),
 adjectives (\texttt{adj}), determiners (\texttt{det}) and pronouns
@@
 each tag in order to increase the linguistic information expressed by the tags.
-For each tag, we first experiment with each of the available features in
-isolation before employing various combinations of them. We base our choices of
-combinations on how promising the features are and what we deem worth
-investigating in terms of linguistic utility, in order to see how the features
-might interact.
+For each tag in turn, we first experiment with each of the available features in
+isolation before testing various combinations in order to see how the features
+might interact. We base our choices of combinations on how promising the
+features are and what we deem worth investigating in terms of linguistic utility.
 

### 2016-05-06 submission.tex
minor editing on baseline experiments

@@
 performance of tagging and parsing on NDT, we trained and tested a full pipeline with both of
-these intital tags sets. The results are reported in Table \ref{inittagseteval}.
+these initial tags sets. The results are reported in Table \ref{inittagseteval}.
 Unsurprisingly, we see that the tagger accuracy
@@
 97.47\% on the original tag set, which is reduced to 93.48\% for the full tag
-set. These results confirm our hypothesis that the very high level linguistic
+set. These results confirm our hypothesis that the very high level of linguistic
 information

### 2016-05-06 submission.tex
minor changes to baseline experiments

@@
 97.47\% on the original tag set, which is reduced to 93.48\% for the full tag
-set. These results confirm our hypothesis that the very high level of linguistic
+set. These results confirm our hypothesis that the very high level linguistic
 information

### 2016-05-06 submission.tex

--- 2016-05-06 submission.tex
@@
 each tag in order to increase the linguistic information expressed by the tags.
-For each tag in turn, we first experiment with each of the available features in
-isolation before testing various combinations in order to see how the features
-might interact. We base our choices of combinations on how promising the
-features are and what we deem worth investigating in terms of linguistic utility.
+For each tag, we in turn first experiment with each of the available features in
+isolation before testing various combinations. We base our choices
+of combinations on how promising the features are and what we deem worth
+For each tag, we first experiment with each of the available features in
+isolation before employing various combinations of them. We base our choices of
+combinations on how promising the features are and what we deem worth
+investigating in terms of linguistic utility, in order to see how the features
+might interact.
 

### 2016-05-06 submission.tex
changes to tag set mapping

@@
 \paragraph{Tag Set Mapping}
-The tag set modifications are realized by mapping the existing tags to new,
-more fine-grained tags "based" on associated morphological features to be
-included in specified tags. The procedure maps the relevant existing tags to
-new, more fine-grained tags including more relevant morphological features for
-the applicable tokens, i.e., tokens that are assigned said tag and feature(s)
-in the gold standard data. Hence, we are in actuality altering the gold
-standard tags by the addition of more relevant linguistic information,
-producing new gold standard annotations of higher quality.
-
-% fixme: tror nok dette bør flyttes og kortes ned. kan bakes inn med det som
-% står under "Tag Set Mapping".
-Our approach is somewhat semi-automatic in that we employ a "hybrid"
-combination of manual annotation and automatic labeling in our experiments. Our
-initial tags are gold standard, and using linguistic insight coupled with
-computational considerations, we introduce new gold standard distinctions to
-the tag set. The data sets with these tags are run through a pipeline with
-automatically assigned tags in both training and testing, where the most
-promising modifications are used "further". Hence, the tag set modifications
-are not deterministic, nor do we try all possible combination; we only
-experiment with linguistically motivated distinctions.
-
-As we want to investigate how we can improve the linguistic quality of a tag
-set, we will only consider tag set modifications we deem linguistically
-sensible.
-%This is done by concatenating the original tag with the
-%specified features for all applicable tokens
-
-%contains the set of features in its morphological features.
+The tag set modifications are realized through a mapping procedure that
+concatenates the existing tag with one or more specified features for all
+applicable tokens, i.e., tokens annotated with said tag and feature(s) in the
+gold standard data.
+%A given tag set modification involves the specification of
+%a tag and the desired features we want to append to the existing,
+%coarse-grained tag.
 
@@
 may affect this interplay.
+
+Hence, we are in actuality altering the gold
+standard tags by the addition of more relevant linguistic information,
+producing new gold standard annotations of higher quality.
+
+% fixme: tror nok dette bør flyttes og kortes ned. kan bakes inn med det som
+% står under "Tag Set Mapping".
+Our approach is somewhat semi-automatic in that we employ a "hybrid"
+combination of manual annotation and automatic labeling in our experiments. Our
+initial tags are gold standard, and using linguistic insight coupled with
+computational considerations, we introduce new gold standard distinctions to
+the tag set. The data sets with these tags are run through a pipeline with
+automatically assigned tags in both training and testing, where the most
+promising modifications are used "further". Hence, the tag set modifications
+are not deterministic, nor do we try all possible combination; we only
+experiment with linguistically motivated distinctions.
+
+As we want to investigate how we can improve the linguistic quality of a tag
+set, we will only consider tag set modifications we deem linguistically
+sensible.
 

### 2016-05-06 submission.tex
summary

--- 2016-05-06 submission.tex
@@
 
-\section{Summary and Future Work}
+\section{Summary}
 \label{sec:summary}
-The improvements in parser performance with our optimized tag set indicate that
-a more fine-grained PoS tag set may assist syntactic parsers in recognizing and
-generalizing syntactic patterns, while potentially compromising the performance
-of PoS taggers.
-
-There are several aspects of this thesis that can be further explored in future
-work, including extrinsic evaluation of the effects of PoS tag sets on other
-downstream NLP applications besides parsing, such as sentiment analysis and
-named entity recognition. These applications often require tagged data, but are
-markedly different from syntactic parsing, hence the evaluation would involve
-investigating an entirely different aspect of the effects of tag set
-granularity.
+This paper has reported on a range of experiments with injecting more
+fine-grained morphological distinctions into an existing PoS tag set, and then
+empirically evaluating the effects both (intrinsically) in terms of tagging
+accuracy and (extrinsically) in terms of parsing accuracy. Our experimental
+results -- based on the annotations of the Norwegian Dependency Treebank using
+the TnT PoS tagger and the Mate dependency parser -- show that the enriched tag
+set leads to significantly improved parsing accuracy, even though PoS tagging
+accuracy in isolation is reduced. The modified tag sets in our experiments are
+defined as combinations of PoS tags and morphological features already
+available in the gold annotations of the treebank. The best performing tag
+configuration is in effect a PoS tag set optimized for dependency parsing of
+Norwegian. However, by using another task for the extrinsic evaluation one can
+of course optimize a tag set specifically for other downstream applications
+instead, such as sentiment analysis, named entity recognition or any other task
+making use of tagged data.
 

### 2016-05-06 submission.tex
Tag Set Mapping

@@
 \paragraph{Tag Set Mapping}
-The tag set modifications are realized through a mapping procedure that
-concatenates the existing tag with one or more specified features for all
-applicable tokens, i.e., tokens annotated with said tag and feature(s) in the
-gold standard data.
+Based only on information already available in the gold annotations of the
+treebank, the tag set modifications are realized through a mapping procedure that
+concatenates existing tags with one or more specified morphological features for all
+applicable tokens.
+ %, i.e., tokens annotated with said tag and feature(s) in the
+% gold standard data.
 %A given tag set modification involves the specification of
@@
     \caption{Results of tagging and parsing with the optimized tag set,
-        compared to the initial tag sets. Trained and tested using
+        compared to the original coarse tag set. Trained and tested using
         automatically assigned tags from TnT.}
@@
 accuracy and (extrinsically) in terms of parsing accuracy. Our experimental
-results -- based on the annotations of the Norwegian Dependency Treebank using
+results -- based on the annotations of the Norwegian Dependency Treebank and using
 the TnT PoS tagger and the Mate dependency parser -- show that the enriched tag

### 2016-05-07 submission.tex
moving tables around
minor changes to caption in table 4

@@
     \caption{Results of tagging and parsing our development section of NDT with
-      the two initial tag sets. From left to right we report the tagger
+        the two initial tag sets. From left to right, we report the tagger
         accuracy of the most-frequent-tag baseline, the tagger accuracy of TnT,
-      and labeled / unlabeled attachment scores for the Mate parser.}
+        and the labeled attachment score and the unlabeled attachment score for
+        the Mate parser.}
     \label{inittagseteval}
@@
 
-%As nouns constitute the largest class in the treebank by far, the effects of
-%the tag set modifications are greatest for noun tokens in terms of overall
-%change in performance.
-The results of experimenting with tag set modifications
-for nouns are shown in Table \ref{substresults} and reveal that, apart from
-case, none of the tag set modifications improve the tagging. However, they all
-give rise to increases in parser accuracy scores.
-The most informative features are definiteness, which
-leads to an increase in LAS by 1.26 percentage points to 88.27\%, and type,
-yielding an LAS of 88.07\%. Turning to combinations of features, we found that
-the combination of type and case, as well as type and definiteness, were the
-most promising, which led us to combine type, case and definiteness in a final
-experiment, resulting in LAS of 88.81\% and UAS of 91.73\%, constituting large
-improvements from parsing with the original tag set, of 1.80 percentage points
-and 1.54 percentage points, respectively.
-
 \begin{table}
@@
 
+The results of experimenting with tag set modifications
+for nouns are shown in Table \ref{substresults} and reveal that, apart from
+case, none of the tag set modifications improve the tagging. However, they all
+give rise to increases in parser accuracy scores.
+The most informative features are definiteness, which
+leads to an increase in LAS by 1.26 percentage points to 88.27\%, and type,
+yielding an LAS of 88.07\%. Turning to combinations of features, we found that
+the combination of type and case, as well as type and definiteness, were the
+most promising, which led us to combine type, case and definiteness in a final
+experiment, resulting in LAS of 88.81\% and UAS of 91.73\%, constituting large
+improvements from parsing with the original tag set, of 1.80 percentage points
+and 1.54 percentage points, respectively.
+
 \paragraph{Verbs}
@@
 in an LAS of 87.12\% and UAS of 90.31\%.
-
-In an additional experiment, we mapped the verb tenses (and mood, in the case
-of imperative) to finiteness. All verbs have finiteness, hence this distinction
-has broad coverage. This mapping is syntactically grounded as finite verbs and
-nonfinite verbs appear in completely different syntactic constructions, and
-proved to greatly improve the parsing, as we saw the highest parser accuracy
-scores of 87.30\% for LAS and 90.43\% for UAS, 0.29 and 0.24 percentage points
-higher than the baseline, respectively. This coincides with the observations
-seen for Swedish in \newcite{Ovr:08}, where finiteness was found
-to be a very beneficial linguistic feature for parsing.
 
@@
 \end{table}
+
+In an additional experiment, we mapped the verb tenses (and mood, in the case
+of imperative) to finiteness. All verbs have finiteness, hence this distinction
+has broad coverage. This mapping is syntactically grounded as finite verbs and
+nonfinite verbs appear in completely different syntactic constructions, and
+proved to greatly improve the parsing, as we saw the highest parser accuracy
+scores of 87.30\% for LAS and 90.43\% for UAS, 0.29 and 0.24 percentage points
+higher than the baseline, respectively. This coincides with the observations
+seen for Swedish in \newcite{Ovr:08}, where finiteness was found
+to be a very beneficial linguistic feature for parsing.
 
@@
 
-An overview of the tags in the optimized tag set can be seen in Table
-\ref{optimizedtagset}, comprising three new tags for adjectives, two for
-determiners, six for pronouns, seven for nouns and two for verbs, totaling 20
-tags. Appending these to the original tag set comprising 19 tags, we reach a
-total of 39 tags for NDT.
-
 \begin{table}
@@
 \end{table}
+
+An overview of the tags in the optimized tag set can be seen in Table
+\ref{optimizedtagset}, comprising three new tags for adjectives, two for
+determiners, six for pronouns, seven for nouns and two for verbs, totaling 20
+tags. Appending these to the original tag set comprising 19 tags, we reach a
+total of 39 tags for NDT.
 

### 2016-05-07 submission.tex
misc

--- 2016-05-06 submission.tex
@@
 
-We argue, however, that it is possible to significantly improve the performance
+In the current paper, we argue that it is neverthelesss possible to significantly improve the performance
 of the downstream application of syntactic parsing by introducing
 linguistically informed, syntactically informative distinctions to a tag set.
-We will now turn to the data that provided the basis for our experiments and
+In the following section we first turn to the treebank that provides the basis for our experiments and
 outline its morphological and syntactic annotation choices.
@@
 excerpts. The annotation process was supported by the rule-based Oslo-Bergen
-Tagger \cite{Hag:Joh:Nok:00} and then manually corrected by annotators, adding
-syntactic dependency analysis to the morphosyntactic annotation.
+Tagger \cite{Hag:Joh:Nok:00} and then manually corrected by human annotators, also adding
+syntactic dependency analyses to the morphosyntactic annotation.
 
@@
 Unsurprisingly, we see that the tagger accuracy
-plummets when we move from the original to the full tag set. The MFT
-baseline for the original tag set is 94.14\%, while it drops by almost 9
+plummets when we move from the original to the full tag set. While the MFT
+tagging baseline for the original tag set is 94.14\%, it drops by almost 9
 percentage points to 85.15\% for the full tag set. TnT reports an accuracy of

### 2016-05-07 submission.tex
Tag set optimization

@@
 
-\paragraph{Tags \& Features} % <-- fixme: rename?
+\paragraph{Predicted vs gold tags}
 As we seek to quantify the effects of PoS tagging in a realistic setting, i.e.,
@@
 assigned PoS tags. For training, however, we have two options: using either
-gold standard or automatically assigned tags. In order to settle on a
-configuration, we conducted experiments with training on both gold standard
-tags and automatically assigned tags to see how the performance differs. The
+gold standard tags or automatically predicted tags. In order to settle on a
+configuration, we conducted experiments with training on both gold tags 
+and predicted tags to see how the performance differs. The
 results of these experiments, shown in Table \ref{tagconfigeval}, reveal that
-the combination of training and testing on automatic tags is clearly superior
-to training on gold standard tags and testing on automatic tags. Consequently,
-the parser is both trained and tested on automatically assigned tags in the
-remainder of our experiments.
+the combination of training and testing on predicted tags is clearly superior
+to training on gold standard tags and testing on predicted tags. Consequently,
+the parser is both trained and tested on automatically predicted tags in the
+remainder of our experiments.\footnote{Though some parsers also make use of
+  morphological features, we removed all morphological features beyond the PoS
+  tags in order to simulate a realistic setting.}
 
@@
     \end{tabular}
-    \caption{Results of parsing with Mate using the various tag configurations,
-        either using gold standard tags or automatically assigned from TnT.
-        \emph{Gold} denotes gold standard tags, \emph{Auto} denotes
-        automatically assigned tags from TnT.}
+    \caption{Results of parsing with Mate using various configurations of PoS
+      tag sources in training and testing: \emph{Gold} denotes gold standard
+      tags and \emph{Auto} denotes tags automatically predicted by TnT.}
     \label{tagconfigeval}
@@
 
-Though some parsers make use of morphological features, % fixme: status for mate?
-we removed all morphological features beyond the PoS tags in order to simulate
-a realistic setting.  Moreover, it is crucial that we remove these features
-when working with automatically assigned tags, as such features would be
-provided by the gold standard.
+% Moreover, it is crucial that we remove these features
+% when working with automatically assigned tags, as such features would be
+% provided by the gold standard.
 
@@
 
-In the following, we report on a set of experiments which evaluate increasingly
-fine-grained PoS tag sets. The PoS tag sets are motivated by morphosyntactic
-properties of Norwegian and are evaluated in terms of both tagger and parser
-accuracy.
+% In the following, we report on a set of experiments which evaluate increasingly
+% fine-grained PoS tag sets. The PoS tag sets are motivated by morphosyntactic
+% properties of Norwegian and are evaluated in terms of both tagger and parser
+% accuracy.
 
@@
 \label{sec:optimization}
-By introducing more fine-grained linguistically motivated distinctions in a tag
+
+The modified tag sets used in the experiments reported in this section are
+defined as combinations of PoS tags and morphological features already
+available in the gold annotations of the treebank. In effect, we redefine the
+tags comprising the gold standard as provided by NDT. The best performing
+configuration can be seens as tag set specifically optimized for dependency
+parsing. Note that the tag selection process itself can be seen as
+semi-automatic; while we for each introduced distinction empirically evaluate
+its impact on parsing performance -- as to see whether it is worth including in
+the final configuration -- the process is manually guided by linguistic
+considerations regarding which combinations to evaluate in the first place.
+Although our expressed goal is to identify a tag set optimized for the
+downstream task of dependency parsing, we will only consider tag set
+modifications we deem linguistically sensible.
+
+By introducing more fine-grained linguistically motivated distinctions in the tag
 set, we increase the linguistic information represented in the tags, which may
 assist the parser in recognizing and generalizing syntactic patterns.
-We would then also increase the complexity of the tagging task though,
+However, this would also increase the complexity of the tagging task, 
 which can be expected to lead to a drop in tagger accuracy.
@@
 best parse, hence it is interesting to investigate how tag set modifications
-may affect this interplay.
+affect this interplay.
 
-Hence, we are in actuality altering the gold
-standard tags by the addition of more relevant linguistic information,
-producing new gold standard annotations of higher quality.
+% Hence, we are in actuality altering the gold
+% standard tags by the addition of more relevant linguistic information,
+% producing new gold standard annotations of higher quality.
 
@@
 % står under "Tag Set Mapping".
-Our approach is somewhat semi-automatic in that we employ a "hybrid"
-combination of manual annotation and automatic labeling in our experiments. Our
-initial tags are gold standard, and using linguistic insight coupled with
-computational considerations, we introduce new gold standard distinctions to
-the tag set. The data sets with these tags are run through a pipeline with
-automatically assigned tags in both training and testing, where the most
-promising modifications are used "further". Hence, the tag set modifications
-are not deterministic, nor do we try all possible combination; we only
-experiment with linguistically motivated distinctions.
-
-As we want to investigate how we can improve the linguistic quality of a tag
-set, we will only consider tag set modifications we deem linguistically
-sensible.
+% Our approach is somewhat semi-automatic in that we employ a "hybrid"
+% combination of manual annotation and automatic labeling in our experiments. Our
+% initial tags are gold standard, and using linguistic insight coupled with
+% computational considerations, we introduce new gold standard distinctions to
+% the tag set. The data sets with these tags are run through a pipeline with
+% automatically assigned tags in both training and testing, where the most
+% promising modifications are used "further". Hence, the tag set modifications
+% are not deterministic, nor do we try all possible combination; we only
+% experiment with linguistically motivated distinctions.
 
@@
 its full set of morphological features, thereby mapping the original tag set to a
-new maximally fine-grained tag set given the annotations available in the
+new maximally fine-grained tag set, given the annotations available in the
 treebank. This resulted in a total of 368 tags,

### 2016-05-07 submission.tex
captions

@@
 compromising tagger accuracy, actually leads to significantly improved parsing
-accuracy. This optimization allows us to bypass the additional step of
+accuracy. This optimization also allows us to bypass the additional step of
 morphological analysis, framing the whole preprocessing problem as a simple
@@
 in application to raw text, we choose to evaluate the parser on automatically
-assigned PoS tags. For training, however, we have two options: using either
+predicted PoS tags. For training, however, we have two options: using either
 gold standard tags or automatically predicted tags. In order to settle on a
@@
 tags comprising the gold standard as provided by NDT. The best performing
-configuration can be seens as tag set specifically optimized for dependency
+configuration can be seen as tag set specifically optimized for dependency
 parsing. Note that the tag selection process itself can be seen as
@@
 
-By introducing more fine-grained linguistically motivated distinctions in the tag
-set, we increase the linguistic information represented in the tags, which may
-assist the parser in recognizing and generalizing syntactic patterns.
-However, this would also increase the complexity of the tagging task, 
-which can be expected to lead to a drop in tagger accuracy.
-However, the best tagging, evaluated in isolation, does not necessarily lead to the
-best parse, hence it is interesting to investigate how tag set modifications
-affect this interplay.
+The hope is that the introduction of more fine-grained linguistically motivated
+distinctions in the tag set may assist the parser in recognizing and
+generalizing syntactic patterns.  However, this would also increase the
+complexity of the tagging task, which can be expected to lead to a drop in
+tagger accuracy.  However, the best tagging, evaluated in isolation, does not
+necessarily lead to the best parse, hence the aim of this section is to
+investigate how tag set modifications affect this interplay.
 
@@
 yields 87.01\% LAS and 90.19\% UAS, which increases to 87.15\% and
-90.39\%, respectively, using the full tag set. As we are looking for
-linguistically informed distinctions that improve the syntactic parsing, these
-results are promising and indicate that additional linguistic
-information assists syntactic parsing, motivating the optimization of the
+90.39\%, respectively, using the full tag set. These
+preliminary results are promising in that thay indicate that additional linguistic
+information assists syntactic parser, motivating further optimization of the
 existing PoS tag set.
@@
     \end{tabular}
-    \caption{Results of experiments with modified PoS tags for nouns.}
+    \caption{Tagging and parsing with modified PoS tags for nouns. The first
+      row corresponds to using the original tag set unmodified.}
     \label{substresults}
@@
     \end{tabular}
-    \caption{Results of experiments with modified PoS tags for verbs.}
+    \caption{Tagging and parsing with modified PoS tags for verbs.}
     \label{verbresults}
@@
     \end{tabular}
-    \caption{Results of experiments with modified PoS tags for adjectives.}
+    \caption{Tagging and parsing with modified PoS tags for adjectives.}
     \label{adjresults}
@@
     \end{tabular}
-    \caption{Results of experiments with modified PoS tags for determiners.}
+    \caption{Tagging and parsing with modified PoS tags for determiners.}
     \label{detresults}
@@
     \end{tabular}
-    \caption{Results of experiments with modified PoS tags for pronouns.}
+    \caption{Tagging and parsing with modified PoS tags for pronouns.}
     \label{pronresults}
@@
     \caption{Results of tagging and parsing with the optimized tag set,
-        compared to the original coarse tag set. Trained and tested using
-        automatically assigned tags from TnT.}
+        compared to the original NDT coarse tag set. The parser is both trained and tested using
+        automatically predicted tags from TnT.}
     \label{finalresults}

### 2016-05-07 submission.tex
litt trimming i baseline-seksjonen

@@
 tagging baseline for the original tag set is 94.14\%, it drops by almost 9
-percentage points to 85.15\% for the full tag set. TnT reports an accuracy of
-97.47\% on the original tag set, which is reduced to 93.48\% for the full tag
-set. These results confirm our hypothesis that the very high level linguistic
+percentage points to 85.15\% for the full tag set. TnT yields 
+97.47\% accuracy on the original tag set, which drops to 93.48\% for the full tag
+set. These results confirm our hypothesis that the high level linguistic
 information
@@
 in the full tag set still improves the parser performance. With the original tag set, Mate
-yields 87.01\% LAS and 90.19\% UAS, which increases to 87.15\% and
-90.39\%, respectively, using the full tag set. These
-preliminary results are promising in that thay indicate that additional linguistic
+achieves 87.01\% LAS and 90.19\% UAS, which for the full tag set increases to 87.15\% and
+90.39\%, respectively. These
+preliminary results are encouraging in indicating that the additional linguistic
 information assists syntactic parser, motivating further optimization of the
-existing PoS tag set.
+tag set.
 

### 2016-05-07 submission.tex
typos

@@
 
-In the current paper, we argue that it is neverthelesss possible to significantly improve the performance
+In the current paper, we argue that it is nevertheless possible to significantly improve the performance
 of the downstream application of syntactic parsing by introducing

### 2016-05-07 submission.tex
fixed typos, minor changes

@@
 both intrinsically, i.e., in terms of PoS tagging accuracy, and extrinsically,
-in terms of parsing accuracy. Our results show that the introduction of
-morphological distinctions not present in the original tag set, whilst
+in terms of syntactic parsing accuracy. Our results show that the introduction
+of morphological distinctions not present in the original tag set, whilst
 compromising tagger accuracy, actually leads to significantly improved parsing
@@
 
-In the current paper, we argue that it is nevertheless possible to significantly improve the performance
-of the downstream application of syntactic parsing by introducing
-linguistically informed, syntactically informative distinctions to a tag set.
-In the following section we first turn to the treebank that provides the basis for our experiments and
-outline its morphological and syntactic annotation choices.
+In the current paper, we argue that it is nevertheless possible to
+significantly improve the performance of the downstream application of
+syntactic parsing by introducing linguistically informed, syntactically
+informative distinctions to a tag set. In the following section, we first turn
+to the treebank that provides the basis for our experiments and outline its
+morphological and syntactic annotation choices.
 
@@
 
-\paragraph{Predicted vs gold tags}
+\paragraph{Predicted vs. Gold Tags}
 As we seek to quantify the effects of PoS tagging in a realistic setting, i.e.,
@@
 tags comprising the gold standard as provided by NDT. The best performing
-configuration can be seen as tag set specifically optimized for dependency
+configuration can be seen as a tag set specifically optimized for dependency
 parsing. Note that the tag selection process itself can be seen as
@@
 complexity of the tagging task, which can be expected to lead to a drop in
-tagger accuracy.  However, the best tagging, evaluated in isolation, does not
+tagger accuracy.  The best tagging, evaluated in isolation, does not
 necessarily lead to the best parse, hence the aim of this section is to
@@
 For each tag, we in turn first experiment with each of the available features in
-isolation before testing various combinations. We base our choices
-of combinations on how promising the features are and what we deem worth
-For each tag, we first experiment with each of the available features in
-isolation before employing various combinations of them. We base our choices of
+isolation before testing various combinations.
+%We base our choices
+%of combinations on how promising the features are and what we deem worth
+We base our choices of
 combinations on how promising the features are and what we deem worth
@@
 \paragraph{Pronouns}
-Pronouns in Norwegian are personal, reciprocal, reflexive and interrogative.
-They can furthermore exhibit gender, number and person, while personal pronouns
-can be distinguished on case (either accusative or nominative).
+Pronouns in Norwegian include personal, reciprocal, reflexive and
+interrogative.  They can furthermore exhibit gender, number and person, while
+personal pronouns can be distinguished on case (either accusative or
+nominative).
 
@@
 modifications, we tested each of the improvements in parser accuracy scores
-from that of the original tag set for statistical significance using Dan
+from those of the original tag set for statistical significance using Dan
 Bikel's randomized parsing evaluation comparator script\footnote{Available as

### 2016-05-07 submission.tex
tabell-fiksing

@@
         \midrule
-        Gold & Gold & 90.15\% & 92.51\% \\
-        Gold & Auto & 85.68\% & 88.98\% \\
-        Auto & Auto & 87.01\% & 90.19\% \\
+        Gold & Gold & 90.15 & 92.51 \\
+        Gold & Auto & 85.68 & 88.98 \\
+        Auto & Auto & 87.01 & 90.19 \\
         \bottomrule
@@
 distinctions in the tag set may assist the parser in recognizing and
-generalizing syntactic patterns.  However, this would also increase the
-complexity of the tagging task, which can be expected to lead to a drop in
-tagger accuracy.  The best tagging, evaluated in isolation, does not
+generalizing syntactic patterns.  This would also increase the
+complexity of the tagging task though, which can be expected to lead to a drop in
+tagger accuracy. However, the best tagging, evaluated in isolation, does not
 necessarily lead to the best parse, hence the aim of this section is to
@@
 97.47\% accuracy on the original tag set, which drops to 93.48\% for the full tag
-set. These results confirm our hypothesis that the high level linguistic
+set. These results confirm our hypothesis that the high level of linguistic
 information
@@
         \midrule
-        Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
-        Full & 85.15\% & 93.48\% & \textbf{87.15\%} & \textbf{90.39\%} \\
+        Original & \textbf{94.14} & \textbf{97.47} & 87.01 & 90.19 \\
+        Full & 85.15 & 93.48 & \textbf{87.15} & \textbf{90.39} \\
         \bottomrule
@@
         \midrule
-        --- & 97.47\% & 87.01\% & 90.19\% \\
-        Case & \textbf{97.48\%} & 87.63\% & 90.72\% \\
-        Definiteness & 97.00\% & 88.27\% & 91.42\% \\
-        Gender & 96.09\% & 87.21\% & 90.36\% \\
-        Number & 96.37\% & 87.97\% & 91.00\% \\
-        Type & 96.92\% & 88.07\% & 91.11\% \\
-        Case \& definiteness & 97.03\% & 88.39\% & 91.44\% \\
-        Type \& case & 96.92\% & 88.46\% & 91.51\% \\
-        Type \& definiteness & 96.99\% & 88.44\% & 91.48\% \\
-        %Type \& number & 96.37\% & 87.95\% & 90.95\% \\
-        Type, case \& definiteness & 97.05\% & \textbf{88.81\%} &
-        \textbf{91.73\%} \\
-        %Type, definiteness \& number & 96.46\% & 88.06\% & 91.12\% \\
+        --- & 97.47 & 87.01 & 90.19 \\
+        Case & \textbf{97.48} & 87.63 & 90.72 \\
+        Definiteness & 97.00 & 88.27 & 91.42 \\
+        Gender & 96.09 & 87.21 & 90.36 \\
+        Number & 96.37 & 87.97 & 91.00 \\
+        Type & 96.92 & 88.07 & 91.11 \\
+        Case \& definiteness & 97.03 & 88.39 & 91.44 \\
+        Type \& case & 96.92 & 88.46 & 91.51 \\
+        Type \& definiteness & 96.99 & 88.44 & 91.48 \\
+        %Type \& number & 96.37 & 87.95 & 90.95 \\
+        Type, case \& definiteness & 97.05 & \textbf{88.81} &
+        \textbf{91.73} \\
+        %Type, definiteness \& number & 96.46 & 88.06 & 91.12 \\
         \bottomrule
@@
         \midrule
-        --- & \textbf{97.47\%} & 87.01\% & 90.19\% \\
-        Mood & 97.43\% & 87.04\% & 90.19\% \\
-        Tense & 97.30\% & 86.97\% & 90.18\% \\
-        Voice & 97.45\% & 86.96\% & 90.09\% \\
-        Mood \& tense & 97.31\% & 87.12\% & 90.31\% \\
-        Voice \& tense & 97.28\% & 86.99\% & 90.15\% \\
-        Mood, tense \& voice & 97.27\% & 86.83\% & 90.05\% \\
-        Finiteness & 97.35\% & \textbf{87.30\%} & \textbf{90.43\%} \\
+        --- & \textbf{97.47} & 87.01 & 90.19 \\
+        Mood & 97.43 & 87.04 & 90.19 \\
+        Tense & 97.30 & 86.97 & 90.18 \\
+        Voice & 97.45 & 86.96 & 90.09 \\
+        Mood \& tense & 97.31 & 87.12 & 90.31 \\
+        Voice \& tense & 97.28 & 86.99 & 90.15 \\
+        Mood, tense \& voice & 97.27 & 86.83 & 90.05 \\
+        Finiteness & 97.35 & \textbf{87.30} & \textbf{90.43} \\
         \bottomrule
@@
         \midrule
-        --- & \textbf{97.47\%} & 87.01\% & 90.19\% \\
-        Definiteness & 96.84\% & 87.14\% & 90.29\% \\
-        Degree & 97.41\% & \textbf{87.29\%} & \textbf{90.44\%} \\
-        Gender & 96.89\% & 87.10\% & 90.25\% \\
-        Number & 96.71\% & 86.99\% & 90.10\% \\
-        Type & 97.40\% & 87.11\% & 90.25\% \\
-        Definiteness \& degree & 96.81\% & 87.23\% & 90.39\% \\
-        Definiteness \& gender & 96.31\% & 87.18\% & 90.39\% \\
-        Definiteness \& number & 96.78\% & 87.27\% & \textbf{90.44\%} \\
-        %Degree \& gender & 96.87\% & 87.00\% & 90.16\% \\
-        %Degree \& number & 96.76\% & 87.13\% & 90.26\% \\
-        %Definiteness, degree \& number & 96.81\% & 87.14\% & 90.30\% \\
+        --- & \textbf{97.47} & 87.01 & 90.19 \\
+        Definiteness & 96.84 & 87.14 & 90.29 \\
+        Degree & 97.41 & \textbf{87.29} & \textbf{90.44} \\
+        Gender & 96.89 & 87.10 & 90.25 \\
+        Number & 96.71 & 86.99 & 90.10 \\
+        Type & 97.40 & 87.11 & 90.25 \\
+        Definiteness \& degree & 96.81 & 87.23 & 90.39 \\
+        Definiteness \& gender & 96.31 & 87.18 & 90.39 \\
+        Definiteness \& number & 96.78 & 87.27 & \textbf{90.44} \\
+        %Degree \& gender & 96.87 & 87.00 & 90.16 \\
+        %Degree \& number & 96.76 & 87.13 & 90.26 \\
+        %Definiteness, degree \& number & 96.81 & 87.14 & 90.30 \\
         \bottomrule
@@
         \midrule
-        --- & 97.47\% & 87.01\% & 90.19\% \\
-        Definiteness & 97.49\% & \textbf{87.30\%} & \textbf{90.42\%} \\
-        Gender & 97.28\% & 87.09\% & 90.31\% \\
-        Number & 97.49\% & 87.04\% & 90.18\% \\
-        Type & \textbf{97.61\%} & 87.00\% & 90.11\% \\
+        --- & 97.47 & 87.01 & 90.19 \\
+        Definiteness & 97.49 & \textbf{87.30} & \textbf{90.42} \\
+        Gender & 97.28 & 87.09 & 90.31 \\
+        Number & 97.49 & 87.04 & 90.18 \\
+        Type & \textbf{97.61} & 87.00 & 90.11 \\
         \bottomrule
@@
         \midrule
-        --- & 97.47\% & 87.01\% & 90.19\% \\
-        Case & 97.50\% & 87.08\% & 90.21\% \\
-        Gender & 97.48\% & 87.06\% & 90.23\% \\
-        Number & 97.49\% & 87.21\% & 90.33\% \\
-        Person & 97.49\% & 87.22\% & 90.32\% \\
-        Type & 97.48\% & 87.19\% & 90.40\% \\
-        Number \& person & 97.49\% & 96.98\% & 90.16\% \\
-        Type \& case & 97.51\% & \textbf{87.30\%} & \textbf{90.41\%} \\
-        Type \& number & 97.49\% & 87.27\% & \textbf{90.41\%} \\
-        Type \& person & 97.49\% & 87.00\% & 90.14\% \\
-        Type, case \& number & \textbf{97.52\%} & 87.11\% & 90.36\% \\
+        --- & 97.47 & 87.01 & 90.19 \\
+        Case & 97.50 & 87.08 & 90.21 \\
+        Gender & 97.48 & 87.06 & 90.23 \\
+        Number & 97.49 & 87.21 & 90.33 \\
+        Person & 97.49 & 87.22 & 90.32 \\
+        Type & 97.48 & 87.19 & 90.40 \\
+        Number \& person & 97.49 & 96.98 & 90.16 \\
+        Type \& case & 97.51 & \textbf{87.30} & \textbf{90.41} \\
+        Type \& number & 97.49 & 87.27 & \textbf{90.41} \\
+        Type \& person & 97.49 & 87.00 & 90.14 \\
+        Type, case \& number & \textbf{97.52} & 87.11 & 90.36 \\
         \bottomrule
@@
         \midrule
-        \emph{Original} & --- & 94.14\% & 97.47\% & 87.01\% & 90.19\% \\
-        Noun & Type, case \& definiteness & 89.61\% & 97.05\% & 88.81\% &
-        91.73\% \\
-        Verb & Finiteness & 93.72\% & 97.35\% & 87.30\% & 90.43\% \\
-        Adjective & Degree & 94.13\% & 97.41\% & 87.29\% & 90.44\% \\
-        Determiner & Definiteness & 94.13\% & 97.49\% & 87.30\% & 90.42\% \\
-        Pronoun & Type \& case & 94.12\% & 97.51\% & 87.30\% & 90.41\% \\
+        \emph{Original} & --- & 94.14 & 97.47 & 87.01 & 90.19 \\
+        Noun & Type, case \& definiteness & 89.61 & 97.05 & 88.81 &
+        91.73 \\
+        Verb & Finiteness & 93.72 & 97.35 & 87.30 & 90.43 \\
+        Adjective & Degree & 94.13 & 97.41 & 87.29 & 90.44 \\
+        Determiner & Definiteness & 94.13 & 97.49 & 87.30 & 90.42 \\
+        Pronoun & Type \& case & 94.12 & 97.51 & 87.30 & 90.41 \\
         \bottomrule
@@
         \texttt{pron|pers} & Personal pronoun \\
-        \texttt{pron|pers|akk} & Personal pronoun, accusative \\
-        \texttt{pron|pers|nom} & Personal pronoun, nominative \\
+        \texttt{pron|pers|akk} & Personal pron., accusative \\
+        \texttt{pron|pers|nom} & Personal pron., nominative \\
         \texttt{pron|refl} & Reflexive pronoun \\
@@
         \texttt{subst|appell|ub} & Common noun, indef. \\
-        \texttt{subst|appell|ub|gen} & Common noun, indef., genitive \\
+        \texttt{subst|appell|ub|gen} & Common noun, indef., gen. \\
         \texttt{subst|prop} & Proper noun \\
@@
         \multirow{2}{*}{Dev}
-        & Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
+        & Original & \textbf{94.14} & \textbf{97.47} & 87.01 & 90.19 \\
         %Full & 85.15\% & 93.48\% & 87.15\% & 90.39\% \\
-        & Optimized & 85.15\% & 96.85\% & \textbf{88.87\%} & \textbf{91.78\%} \\
+        & Optimized & 85.15 & 96.85 & \textbf{88.87} & \textbf{91.78} \\
         \midrule
         \multirow{2}{*}{Test}
-        & Original & \textbf{94.22\%} & \textbf{97.30\%} & 86.64\% & 90.07\% \\
-        %& Full & 84.27\% & 92.79\% & 87.03\% & 90.16\% \\
-        & Optimized & 88.08\% & 96.35\% & \textbf{88.55\%} & \textbf{91.41\%} \\
+        & Original & \textbf{94.22} & \textbf{97.30} & 86.64 & 90.07 \\
+        %& Full & 84.27\% & 92.79\% & 87.03\% & 90.16 \\
+        & Optimized & 88.08 & 96.35 & \textbf{88.55} & \textbf{91.41} \\
         \bottomrule
